---
title: "Final Project"
author: "Makayla Whitney"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
py_install("cleannlp")
system("pip install cleannlp")
cnlp_init_spacy(model_name = "en")
```

```{r code_setup}

#install.packages("cleanNLP")
library(cleanNLP)
library(dplyr)
library(tidyr)
#install.packages("gtsummary")
library(gtsummary)
library(magrittr)
library(ggplot2)
cnlp_init_udpipe()

```

``` {r test}

#this code chunk was to test and learn the package

annotate_test <- cnlp_annotate(input = c(
  "Here is the first text. It is short.",
  "Here's the second. It is short too!",
  "The third text is the shortest."))


d <- annotate_test$token

d %>% 
  count(xpos) %>% 
  filter(xpos != ".") %>% 
  ggplot(aes(n, xpos)) +
  geom_col()


lapply(annotate_test, head)


cat("This is a normal string in R",
  "This is some text \"That has a quote\" and then some other text",
  sep = "\n")

```

#Introduction

14% of adults in the United States are functionally illiterate. "Functionally illiterate" is defined as reading below the 4th grade reading level. Many adults enroll in literacy tutoring programs that assist them with increasing their reading level to either, 1) complete their GED, 2) enroll in college or enter the workforce, 3) assist their children at home, or 4) other personal learning goals. These agencies are not equip with adult reading materials purely because they don't exist. Instead, agencies provide their tutors with children's books that are complementary to their adult learner's reading level. Children books are not topically engaging or relevant to most adult learners. I examine the parts of speech found within *The Giver* - a book found in most 8th grade curriculum across the country - to help conceptualize what concepts should be taught at the 8th grade level. This project is the beginning of a large-scale examination of 4th through 8th grade books to assist with creating a formula to be used in establishing adult reading materials. 

##Chapter 1

```{r ch1}
annotate_ch1 <- cnlp_annotate(input = c("It was almost December, and Jonas was beginning to be frightened.",
                                        "No. Wrong word, Jonas thought.",
                                        "Frightened meant that deep, sickening feeling of something terrible about to happen.",
                                        "Frightened was the way he had felt a year ago when an unidentified aircraft had overflown the community twice.",
                                        "He had seen it both times.",
                                        "Squinting toward the sky, he had seen the sleek jet, almost a blur at its high speed, go past, and a second later heard the blast of sound that followed.",
                                        "Then one more time, a moment later, from the opposite direction, the same plane.",
                                        "At first, he had been only fascinated.",
                                        "He had never seen aircraft so close, for it was against the rules for Pilots to fly over the community.",
                                        "Occasionally, when supplies were delivered by cargo planes to the landing field across the river, the children rode their bicycles to the riverbank and watched, intrigued, the unloading and then the takeoff directed to the west, always away from the community.",
                                        "But the aircraft a year ago had been different.",
                                        "It was not a squat, fat-bellied cargo plane but a needle-nosed single-pilot jet.",
                                        "Jonas, looking around anxiously, had seen others--adults as well as children--stop what they were doing and wait, confused, for an explanation of the frightening event.",
                                        "Then all of the citizens had been ordered to go into the nearest building and stay there. IMMEDIATELY, the rasping voice through the speakers had said. LEAVE YOUR BICYCLES WHERE THEY ARE.",
                                        "Instantly, obediently, Jonas had dropped his bike on its side on the path behind his family's dwelling. He had run indoors and stayed there, alone. His parents were both at work, and his little sister, Lily, was at the Childcare Center where she spent her after-school hours.",
                                        "Looking through the front window, he had seen no people: none of the busy afternoon crew of Street Cleaners, Landscape Workers, and Food Delivery people who usually populated the community at that time of day. He saw only the abandoned bikes here and there on their sides; an upturned wheel on one was still revolving slowly.",
                                        "He had been frightened then. The sense of his own community silent, waiting, had made his stomach churn. He had trembled.",
                                        "But it had been nothing. Within minutes the speakers had crackled again, and the voice, reassuring now and less urgent, had explained that a Pilot-in-Training had misread his navigational instructions and made a wrong turn. Desperately the Pilot had been trying to make his way back before his error was noticed.",
                                        "NEEDLESS TO SAY, HE WILL BE RELEASED, the voice had said, followed by silence. There was an ironic tone to that final message, as if the Speaker found it amusing; and Jonas had smiled a little, though he knew what a grim statement it had been. For a contributing citizen to be released from the community was a final decision, a terrible punishment, an overwhelming statement of failure.",
                                        "Even the children were scolded if they used the term lightly at play, jeering at a teammate who missed a catch or stumbled in a race.",
                                        "Jonas had done it once, had shouted at his best friend, \"That's it, Asher! You're released!\" when Asher's clumsy error had lost a match for his team. He had been taken aside for a brief and serious talk by the coach, had hung his head with guilt and embarrassment, and apologized to Asher after the game.",
                                        "Now, thinking about the feeling of fear as he pedaled home along the river path, he remembered that moment of palpable, stomach- sinking terror when the aircraft had streaked above. It was not what he was feeling now with December approaching. He searched for the right word to describe his own feeling.",
                                        "Jonas was careful about language. Not like his friend, Asher, who talked too fast and mixed things up, scrambling words and phrases until they were barely recognizable and often very funny.",
                                        "Jonas grinned, remembering the morning that Asher had dashed into the classroom, late as usual, arriving breathlessly in the middle of the chanting of the morning anthem. When the class took their seats at the conclusion of the patriotic hymn, Asher remained standing to make his public apology as was required.",
                                        "\"I apologize for inconveniencing my learning community.\" Asher ran through the standard apology phrase rapidly, still catching his breath. The Instructor and class waited patiently for his explanation. The students had all been grinning, because they had listened to Asher's explanations so many times before.",
                                        "\"I left home at the correct time but when I was riding along near the hatchery, the crew was separating some salmon. I guess I just got distraught, watching them.\"",
                                        "\"I apologize to my classmates,\" Asher concluded. He smoothed his rumpled tunic and sat down.",
                                        "\"We accept your apology, Asher.\" The class recited the standard response in unison. Many of the students were biting their lips to keep from laughing.",
                                        "\"I accept your apology, Asher,\" the Instructor said. He was smiling. \"And I thank you, because once again you have provided an opportunity for a lesson in language. 'Distraught' is too strong an adjective to describe salmon-viewing.\" He turned and wrote \"distraught\" on the instructional board. Beside it he wrote \"distracted.\"",
                                        "Jonas, nearing his home now, smiled at the recollection. Thinking, still, as he wheeled his bike into its narrow port beside the door, he realized that frightened was the wrong word to describe his feelings, now that December was almost here. It was too strong an adjective.",
                                        "He had waited a long time for this special December. Now that it was almost upon him, he wasn't frightened, but he was ... eager, he decided. He was eager for it to come. And he was excited, certainly. All of the Elevens were excited about the event that would be coming so soon.",
                                        "But there was a little shudder of nervousness when he thought about it, about what might happen.",
                                        "Apprehensive, Jonas decided. That's what I am.",
                                        "\"Who wants to be the first tonight, for feelings?\" Jonas's father asked, at the conclusion of their evening meal.",
                                        "It was one of the rituals, the evening telling of feelings.",
                                        "Jonas and his sister, Lily, argued over turns, over who would get to go first. Their parents, of course, were part of the ritual; they, too, told their feelings each evening. But like all parents - all adults - they didn't fight and wheedle for their turn.",
                                        "Nor did Jonas, tonight. His feelings were too complicated this evening. He wanted to share them, but he wasn't eager to begin the process of sifting through his own complicated emotions, even with the help that he knew his parents could give.",
                                        "\"You go, Lily,\" he said, seeing his sister, who was much younger-- only a Seven--wiggling with impatience in her chair.",
                                        "\"I felt very angry this afternoon,\" Lily announced. \"My Childcare group was at the play area, and we had a visiting group of Sevens, and they didn't obey the rules at all. One of them--a male; I don't know his name--kept going right to the front of the line for the slide, even though the rest of us were all waiting. I felt so angry at him. I made my hand into a fist, like this.\" She held up a clenched fist and the rest of the family smiled at her small defiant gesture.",
                                        "\"Why do you think the visitors didn't obey the rules?\" Mother asked.",
                                        "Lily considered, and shook her head. \"I don't know. They acted like...like...\"",
                                        "\"Animals?\" Jonas suggested. He laughed.",
                                        "\"That's right,\" Lily said, laughing too. \"Like animals.\" Neither child knew what the word meant, exactly, but it was often used to describe someone uneducated or clumsy, someone who didn't fit in.",
                                        "\"Where were the visitors from?\" Father asked.",
                                        "Lily frowned, trying to remember. \"Our leader told us, when he made the welcome speech, but I can't remember. I guess I wasn't paying attention. It was from another community. They had to leave very early, and they had their midday meal on the bus.\"",
                                        "Mother nodded. \"Do you think it's possible that their rules may be different? And so they simply didn't know what your play area rules were?\"",
                                        "Lily shrugged, and nodded. \"I suppose.\"",
                                        "\"You've visited other communities, haven't you?\" Jonas asked. \"My group has, often.\"",
                                        "Lily nodded again. \"When we were Sixes, we went and shared a whole school day with a group of Sixes in their community.\"",
                                        "\"How did you feel when you were there?\"",
                                        "Lily frowned. \"I felt strange. Because their methods were different. They were learning usages that my group hadn't learned yet, so we felt stupid.\"",
                                        "Father was listening with interest. \"I'm thinking, Lily,\" he said, \"about the boy who didn't obey the rules today. Do you think it's possible that he felt strange and stupid, being in a new place with rules that he didn't know about?\"",
                                        "Lily pondered that. \"Yes,\" she said, finally.",
                                        "\"I feel a little sorry for him,\" Jonas said, \"even though I don't even know him. I feel sorry for anyone who is in a place where he feels strange and stupid.\"",
                                        "\"How do you feel now, Lily?\" Father asked. \"Still angry?\"",
                                        "\"I guess not,\" Lily decided. \"I guess I feel a little sorry for him. And sorry I made a fist.\" She grinned.",
                                        "Jonas smiled back at his sister. Lily's feelings were always straightforward, fairly simple, usually easy to resolve. He guessed that his own had been, too, when he was a Seven.",
                                        "He listened politely, though not very attentively, while his father took his turn, describing a feeling of worry that he'd had that day at work: a concern about one of the newchildren who wasn't doing well. Jonas's father's title was Nurturer. He and the other Nurturers were responsible for all the physical and emotional needs of every newchild during its earliest life. It was a very important job, Jonas knew, but it wasn't one that interested him much.",
                                        "\"What gender is it?\" Lily asked.",
                                        "\"Male,\" Father said. \"He's a sweet little male with a lovely disposition. But he isn't growing as fast as he should, and he doesn't sleep soundly. We have him in the extra care section for supplementary nurturing, but the committee's beginning to talk about releasing him.\"",
                                        "\"Oh, no,\" Mother murmured sympathetically. \"I know how sad that must make you feel.\"",
                                        "Jonas and Lily both nodded sympathetically as well. Release of newchildren was always sad, because they hadn't had a chance to enjoy life within the community yet. And they hadn't done anything wrong.",
                                        "There were only two occasions of release which were not punishment. Release of the elderly, which was a time of celebration for a life well and fully lived; and release of a newchild, which always brought a sense of what-could-we- have-done. This was especially troubling for the Nurturers, like Father, who felt they had failed somehow. But it happened very rarely.",
                                        "\"Well,\" Father said, \"I'm going to keep trying. I may ask the committee for permission to bring him here at night, if you don't mind. You know what the night-crew Nurturers are like. I think this little guy needs something extra.\"",
                                        "\"Of course,\" Mother said, and Jonas and Lily nodded. They had heard Father complain about the night crew before. It was a lesser job, night-crew nurturing, assigned to those who lacked the interest or skills or insight for the more vital jobs of the daytime hours. Most of the people on the night crew had not even been given spouses because they lacked, somehow, the essential capacity to connect to others, which was required for the creation of a family unit.",
                                        "\"Maybe we could even keep him,\" Lily suggested sweetly, trying to look innocent. The look was fake, Jonas knew; they all knew.",
                                        "\"Lily,\" Mother reminded her, smiling, \"you know the rules.\"",
                                        "Two children--one male, one female--to each family unit. It was written very clearly in the rules.",
                                        "Lily giggled. \"Well,\" she said, \"I thought maybe just this once.\"",
                                        "Next, Mother, who held a prominent position at the Department of Justice, talked about her feelings. Today a repeat offender had been brought before her, someone who had broken the rules before. Someone who she hoped had been adequately and fairly punished, and who had been restored to his place: to his job, his home, his family unit. To see him brought before her a second time caused her overwhelming feelings of frustration and anger. And even guilt, that she hadn't made a difference in his life.",
                                        "\"I feel frightened, too, for him,\" she confessed. \"You know that there's no third chance. The rules say that if there's a third transgression, he simply has to be released.\" Jonas shivered. He knew it happened. There was even a boy in his group of Elevens whose father had been released years before. No one ever mentioned it; the disgrace was unspeakable. It was hard to imagine.",
                                        "Lily stood up and went to her mother. She stroked her mother's arm.",
                                        "From his place at the table, Father reached over and took her hand. Jonas reached for the other.",
                                        "One by one, they comforted her. Soon she smiled, thanked them, and murmured that she felt soothed.",
                                        "The ritual continued. \"Jonas?\" Father asked. \"You're last, tonight.\"",
                                        "Jonas sighed. This evening he almost would have preferred to keep his feelings hidden. But it was, of course, against the rules.",
                                        "\"I'm feeling apprehensive,\" he confessed, glad that the appropriate descriptive word had finally come to him. \"Why is that, son?\" His father looked concerned.",
                                        "\"I know there's really nothing to worry about,\" Jonas explained, \"and that every adult has been through it. I know you have, Father, and you too, Mother. But it's the Ceremony that I'm apprehensive about. It's almost December.\"",
                                        "Lily looked up, her eyes wide. \"The Ceremony of Twelve,\" she whispered in an awed voice. Even the smallest children--Lily's age and younger--knew that it lay in the future for each of them.",
                                        "\"I'm glad you told us of your feelings,\" Father said.",
                                        "\"Lily,\" Mother said, beckoning to the little girl, \"Go on now and get into your nightclothes. Father and I are going to stay here and talk to Jonas for a while.\"",
                                        "Lily sighed, but obediently she got down from her chair. \"Privately?\" she asked.",
                                        "Mother nodded. \"Yes,\" she said, \"this talk will be a private one with Jonas.\"" ))


```


```{r ch1plot}

ch1 <- annotate_ch1$token

ch1 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Quantity",
       y = "Parts of Speech")


lapply(annotate_ch1, head)

#attempting to summarize the percentages of parts of speech present
ch1 %>% 
  select(lemma, upos) %>%
  tbl_summary(ch1)

```
The above bar chart displays a count of each part of speech within Chapter 1 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

*I'm having an issue when running this code; it seems that it wants me to choose a by= that's either lemma or upos, and I'm not familiar enough with NLP in python to make this decsion...*
*That being said, pivoting to horizontal was a great choice for this visualization, and I see that you're planning on adding color, which will definitely help clarify for a reader!*


##Chapter 11
```{r ch11}

annotate_ch11 <- cnlp_annotate(input = c("Jonas felt nothing unusual at first. He felt only the light touch of the old man's hands on his back.",
                                         "He tried to relax, to breathe evenly. The room was absolutely silent, and for a moment Jonas feared that he might disgrace himself now, on the first day of his training, by falling asleep.",
                                         "Then he shivered. He realized that the touch of the hands felt, suddenly, cold. At the same instant, breathing in, he felt the air change, and his very breath was cold. He licked his lips, and in doing so, his tongue touched the suddenly chilled air.",
                                         "It was very startling; but he was not at all frightened, now. He was filled with energy, and he breathed again, feeling the sharp intake of frigid air. Now, too, he could feel cold air swirling around his entire body. He felt it blow against his hands where they lay at his sides, and over his back.",
                                         "The touch of the man's hands seemed to have disappeared.",
                                         "Now he became aware of an entirely new sensation: pinpricks? No, because they were soft and without pain. Tiny, cold, featherlike feelings peppered his body and face. He put out his tongue again, and caught one of the dots of cold upon it. It disappeared from his awareness instantly; but he caught another, and another. The sensation made him smile.",
                                         "One part of his consciousness knew that he was still lying there, on the bed, in the Annex room. Yet another, separate part of his being was upright now, in a sitting position, and beneath him he could feel that he was not on the soft decorated bedcovering at all, but rather seated on a flat, hard surface. His hands now held (though at the same time they were still motionless at his sides) a rough, damp rope.",
                                         "And he could see, though his eyes were closed. He could see a bright, whirling torrent of crystals in the air around him, and he could see them gather on the backs of his hands, like cold fur.",
                                         "His breath was visible.",
                                         "Beyond, through the swirl of what he now, somehow, perceived was the thing the old man had spoken of-- snow-- he could look out and down a great distance. He was up high someplace. The ground was thick with the furry snow, but he sat slightly above it on a hard, flat object.",
                                         "Sled, he knew abruptly. He was sitting on a thing called sled. And the sled itself seemed to be poised at the top of a long, extended mound that rose from the very land where he was. Even as he thought the word \"mound,\" his new consciousness told him hill.",
                                         "Then the sled, with Jonas himself upon it, began to move through the snowfall, and he understood instantly that now he was going downhill. No voice made an explanation. The experience explained itself to him.",
                                         "His face cut through the frigid air as he began the descent, moving through the substance called snow on the vehicle called sled, which propelled itself on what he now knew without doubt to be runners.",
                                         "Comprehending all of those things as he sped downward, he was free to enjoy the breathless glee that overwhelmed him: the speed, the clear cold air, the total silence, the feeling of balance and excitement and peace.",
                                         "Then, as the angle of incline lessened, as the mound--the hill-- flattened, nearing the bottom, the sled's forward motion slowed. The snow was piled now around it, and he pushed with his body, moving it forward, not wanting the exhilarating ride to end.",
                                         "Finally the obstruction of the piled snow was too much for the thin runners of the sled, and he came to a stop. He sat there for a moment, panting, holding the rope in his cold hands. Tentatively he opened his eyes--not his snow-hill-sled eyes, for they had been open throughout the strange ride. He opened his ordinary eyes, and saw that he was still on the bed, that he had not moved at all.",
                                         "The old man, still beside the bed, was watching him. \"How do you feel?\" he asked.",
                                         "Jonas sat up and tried to answer honestly. \"Surprised,\" he said, after a moment.",
                                         "The old man wiped his forehead with his sleeve. \"Whew,\" he said. \"It was exhausting. But you know, even transmitting that tiny memory to you--I think it lightened me just a little.\"",
                                         "\"Do you mean--you did say I could ask questions?\" The man nodded, encouraging his question.",
                                         "\"Do you mean that now you don't have the memory of it--of that ride on the sled--anymore?\"",
                                         "\"That's right. A little weight off this old body.\"",
                                         "\"But it was such fun! And now you don't have it anymore! I took it from you!\"",
                                         "But the old man laughed. \"All I gave you was one ride, on one sled, in one snow, on one hill. I have a whole world of them in my memory. I could give them to you one by one, a thousand times, and there would still be more.\"",
                                         
                                         "\"Are you saying that I--I mean we--could do it again?\" Jonas asked. \"I'd really like to. I think I could steer, by pulling the rope. I didn't try this time, because it was so new.\"",
                                         "The old man, laughing, shook his head. \"Maybe another day, for a treat. But there's no time, really, just to play. I only wanted to begin by showing you how it works.\"",
                                         "\"Now,\" he said, turning businesslike, \"lie back down. I want to --\"",
                                         "Jonas did. He was eager for whatever experience would come next. But he had, suddenly, so many questions.",
                                         "\"Why don't we have snow, and sleds, and hills?\" he asked. \"And when did we, in the past? Did my parents have sleds when they were young? Did you?\"",
                                         "The old man shrugged and gave a short laugh. \"No,\" he told Jonas. \"It's a very distant memory. That's why it was so exhausting--I had to tug it forward from many generations back. It was given to me when I was a new Receiver, and the previous Receiver had to pull it through a long time period, too.\"",
                                         "\"But what happened to those things? Snow, and the rest of it?\"",
                                         "\"Climate Control. Snow made growing food difficult, limited the agricultural periods. And unpredictable weather made transportation almost impossible at times. It wasn't a practical thing, so it became obsolete when we went to Sameness.\"",
                                         "\"And hills, too,\" he added. \"They made conveyance of goods unwieldy. Trucks; buses. Slowed them down. So--\" He waved his hand, as if a gesture had caused hills to disappear. \"Sameness,\" he concluded.",
                                         "Jonas frowned. \"I wish we had those things, still. Just now and then.\"",
                                         "The old man smiled. \"So do I,\" he said. \"But that choice is not ours.\" \"But sir,\" Jonas suggested, \"since you have so much power--\"",
                                         "The man corrected him. \"Honor,\" he said firmly. \"I have great honor. So will you. But you will find that that is not the same as power.\"",
                                         "\"Lie quietly now. Since we've entered into the topic of climate, let me give you something else. And this time I'm not going to tell you the name of it, because I want to test the receiving. You should be able to perceive the name without being told. I gave away snow and sled and downhill and runners by telling them to you in advance.\"",
                                         "Without being instructed, Jonas closed his eyes again. He felt the hands on his back again. He waited.",
                                         "Now it came more quickly, the feelings. This time the hands didn't become cold, but instead began to feel warm on his body. They moistened a little. The warmth spread, extending across his shoulders, up his neck, onto the side of his face. He could feel it through his clothed parts, too: a pleasant, all-over sensation; and when he licked his lips this time, the air was hot and heavy.",
                                        "He didn't move. There was no sled. His posture didn't change. He was simply alone someplace, out of doors, lying down, and the warmth came from far above. It was not as exciting as the ride through the snowy air; but it was pleasurable and comforting.",
                                        "Suddenly he perceived the word for it: sunshine. He perceived that it came from the sky. Then it ended.",
                                        "\"Sunshine,\" he said aloud, opening his eyes.",
                                        "\"Good. You did get the word. That makes my job easier. Not so much explaining.\"",
                                        "\"And it came from the sky.\" \"That's right,\" the old man said. \"Just the way it used to.\"",
                                        "\"Before Sameness. Before Climate Control,\" Jonas added.",
                                        "The man laughed. \"You receive well, and learn quickly. I'm very pleased with you. That's enough for today, I think. We're off to a good start.\"",
                                        "There was a question bothering Jonas. \"Sir,\" he said, \"The Chief Elder told me--she told everyone--and you told me, too, that it would be painful. So I was a little scared. But it didn't hurt at all. I really enjoyed it.\" He looked quizzically at the old man.",
                                        "The man sighed. \"I started you with memories of pleasure. My previous failure gave me the wisdom to do that.\" He took a few deep breaths.\"Jonas,\" he said, \"it will be painful. But it need not be painful yet.\"",
                                        "\"I'm brave. I really am.\" Jonas sat up a little straighter. ",
                                        "The old man looked at him for a moment. He smiled.",
                                        "\"I can see that,\" he said. \"Well, since you asked the question--I think I have enough energy for one more transmission.\"",
                                        "\"Lie down once more. This will be the last today.\"",
                                        "Jonas obeyed cheerfully. He dosed his eyes, waiting, and felt the hands again; then he felt the warmth again, the sunshine again, coming from the sky of this other consciousness that was so new to him. This time, as he lay basking in the wonderful warmth, he felt the passage of time. His real self was aware that it was only a minute or two; but his other, memory-receiving self felt hours pass in the sun. His skin began to sting. Restlessly he moved one arm, bending it, and felt a sharp pain in the crease of his inner arm at the elbow.",
                                        "\"Ouch,\" he said loudly, and shifted on the bed. \"Owwww,\" he said, wincing at the shift, and even moving his mouth to speak made his face hurt.",
                                        "He knew there was a word, but the pain kept him from grasping it.",
                                        "Then it ended. He opened his eyes, wincing with discomfort. \"It hurt,\" he told the man, \"and I couldn't get the word for it.\"",
                                        "\"It was sunburn,\" the old man told him.",
                                        "\"It hurt a lot,\" Jonas said, \"but I'm glad you gave it to me. It was interesting. And now I understand better, what it meant, that there would be pain.\"",
                                        "The man didn't respond. He sat silently for a second. Finally he said, \"Get up, now. It's time for you to go home.\"",
                                        "They both walked to the center of the room. Jonas put his tunic back on. \"Goodbye, sir,\" he said. \"Thank you for my first day.\"",
                                        "The old man nodded to him. He looked drained, and a little sad. \"Sir?\" Jonas said shyly. \"Yes? Do you have a question?\"",
                                        "\"It's just that I don't know your name. I thought you were The Receiver, but you say that now I'm The Receiver. So I don't know what to call you.\"",
                                        "The man had sat back down in the comfortable upholstered chair. He moved his shoulders around as if to ease away an aching sensation. He seemed terribly weary.",
                                        "\"Call me The Giver,\" he told Jonas."))

```

```{r ch11plot}

ch11 <- annotate_ch11$token

ch11 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Quantity",
       y = "Parts of Speech")


lapply(annotate_ch11, head)

```
The above bar chart displays a count of each part of speech within Chapter 11 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

##Chapter 23

```{r ch23}
annotate_ch23 <- cnlp_annotate(input = c("Jonas felt more and more certain that the destination lay ahead of him, very near now in the night that was approaching. None of his senses confirmed it. He saw nothing ahead except the endless ribbon of road unfolding in twisting narrow curves. He heard no sound ahead.",
                                         "Yet he felt it: felt that Elsewhere was not far away. But he had little hope left that he would be able to reach it. His hope diminished further when the sharp, cold air began to blur and thicken with swirling white.",
                                         "Gabriel, wrapped in his inadequate blanket, was hunched, shivering, and silent in his little seat. Jonas stopped the bike wearily, lifted the child down, and realized with heartbreak how cold and weak Gabe had become.",
                                         "Standing in the freezing mound that was thickening around his numb feet, Jonas opened his own tunic, held Gabriel to his bare chest, and tied the torn and dirty blanket around them both. Gabriel moved feebly against him and whimpered briefly into the silence that surrounded them.",
                                         "Dimly, from a nearly forgotten perception as blurred as the substance itself, Jonas recalled what the whiteness was.",
                                         "\"It's called snow, Gabe,\" Jonas whispered. \"Snowflakes. They fall down from the sky, and they're very beautiful.\"",
                                         "There was no response from the child who had once been so curious and alert. Jonas looked down through the dusk at the little head against his chest. Gabriel's curly hair was matted and filthy, and there were tearstains outlined in dirt on his pale cheeks. His eyes were closed. As Jonas watched, a snowflake drifted down and was caught briefly for a moment's sparkle in the tiny fluttering eyelashes.",
                                         "Wearily he remounted the bicycle. A steep hill loomed ahead. In the best of conditions, the hill would have been a difficult, demanding ride. But now the rapidly deepening snow obscured the narrow road and made the ride impossible. His front wheel moved forward imperceptibly as he pushed on the pedals with his numb, exhausted legs. But the bicycle stopped. It would not move.",
                                         "He got off and let it drop sideways into the snow. For a moment he thought how easy it would be to drop beside it himself, to let himself and Gabriel slide into the softness of snow, the darkness of night, the warm comfort of sleep. But he had come this far. He must try to go on.",
                                         "The memories had fallen behind him now, escaping from his protection to return to the people of his community. Were there any left at all? Could he hold onto a last bit of warmth? Did he still have the strength to Give? Could Gabriel still Receive?",
                                         "He pressed his hands into Gabriel's back and tried to remember sunshine. For a moment it seemed that nothing came to him, that his power was completely gone. Then flickered suddenly, and he felt tiny tongues of heat begin to creep across and into his frozen feet and legs. He felt his face begin to glow and the tense, cold skin of his arms and hands relax. For a fleeting second he felt that he wanted to keep it for himself, to let himself bathe in sunlight, unburdened by anything or anyone else.",
                                         "But the moment passed and was followed by an urge, a need, a passionate yearning to share the warmth with the one person left for him to love. Aching from the effort, he forced the memory of warmth into the thin, shivering body in his arms.",
                                         "Gabriel stirred. For a moment they both were bathed in warmth and renewed strength as they stood hugging each other in the blinding snow.",
                                         "Jonas began to walk up the hill.",
                                         "The memory was agonizingly brief. He had trudged no more than a few yards through the night when it was gone and they were cold again.",
                                         "But his mind was alert now. Warming himself ever so briefly had shaken away the lethargy and resignation and restored his will to survive. He began to walk faster on feet that he could no longer feel. But the hill was treacherously steep; he was impeded by the snow and his own lack of strength. He didn't make it very far before he stumbled and fell forward.",
                                         "On his knees, unable to rise, Jonas tried a second time. His consciousness grasped at a wisp of another warm memory, and tried desperately to hold it there, to enlarge it, and pass it into Gabriel. His spirits and strength lifted with the momentary warmth and he stood. Again, Gabriel stirred against him as he began to climb.",
                                         "But the memory faded, leaving him colder than before. If only he had had time to receive more warmth from The Giver before he escaped! Maybe there would be more left for him now. But there was no purpose in if-onlys. His entire concentration now had to be on moving his feet, warming Gabriel and himself, and going forward.",
                                         "He climbed, stopped, and warmed them both briefly again, with a tiny scrap of memory that seemed certainly to be all he had left.",
                                         "The top of the hill seemed so far away, and he did not know what lay beyond. But there was nothing left to do but continue. He trudged upward.",
                                         "As he approached the summit of the hill at last, something began to happen. He was not warmer; if anything, he felt more numb and more cold. He was not less exhausted; on the contrary, his steps were leaden, and he could barely move his freezing, tired legs.",
                                         "But he began, suddenly, to feel happy. He began to recall happy times. He remembered his parents and his sister. He remembered his friends, Asher and Fiona. He remembered The Giver.",
                                         "Memories of joy flooded through him suddenly.",
                                         "He reached the place where the hill crested and he could feel the ground under his snow-covered feet become level. It would not be uphill anymore.",
                                         "\"We're almost there, Gabriel,\" he whispered, feeling quite certain without knowing why. \"I remember this place, Gabe.\" And it was true. But it was not a grasping of a thin and burdensome recollection; this was different. This was something that he could keep. It was a memory of his own.",
                                         "He hugged Gabriel and rubbed him briskly, warming him, to keep him alive. The wind was bitterly cold. The snow swirled, blurring his vision. But somewhere ahead, through the blinding storm, he knew there was warmth and light.",
                                         "Using his final strength, and a special knowledge that was deep inside him, Jonas found the sled that was waiting for them at the top of the hill. Numbly his hands fumbled for the rope.",
                                         "He settled himself on the sled and hugged Gabe close. The hill was steep but the snow was powdery and soft, and he knew that this time there would be no ice, no fall, no pain. Inside his freezing body, his heart surged with hope.",
                                         "They started down.",
                                         "Jonas felt himself losing consciousness and with his whole being willed himself to stay upright atop the sled, clutching Gabriel, keeping him safe. The runners sliced through the snow and the wind whipped at his face as they sped in a straight line through an incision that seemed to lead to the final destination, the place that he had always felt was waiting, the Elsewhere that held their future and their past.",
                                         "He forced his eyes open as they went downward, downward, sliding, and all at once he could see lights, and he recognized them now. He knew they were shining through the windows of rooms, that they were the red, blue, and yellow lights that twinkled from trees in places where families created and kept memories, where they celebrated love.",
                                         "Downward, downward, faster and faster. Suddenly he was aware with certainty and joy that below, ahead, they were waiting for him; and that they were waiting, too, for the baby. For the first time, he heard something that he knew to be music. He heard people singing.",
                                         "Behind him, across vast distances of space and time, from the place he had left, he thought he heard music too. But perhaps it was only an echo."))

```

```{r ch23plot}

ch23 <- annotate_ch23$token

ch23 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Quantity",
       y = "Parts of Speech")

lapply(annotate_ch23, head)

```
The above bar chart displays a count of each part of speech within Chapter 23 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

##Comparison Across Chapters

```{r verbcomparison}

verbch1 <- ch1 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch1")

verbch11 <- ch11 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch11")

verbch23 <- ch23 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch23")

verbcombined <- rbind(verbch1, verbch11, verbch23) %>%
  rename(chapters = new_col) %>%
  select(upos, chapters)
  
verbcombined %>% 
  count(chapters) %>% 
  ggplot(aes(n, chapters)) +
  geom_col() +
  scale_y_discrete(breaks = paste0("ch", c("1", "11", "23")),
                   labels = c("1", "11", "23")) +
  labs(x = "Number of verbs",
       y = "Chapter")

```
The above graph displays the amount of verbs present in chapters 1, 11, and 23. There is a downward trend in the amount of verbs as you move through the book. 

*I really love this figure! It raises so many questions! I think the use of bar plots was particularly effective here. I'm curious about the possibilities of adding an additional variable to this figure - something like "complexity" or a different metric that shows whether the verbs change in quality as well as quantity over the course of the book. Also, I wonder whether this is impacted by a change in length (i.e., is chapter 1 super long?). That may be a fun aspect to play with in a future draft (or just with this data set)*

*I think you picked great forms for the data you picked, and did a really good job navigating complex data - NLP is tough stuff! I haven't poked around NLP done in R before, so definitely learned a lot examining your code!*
