---
title: "Final Project"
author: "Makayla Whitney"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
#py_install("cleannlp")
#system("pip install cleannlp")
#cnlp_init_spacy(model_name = "en")

#These weren't working for me?? not sure what they are supposed to be doing but i was able to run the rest of the code installing cleannlp from r and not python
```

```{r code_setup}

#install.packages("cleanNLP")
library(cleanNLP)
library(dplyr)
library(tidyr)
#install.packages("gtsummary")
library(gtsummary)
library(magrittr)
library(ggplot2)
library(forcats)
cnlp_init_udpipe()


```

``` {r test}

#this code chunk was to test and learn the package

annotate_test <- cnlp_annotate(input = c(
  "Here is the first text. It is short.",
  "Here's the second. It is short too!",
  "The third text is the shortest."))


d <- annotate_test$token

d %>% 
  count(xpos) %>% 
  filter(xpos != ".") %>% 
  ggplot(aes(n, xpos)) +
  geom_col()


lapply(annotate_test, head)


cat("This is a normal string in R",
  "This is some text \"That has a quote\" and then some other text",
  sep = "\n")

```

#Introduction

14% of adults in the United States are functionally illiterate. "Functionally illiterate" is defined as reading below the 4th grade reading level. Many adults enroll in literacy tutoring programs that assist them with increasing their reading level to either, 1) complete their GED, 2) enroll in college or enter the workforce, 3) assist their children at home, or 4) other personal learning goals. These agencies are not equip with adult reading materials purely because they don't exist. Instead, agencies provide their tutors with children's books that are complementary to their adult learner's reading level. Children books are not topically engaging or relevant to most adult learners. I examine the parts of speech found within *The Giver* - a book found in most 8th grade curriculum across the country - to help conceptualize what concepts should be taught at the 8th grade level. This project is the beginning of a large-scale examination of 4th through 8th grade books to assist with creating a formula to be used in establishing adult reading materials. 

##Chapter 1

```{r ch1}
annotate_ch1 <- cnlp_annotate(input = c("It was almost December, and Jonas was beginning to be frightened.",
                                        "No. Wrong word, Jonas thought.",
                                        "Frightened meant that deep, sickening feeling of something terrible about to happen.",
                                        "Frightened was the way he had felt a year ago when an unidentified aircraft had overflown the community twice.",
                                        "He had seen it both times.",
                                        "Squinting toward the sky, he had seen the sleek jet, almost a blur at its high speed, go past, and a second later heard the blast of sound that followed.",
                                        "Then one more time, a moment later, from the opposite direction, the same plane.",
                                        "At first, he had been only fascinated.",
                                        "He had never seen aircraft so close, for it was against the rules for Pilots to fly over the community.",
                                        "Occasionally, when supplies were delivered by cargo planes to the landing field across the river, the children rode their bicycles to the riverbank and watched, intrigued, the unloading and then the takeoff directed to the west, always away from the community.",
                                        "But the aircraft a year ago had been different.",
                                        "It was not a squat, fat-bellied cargo plane but a needle-nosed single-pilot jet.",
                                        "Jonas, looking around anxiously, had seen others--adults as well as children--stop what they were doing and wait, confused, for an explanation of the frightening event.",
                                        "Then all of the citizens had been ordered to go into the nearest building and stay there. IMMEDIATELY, the rasping voice through the speakers had said. LEAVE YOUR BICYCLES WHERE THEY ARE.",
                                        "Instantly, obediently, Jonas had dropped his bike on its side on the path behind his family's dwelling. He had run indoors and stayed there, alone. His parents were both at work, and his little sister, Lily, was at the Childcare Center where she spent her after-school hours.",
                                        "Looking through the front window, he had seen no people: none of the busy afternoon crew of Street Cleaners, Landscape Workers, and Food Delivery people who usually populated the community at that time of day. He saw only the abandoned bikes here and there on their sides; an upturned wheel on one was still revolving slowly.",
                                        "He had been frightened then. The sense of his own community silent, waiting, had made his stomach churn. He had trembled.",
                                        "But it had been nothing. Within minutes the speakers had crackled again, and the voice, reassuring now and less urgent, had explained that a Pilot-in-Training had misread his navigational instructions and made a wrong turn. Desperately the Pilot had been trying to make his way back before his error was noticed.",
                                        "NEEDLESS TO SAY, HE WILL BE RELEASED, the voice had said, followed by silence. There was an ironic tone to that final message, as if the Speaker found it amusing; and Jonas had smiled a little, though he knew what a grim statement it had been. For a contributing citizen to be released from the community was a final decision, a terrible punishment, an overwhelming statement of failure.",
                                        "Even the children were scolded if they used the term lightly at play, jeering at a teammate who missed a catch or stumbled in a race.",
                                        "Jonas had done it once, had shouted at his best friend, \"That's it, Asher! You're released!\" when Asher's clumsy error had lost a match for his team. He had been taken aside for a brief and serious talk by the coach, had hung his head with guilt and embarrassment, and apologized to Asher after the game.",
                                        "Now, thinking about the feeling of fear as he pedaled home along the river path, he remembered that moment of palpable, stomach- sinking terror when the aircraft had streaked above. It was not what he was feeling now with December approaching. He searched for the right word to describe his own feeling.",
                                        "Jonas was careful about language. Not like his friend, Asher, who talked too fast and mixed things up, scrambling words and phrases until they were barely recognizable and often very funny.",
                                        "Jonas grinned, remembering the morning that Asher had dashed into the classroom, late as usual, arriving breathlessly in the middle of the chanting of the morning anthem. When the class took their seats at the conclusion of the patriotic hymn, Asher remained standing to make his public apology as was required.",
                                        "\"I apologize for inconveniencing my learning community.\" Asher ran through the standard apology phrase rapidly, still catching his breath. The Instructor and class waited patiently for his explanation. The students had all been grinning, because they had listened to Asher's explanations so many times before.",
                                        "\"I left home at the correct time but when I was riding along near the hatchery, the crew was separating some salmon. I guess I just got distraught, watching them.\"",
                                        "\"I apologize to my classmates,\" Asher concluded. He smoothed his rumpled tunic and sat down.",
                                        "\"We accept your apology, Asher.\" The class recited the standard response in unison. Many of the students were biting their lips to keep from laughing.",
                                        "\"I accept your apology, Asher,\" the Instructor said. He was smiling. \"And I thank you, because once again you have provided an opportunity for a lesson in language. 'Distraught' is too strong an adjective to describe salmon-viewing.\" He turned and wrote \"distraught\" on the instructional board. Beside it he wrote \"distracted.\"",
                                        "Jonas, nearing his home now, smiled at the recollection. Thinking, still, as he wheeled his bike into its narrow port beside the door, he realized that frightened was the wrong word to describe his feelings, now that December was almost here. It was too strong an adjective.",
                                        "He had waited a long time for this special December. Now that it was almost upon him, he wasn't frightened, but he was ... eager, he decided. He was eager for it to come. And he was excited, certainly. All of the Elevens were excited about the event that would be coming so soon.",
                                        "But there was a little shudder of nervousness when he thought about it, about what might happen.",
                                        "Apprehensive, Jonas decided. That's what I am.",
                                        "\"Who wants to be the first tonight, for feelings?\" Jonas's father asked, at the conclusion of their evening meal.",
                                        "It was one of the rituals, the evening telling of feelings.",
                                        "Jonas and his sister, Lily, argued over turns, over who would get to go first. Their parents, of course, were part of the ritual; they, too, told their feelings each evening. But like all parents - all adults - they didn't fight and wheedle for their turn.",
                                        "Nor did Jonas, tonight. His feelings were too complicated this evening. He wanted to share them, but he wasn't eager to begin the process of sifting through his own complicated emotions, even with the help that he knew his parents could give.",
                                        "\"You go, Lily,\" he said, seeing his sister, who was much younger-- only a Seven--wiggling with impatience in her chair.",
                                        "\"I felt very angry this afternoon,\" Lily announced. \"My Childcare group was at the play area, and we had a visiting group of Sevens, and they didn't obey the rules at all. One of them--a male; I don't know his name--kept going right to the front of the line for the slide, even though the rest of us were all waiting. I felt so angry at him. I made my hand into a fist, like this.\" She held up a clenched fist and the rest of the family smiled at her small defiant gesture.",
                                        "\"Why do you think the visitors didn't obey the rules?\" Mother asked.",
                                        "Lily considered, and shook her head. \"I don't know. They acted like...like...\"",
                                        "\"Animals?\" Jonas suggested. He laughed.",
                                        "\"That's right,\" Lily said, laughing too. \"Like animals.\" Neither child knew what the word meant, exactly, but it was often used to describe someone uneducated or clumsy, someone who didn't fit in.",
                                        "\"Where were the visitors from?\" Father asked.",
                                        "Lily frowned, trying to remember. \"Our leader told us, when he made the welcome speech, but I can't remember. I guess I wasn't paying attention. It was from another community. They had to leave very early, and they had their midday meal on the bus.\"",
                                        "Mother nodded. \"Do you think it's possible that their rules may be different? And so they simply didn't know what your play area rules were?\"",
                                        "Lily shrugged, and nodded. \"I suppose.\"",
                                        "\"You've visited other communities, haven't you?\" Jonas asked. \"My group has, often.\"",
                                        "Lily nodded again. \"When we were Sixes, we went and shared a whole school day with a group of Sixes in their community.\"",
                                        "\"How did you feel when you were there?\"",
                                        "Lily frowned. \"I felt strange. Because their methods were different. They were learning usages that my group hadn't learned yet, so we felt stupid.\"",
                                        "Father was listening with interest. \"I'm thinking, Lily,\" he said, \"about the boy who didn't obey the rules today. Do you think it's possible that he felt strange and stupid, being in a new place with rules that he didn't know about?\"",
                                        "Lily pondered that. \"Yes,\" she said, finally.",
                                        "\"I feel a little sorry for him,\" Jonas said, \"even though I don't even know him. I feel sorry for anyone who is in a place where he feels strange and stupid.\"",
                                        "\"How do you feel now, Lily?\" Father asked. \"Still angry?\"",
                                        "\"I guess not,\" Lily decided. \"I guess I feel a little sorry for him. And sorry I made a fist.\" She grinned.",
                                        "Jonas smiled back at his sister. Lily's feelings were always straightforward, fairly simple, usually easy to resolve. He guessed that his own had been, too, when he was a Seven.",
                                        "He listened politely, though not very attentively, while his father took his turn, describing a feeling of worry that he'd had that day at work: a concern about one of the newchildren who wasn't doing well. Jonas's father's title was Nurturer. He and the other Nurturers were responsible for all the physical and emotional needs of every newchild during its earliest life. It was a very important job, Jonas knew, but it wasn't one that interested him much.",
                                        "\"What gender is it?\" Lily asked.",
                                        "\"Male,\" Father said. \"He's a sweet little male with a lovely disposition. But he isn't growing as fast as he should, and he doesn't sleep soundly. We have him in the extra care section for supplementary nurturing, but the committee's beginning to talk about releasing him.\"",
                                        "\"Oh, no,\" Mother murmured sympathetically. \"I know how sad that must make you feel.\"",
                                        "Jonas and Lily both nodded sympathetically as well. Release of newchildren was always sad, because they hadn't had a chance to enjoy life within the community yet. And they hadn't done anything wrong.",
                                        "There were only two occasions of release which were not punishment. Release of the elderly, which was a time of celebration for a life well and fully lived; and release of a newchild, which always brought a sense of what-could-we- have-done. This was especially troubling for the Nurturers, like Father, who felt they had failed somehow. But it happened very rarely.",
                                        "\"Well,\" Father said, \"I'm going to keep trying. I may ask the committee for permission to bring him here at night, if you don't mind. You know what the night-crew Nurturers are like. I think this little guy needs something extra.\"",
                                        "\"Of course,\" Mother said, and Jonas and Lily nodded. They had heard Father complain about the night crew before. It was a lesser job, night-crew nurturing, assigned to those who lacked the interest or skills or insight for the more vital jobs of the daytime hours. Most of the people on the night crew had not even been given spouses because they lacked, somehow, the essential capacity to connect to others, which was required for the creation of a family unit.",
                                        "\"Maybe we could even keep him,\" Lily suggested sweetly, trying to look innocent. The look was fake, Jonas knew; they all knew.",
                                        "\"Lily,\" Mother reminded her, smiling, \"you know the rules.\"",
                                        "Two children--one male, one female--to each family unit. It was written very clearly in the rules.",
                                        "Lily giggled. \"Well,\" she said, \"I thought maybe just this once.\"",
                                        "Next, Mother, who held a prominent position at the Department of Justice, talked about her feelings. Today a repeat offender had been brought before her, someone who had broken the rules before. Someone who she hoped had been adequately and fairly punished, and who had been restored to his place: to his job, his home, his family unit. To see him brought before her a second time caused her overwhelming feelings of frustration and anger. And even guilt, that she hadn't made a difference in his life.",
                                        "\"I feel frightened, too, for him,\" she confessed. \"You know that there's no third chance. The rules say that if there's a third transgression, he simply has to be released.\" Jonas shivered. He knew it happened. There was even a boy in his group of Elevens whose father had been released years before. No one ever mentioned it; the disgrace was unspeakable. It was hard to imagine.",
                                        "Lily stood up and went to her mother. She stroked her mother's arm.",
                                        "From his place at the table, Father reached over and took her hand. Jonas reached for the other.",
                                        "One by one, they comforted her. Soon she smiled, thanked them, and murmured that she felt soothed.",
                                        "The ritual continued. \"Jonas?\" Father asked. \"You're last, tonight.\"",
                                        "Jonas sighed. This evening he almost would have preferred to keep his feelings hidden. But it was, of course, against the rules.",
                                        "\"I'm feeling apprehensive,\" he confessed, glad that the appropriate descriptive word had finally come to him. \"Why is that, son?\" His father looked concerned.",
                                        "\"I know there's really nothing to worry about,\" Jonas explained, \"and that every adult has been through it. I know you have, Father, and you too, Mother. But it's the Ceremony that I'm apprehensive about. It's almost December.\"",
                                        "Lily looked up, her eyes wide. \"The Ceremony of Twelve,\" she whispered in an awed voice. Even the smallest children--Lily's age and younger--knew that it lay in the future for each of them.",
                                        "\"I'm glad you told us of your feelings,\" Father said.",
                                        "\"Lily,\" Mother said, beckoning to the little girl, \"Go on now and get into your nightclothes. Father and I are going to stay here and talk to Jonas for a while.\"",
                                        "Lily sighed, but obediently she got down from her chair. \"Privately?\" she asked.",
                                        "Mother nodded. \"Yes,\" she said, \"this talk will be a private one with Jonas.\"" ))
```

```{r ch1plot}

ch1 <- annotate_ch1$token

#version 1
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".") %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "",
       y = "",
       title = "Parts of Speech")

#version 2
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 1 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 1 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch1, head)

#summarize the percentages of parts of speech present in ch1 
ch1 %>% 
  select(upos) %>%
  tbl_summary()


```
The above bar chart displays a count of each part of speech within Chapter 1 of *The Giver*. 

##Chapter 2
```{r ch2}

annotate_ch2 <- cnlp_annotate(input = c("Jonas watched as his father poured a fresh cup of coffee. He waited.",
                                        "\"You know,\" his father finally said, \"every December was exciting to me when I was young. And it has been for you and Lily, too, I'm sure. Each December brings such changes.\"",
                                        "Jonas nodded. He could remember the Decembers back to when he had become, well, probably a Four. The earlier ones were lost to him. But he observed them each year, and he remembered Lily's earliest Decembers. He remembered when his family received Lily, the day she was named, the day that she had become a One.",
                                        "The Ceremony for the Ones was always noisy and fun. Each December, all the newchildren born in the previous year turned One. One at a time--there were always fifty in each year's group, if none had been released --they had been brought to the stage by the Nurturers who had cared for them since birth. Some were already walking, wobbly on their unsteady legs; others were no more than a few days old, wrapped in blankets, held by their Nurturers.",
                                        "\"I enjoy the Naming,\" Jonas said.",
                                        "His mother agreed, smiling. \"The year we got Lily, we knew, of course, that we'd receive our female, because we'd made our application and been approved. But I'd been wondering and wondering what her name would be.\"",
                                        "\"I could have sneaked a look at the list prior to the ceremony,\" Father confided. \"The committee always makes the list in advance, and it's right there in the office at the Nurturing Center.\"",
                                        "\"As a matter of fact,\" he went on, \"I feel a little guilty about this. But I did go in this afternoon and looked to see if this year's Naming list had been made yet. It was right there in the office, and I looked up number Thirty-six--that's the little guy I've been concerned about-- because it occurred to me that it might enhance his nurturing if I could call him by a name. Just privately, of course, when no one else is around.\"",
                                        "\"Did you find it?\" Jonas asked. He was fascinated. It didn't seem a terribly important rule, but the fact that his father had broken a rule at all awed him. He glanced at his mother, the one responsible for adherence to the rules, and was relieved that she was smiling.",
                                        "His father nodded. \"His name--if he makes it to the Naming without being released, of course--is to be Gabriel. So I whisper that to him when I feed him every four hours, and during exercise and playtime. If no one can hear me.\"",
                                        "\"I call him Gabe, actually,\" he said, and grinned. \"Gabe.\" Jonas tried it out. A good name, he decided. Though Jonas had only become a Five the year that they acquired Lily and learned her name, he remembered the excitement, the conversations at home, wondering about her: how she would look, who she would be, how she would fit into their established family unit. He remembered climbing the steps to the stage with his parents, his father by his side that year instead of with the Nurturers, since it was the year that he would be given a newchild of his own.",
                                        "He remembered his mother taking the newchild, his sister, into her arms, while the document was read to the assembled family units. \"Newchild Twenty-three,\" the Namer had read. \"Lily.\"",
                                        "He remembered his father's look of delight, and that his father had whispered, \"She's one of my favorites. I was hoping for her to be the one.\" The crowd had clapped, and Jonas had grinned. He liked his sister's name. Lily, barely awake, had waved her small fist. Then they had stepped down to make room for the next family unit.",
                                        "\"When I was an Eleven,\" his father said now, \"as you are, Jonas, I was very impatient, waiting for the Ceremony of Twelve. It's a long two days. I remember that I enjoyed the Ones, as I always do, but that I didn't pay much attention to the other ceremonies, except for my sister's. She became a Nine that year, and got her bicycle. I'd been teaching her to ride mine, even though technically I wasn't supposed to.\"",
                                        "Jonas laughed. It was one of the few rules that was not taken very seriously and was almost always broken. The children all received their bicycles at Nine; they were not allowed to ride bicycles before then. But almost always, the older brothers and sisters had secretly taught the younger ones. Jonas had been thinking already about teaching Lily.",
                                        "There was talk about changing the rule and giving the bicycles at an earlier age. A committee was studying the idea. When something went to a committee for study, the people always joked about it. They said that the committee members would become Elders by the time the rule change was made.",
                                        "Rules were very hard to change. Sometimes, if it was a very important rule--unlike the one governing the age for bicycles--it would have to go, eventually, to The Receiver for a decision. The Receiver was the most important Elder. Jonas had never even seen him, that he knew of; someone in a position of such importance lived and worked alone. But the committee would never bother The Receiver with a question about bicycles; they would simply fret and argue about it themselves for years, until the citizens forgot that it had ever gone to them for study.",
                                        "His father continued. \"So I watched and cheered when my sister, Katya, became a Nine and removed her hair ribbons and got her bicycle,\" Father went on. \"Then I didn't pay much attention to the Tens and Elevens. And finally, at the end of the second day, which seemed to go on forever, it was my turn. It was the Ceremony of Twelve.\"",
                                        "Jonas shivered. He pictured his father, who must have been a shy and quiet boy, for he was a shy and quiet man, seated with his group, waiting to be called to the stage. The Ceremony of Twelve was the last of the Ceremonies. The most important.",
                                        "I remember how proud my parents looked--and my sister, too; even though she wanted to be out riding the bicycle publicly, she stopped fidgeting and was very still and attentive when my turn came.",
                                        "\But to be honest, Jonas,\" his father said, \"for me there was not the element of suspense that there is with your Ceremony. Because I was already fairly certain of what my Assignment was to be.\"",
                                        "Jonas was surprised. There was no way, really, to know in advance. It was a secret selection, made by the leaders of the community, the Committee of Elders, who took the responsibility so seriously that there were never even any jokes made about Assignments.",
                                        "His mother seemed surprised, too. \"How could you have known?\" she asked.",
                                        "His father smiled his gentle smile. \"Well, it was clear to me--and my parents later confessed that it had been obvious to them, too--what my aptitude was. I had always loved the newchildren more than anything. When my friends in my age group were holding bicycle races, or building toy vehicles or bridges with their construction sets, or--\"",
                                        "\"All the things I do with my friends,\" Jonas pointed out, and his mother nodded in agreement.",
                                        "\"I always participated, of course, because as children we must experience all of those things. And I studied hard in school, as you do, Jonas. But again and again, during free time, I found myself drawn to the newchildren. I spent almost all of my volunteer hours helping in the Nurturing Center. Of course the Elders knew that, from their observation.\"",
                                        "Jonas nodded. During the past year he had been aware of the increasing level of observation. In school, at recreation time, and during volunteer hours, he had noticed the Elders watching him and the other Elevens. He had seen them taking notes. He knew, too, that the Elders were meeting for long hours with all of the instructors that he and the other Elevens had had during their years of school.",
                                        "\"So I expected it, and I was pleased, but not at all surprised, when my Assignment was announced as Nurturer,\" Father explained.",
                                        "\"Did everyone applaud, even though they weren't surprised?\" Jonas asked.",
                                        "\"Oh, of course. They were happy for me, that my Assignment was what I wanted most. I felt very fortunate.\" His father smiled.",
                                        "\"Were any of the Elevens disappointed, your year?\" Jonas asked. Unlike his father, he had no idea what his Assignment would be. But he knew that some would disappoint him. Though he respected his father's work, Nurturer would not be his wish. And he didn't envy Laborers at all.",
                                        "His father thought. \"No, I don't think so. Of course the Elders are so careful in their observations and selections.\"",
                                        "\"I think it's probably the most important job in our community,\" his mother commented.",
                                        "\"My friend Yoshiko was surprised by her selection as Doctor,\" Father said, \"but she was thrilled. And let's see, there was Andrei--I remember that when we were boys he never wanted to do physical things. He spent all the recreation time he could with his construction set, and his volunteer hours were always on building sites. The Elders knew that, of course. Andrei was given the Assignment of Engineer and he was delighted.\"",
                                        "\"Andrei later designed the bridge that crosses the river to the west of town,\" Jonas's mother said. \"It wasn't there when we were children.\"",
                                        "\"There are very rarely disappointments, Jonas. I don't think you need to worry about that,\" his father reassured him. \"And if there are, you know there's an appeal process.\" But they all laughed at that--an appeal went to a committee for study.",
                                        "\"I worry a little about Asher's Assignment,\" Jonas confessed. \"Asher's such fun. But he doesn't really have any serious interests. He makes a game out of everything.\"",
                                        "His father chuckled. \"You know,\" he said, \"I remember when Asher was a newchild at the Nurturing Center, before he was named. He never cried. He giggled and laughed at everything. All of us on the staff enjoyed nurturing Asher.\"",
                                        "\"The Elders know Asher,\" his mother said. \"They'll find exactly the right Assignment for him. I don't think you need to worry about him. But, Jonas, let me warn you about something that may not have occurred to you. I know I didn't think about it until after my Ceremony of Twelve.\"",
                                        "\"What's that?\"",
                                        "\"Well, it's the last of the Ceremonies, as you know. After Twelve, age isn't important. Most of us even lose track of how old we are as time passes, though the information is in the Hall of Open Records, and we could go and look it up if we wanted to. What's important is the preparation for adult life, and the training you'll receive in your Assignment.\"",
                                        "\"I know that,\" Jonas said. \"Everyone knows that.\"",
                                        "\"But it means,\" his mother went on, \"that you'll move into a new group. And each of your friends will. You'll no longer be spending your time with your group of Elevens. After the Ceremony of Twelve, you'll be with your Assignment group, with those in training. No more volunteer hours. No more recreation hours. So your friends will no longer be as close.\"",
                                        "Jonas shook his head. \"Asher and I will always be friends,\" he said firmly. \"And there will still be school.\"",
                                        "\"That's true,\" his father agreed. \"But what your mother said is true as well. There will be changes.\"",
                                        "\"Good changes, though,\" his mother pointed out. \"After my Ceremony of Twelve, I missed my childhood recreation. But when I entered my training for Law and Justice, I found myself with people who shared my interests. I made friends on a new level, friends of all ages.\"",
                                        "\"Did you still play at all, after Twelve?\" Jonas asked.",
                                        "\"Occasionally,\" his mother replied. \"But it didn't seem as important to me.\"",
                                        "\"I did,\" his father said, laughing. \"I still do. Every day, at the Nurturing Center, I play bounce-on-the-knee, and peek-a-boo, and hug-the-teddy.\" He reached over and stroked Jonas's neatly trimmed hair. \"Fun doesn't end when you become Twelve.\"",
                                        "Lily appeared, wearing her nightclothes, in the doorway. She gave an impatient sigh. \"This is certainly a very long private conversation,\" she said. \"And there are certain people waiting for their comfort object.\"",
                                        "\"Lily,\" her mother said fondly, \"you're very dose to being an Eight, and when you're an Eight, your comfort object will be taken away. It will be recycled to the younger children. You should be starting to go off to sleep without it.\"",
                                        "But her father had already gone to the shelf and taken down the stuffed elephant which was kept there. Many of the comfort objects, like Lily's, were soft, stuffed, imaginary creatures. Jonas's had been called a bear.",
                                        "\"Here you are, Lily-billy,\" he said. \"I'll come help you remove your hair ribbons.\"",
                                        "Jonas and his mother rolled their eyes, yet they watched affectionately as Lily and her father headed to her sleeping-room with the stuffed elephant that had been given to her as her comfort object when she was born. His mother moved to her big desk and opened her briefcase; her work never seemed to end, even when she was at home in the evening. Jonas went to his own desk and began to sort through his school papers for the evening's assignment. But his mind was still on December and the coming Ceremony.",
                                        "Though he had been reassured by the talk with his parents, he hadn't the slightest idea what Assignment the Elders would be selecting for his future, or how he might feel about it when the day came."))

```

##Chapter 3
```{r ch3}
annotate_ch3 <- cnlp_annotate(input = c("\"Oh, look!\" Lily squealed in delight. \"Isn't he cute? Look how tiny he is! And he has funny eyes like yours, Jonas!\" Jonas glared at her. He didn't like it that she had mentioned his eyes. He waited for his father to chastise Lily. But Father was busy unstrapping the carrying basket from the back of his bicycle. Jonas walked over to look.",
                                        "It was the first thing Jonas noticed as he looked at the newchild peering up curiously from the basket. The pale eyes.",
                                        "Almost every citizen in the community had dark eyes. His parents did, and Lily did, and so did all of his group members and friends. But there were a few exceptions: Jonas himself, and a female Five who he had noticed had the different, lighter eyes. No one mentioned such things; it was not a rule, but was considered rude to call attention to things that were unsettling or different about individuals. Lily, he decided, would have to learn that soon, or she would be called in for chastisement because of her insensitive chatter.",
                                        "Father put his bike into its port. Then he picked up the basket and carried it into the house. Lily followed behind, but she glanced back over her shoulder at Jonas and teased, \"Maybe he had the same Birthmother as you.\"",
                                        "Jonas shrugged. He followed them inside. But he had been startled by the newchild's eyes. Mirrors were rare in the community; they weren't forbidden, but there was no real need of them, and Jonas had simply never bothered to look at himself very often even when he found himself in a location where a mirror existed. Now, seeing the newchild and its expression, he was reminded that the light eyes were not only a rarity but gave the one who had them a certain look--what was it? Depth, he decided; as if one were looking into the clear water of the river, down to the bottom, where things might lurk which hadn't been discovered yet. He felt self-conscious, realizing that he, too, had that look.",
                                        "He went to his desk, pretending not to be interested in the newchild. On the other side of the room, Mother and Lily were bending over to watch as Father unwrapped its blanket.",
                                        "\"What's his comfort object called?\" Lily asked, picking up the stuffed creature which had been placed beside the newchild in his basket.",
                                        "Father glanced at it. \"Hippo,\" he said.",
                                        "Lily giggled at the strange word. \"Hippo,\" she repeated, and put the comfort object down again. She peered at the unwrapped newchild, who waved his arms.",
                                        "\"I think newchildren are so cute,\" Lily sighed. \"I hope I get assigned to be a Birthmother.\"",
                                        "\"Lily!\" Mother spoke very sharply. \"Don't say that. There's very little honor in that Assignment.\"",
                                        "\"But I was talking to Natasha. You know the Ten who lives around the corner? She does some of her volunteer hours at the Birthing Center. And she told me that the Birthmothers get wonderful food, and they have very gentle exercise periods, and most of the time they just play games and amuse themselves while they're waiting. 1 think I'd like that,\" Lily said petulantly.",
                                        "\"Three years,\" Mother told her firmly. \"Three births and that's all. After that they are Laborers for the rest of their adult lives, until the day that they enter the House of the Old. Is that what you want, Lily? Three lazy years, and then hard physical labor until you are old?\"",
                                        "\"Well, no, I guess not,\" Lily acknowledged reluctantly Father turned the newchild onto his tummy in the basket. He sat beside it and rubbed its small back with a rhythmic motion. \"Anyway, Lily-billy,\" he said affectionately, \"the Birthmothers never even get to see newchildren. If you enjoy the little ones so much, you should hope for an Assignment as Nurturer.\"",
                                        "\"When you're an Eight and start your volunteer hour, you can try some at the Nurturing Center,\" Mother suggested.",
                                        "\"Yes, I think I will,\" Lily said. She knelt beside the basket. \"What did you say his name is? Gabriel? Hello, Gabriel,\" she said in a singsong voice. Then she giggled \"Ooops,\" she whispered. \"I think he's alseep. I guess I'd better be quiet.\"",
                                        "Jonas turned to the school assignments on his desk Some chance of that, he thought. Lily was never quiet Probably she should hope for an Assignment as Speaker so that she could sit in the office with the microphone a1l day, making announcements. He laughed silently to himself, picturing his sister droning on in the self- important voice that all the Speakers seemed to develop, saying things like, ATTENTION. THIS IS A REMINDER TO FEMALES UNDER NINE THAT HAIR RIBBONS ARE TO BE NEATLY TIED AT ALL TIMES.",
                                        "He turned toward Lily and noticed to his satisfaction that her ribbons were, as usual, undone and dangling. There would be an announcement like that quite soon, he felt certain, and it would be directed mainly at Lily, though her name, of course, would not be mentioned. Everyone would know.",
                                        "Everyone had known, he remembered with humiliation, that the announcement ATTENTION. THIS IS A REMINDER TO MALE ELEVENS THAT OBJECTS ARE NOT TO BE REMOVED FROM THE RECREATION AREA AND THAT SNACKS ARE TO BE EATEN, NOT HOARDED had been specifically directed at him, the day last month that he had taken an apple home. No one had mentioned it, not even his parents, because the public announcement had been sufficient to produce the appropriate remorse. He had, of course, disposed of the apple and made his apology to the Recreation Director the next morning, before school.",
                                        "Jonas thought again about that incident. He was still bewildered by it. Not by the announcement or the necessary apology; those were standard procedures, and he had deserved them--but by the incident itself. He probably should have brought up his feeling of bewilderment that very evening when the family unit had shared their feelings of the day. But he had not been able to sort out and put words to the source of his confusion, so he had let it pass.",
                                        "It had happened during the recreation period, when he had been playing with Asher. Jonas had casually picked up an apple from the basket where the snacks were kept, and had thrown it to his friend. Asher had thrown it back, and they had begun a simple game of catch.",
                                        "There had been nothing special about it; it was an activity that he had performed countless times: throw, catch; throw, catch. It was effortless for Jonas, and even boring, though Asher enjoyed it, and playing catch was a required activity for Asher because it would improve his hand-eye coordination, which was not up to standards.",
                                        "But suddenly Jonas had noticed, following the path of the apple through the air with his eyes, that the piece of fruit had--well, this was the part that he couldn't adequately understand--the apple had changed. Just for an instant. It had changed in mid-air, he remembered. Then it was in his hand, and he looked at it carefully, but it was the same apple. Unchanged. The same size and shape: a perfect sphere. The same nondescript shade, about the same shade as his own tunic.",
                                        "There was absolutely nothing remarkable about that apple. He had tossed it back and forth between his hands a few times, then thrown it again to Asher. And again--in the air, for an instant only--it had changed.",
                                        "It had happened four times. Jonas had blinked, looked around, and then tested his eyesight, squinting at the small print on the identification badge attached to his tunic. He read his name quite clearly. He could also clearly see Asher at the other end of the throwing area. And he had had no problem catching the apple.",
                                        "Jonas had been completely mystified.",
                                        "\"Ash?\" he had called. \"Does anything seem strange to you? About the apple?\"",
                                        "\"Yes,\" Asher called back, laughing. \"It jumps out of my hand onto the ground!\" Asher had just dropped it once again.",
                                        "So Jonas laughed too, and with his laughter tried to ignore his uneasy conviction that something had happened. But he had taken the apple home, against the recreation area rules. That evening, before his parents and Lily arrived at the dwelling, he had held it in his hands and looked at it carefully. It was slightly bruised now, because Asher had dropped it several times. But there was nothing at all unusual about the apple.",
                                        "He had held a magnifying glass to it. He had tossed it several times across the room, watching, and then rolled it around and around on his desktop, waiting for the thing to happen again.",
                                        "But it hadn't. The only thing that happened was the announcement later that evening over the speaker, the announcement that had singled him out without using his name, that had caused both of his parents to glance meaningfully at his desk where the apple still lay.",
                                        "Now, sitting at his desk, staring at his schoolwork as his family hovered over the newchild in its basket, he shook his head, trying to forget the odd incident. He forced himself to arrange his papers and try to study a little before the evening meal. The newchild, Gabriel, stirred and whimpered, and Father spoke softly to Lily, explaining the feeding procedure as he opened the container that held the formula and equipment.",
                                        "The evening proceeded as all evenings did in the family unit, in the dwelling, in the community: quiet, reflective, a time for renewal and preparation for the day to come. It was different only in the addition to it of the newchild with his pale, solemn, knowing eyes."))
```

##Chapter 4
```{r ch4}
annotate_ch4 <- cnlp_annotate(input = c("Jonas rode at a leisurely pace, glancing at the bikeports beside the buildings to see if he could spot Asher's. He didn't often do his volunteer hours with his friend because Asher frequently fooled around and made serious work a little difficult. But now, with Twelve coming so soon and the volunteer hours ending, it didn't seem to matter.",
                                        "The freedom to choose where to spend those hours had always seemed a wonderful luxury to Jonas; other hours of the day were so carefully regulated.",
                                        "He remembered when he had become an Eight, as Lily would do shortly, and had been faced with that freedom of choice. The Eights always set out on their first volunteer hour a little nervously, giggling and staying in groups of friends. They almost invariably did their hours on Recreation Duty first, helping with the younger ones in a place where they still felt comfortable. But with guidance, as they developed self-confidence and maturity, they moved on to other jobs, gravitating toward those that would suit their own interests and skills.",
                                        "A male Eleven named Benjamin had done his entire nearly-Four years in the Rehabilitation Center, working with citizens who had been injured. It was rumored that he was as skilled now as the Rehabilitation Directors themselves, and that he had even developed some machines and methods to hasten rehabilitation. There was no doubt that Benjamin would receive his Assignment to that field and would probably be permitted to bypass most of the training.",
                                        "Jonas was impressed by the things Benjamin had achieved. He knew him, of course, since they had always been groupmates, but they had never talked about the boy's accomplishments because such a conversation would have been awkward for Benjamin.
There was never any comfortable way to mention or discuss one's successes without breaking the rule against bragging, even if one didn't mean to. It was a minor rule, rather like rudeness, punishable only by gentle chastisement. But still. Better to steer clear of an occasion governed by a rule which would be so easy to break.",
                                        "The area of dwellings behind him, Jonas rode past the community structures, hoping to spot Asher's bicycle parked beside one of the small factories or office buildings. He passed the Childcare Center where Lily stayed after school, and the play areas surrounding it.
He rode through the Central Plaza and the large Auditorium where public meetings were held.",
                                        "Jonas slowed and looked at the nametags on the bicycles lined up outside the Nurturing Center. Then he checked those outside Food Distribution; it was always fun to help with the deliveries, and he hoped he would find his friend there so that they could go together on the daily rounds, carrying the cartons of supplies into the dwellings of the community. But he finally found Asher's bicycle-- leaning, as usual, instead of upright in its port, as it should have been--at the House of the Old.",
                                        "There was only one other child's bicycle there, that of a female Eleven named Fiona. Jonas liked Fiona. She was a good student, quiet and polite, but she had a sense of fun as well, and it didn't surprise him that she was working with Asher today. He parked his bicycle neatly in the port beside theirs and entered the building.",
                                        "\"Hello, Jonas,\" the attendant at the front desk said. She handed him the sign-up sheet and stamped her own official seal beside his signature. All of his volunteer hours would be carefully tabulated at the Hall of Open Records. Once, long ago, it was whispered among the children, an Eleven had arrived at the Ceremony of Twelve only to hear a public announcement that he had not completed the required number of volunteer hours and would not, therefore, be given his Assignment. He had been permitted an additional month in which to complete the hours, and then given his Assignment privately, with no applause, no celebration: a disgrace that had clouded his entire future.",
                                        "\"It's good to have some volunteers here today,\" the attendant told him. \"We celebrated a release this morning, and that always throws the schedule off a little, so things get backed up.\" She looked at a printed sheet. \"Let's see. Asher and Fiona are helping in the bathing room. Why don't you join them there? You know where it is, don't you?\"",
                                        "Jonas nodded, thanked her, and walked down the long hallway. He glanced into the rooms on either side. The Old were sitting quietly, some visiting and talking with one another, others doing handwork and simple crafts. A few were asleep. Each room was comfortably furnished, the floors covered with thick carpeting. It was a serene and slow-paced place, unlike the busy centers of manufacture and distribution where the daily work of the community occurred.",
                                        "Jonas was glad that he had, over the years, chosen to do his hours in a variety of places so that he could experience the differences. He realized, though, that not focusing on one area meant he was left with not the slightest idea-not even a guess of what his Assignment would be.",
                                        "He laughed softly. Thinking about the Ceremony again, Jonas? he teased himself. But he suspected that with the date so near, probably all of his friends were, too.",
                                        "He passed a Caretaker walking slowly with one of the Old in the hall. \"Hello, Jonas,\" the young uniformed man said, smiling pleasantly. The woman beside him, whose arm he held, was hunched over as she shuffled along in her soft slippers. She looked toward Jonas and smiled, but her dark eyes were clouded and blank. He realized she was blind.",
                                        "He entered the bathing room with its warm moist air and scent of cleansing lotions. He removed his tunic, hung it carefully on a wall hook, and put on the volunteer's smock that was folded on a shelf.",
                                        "\"Hi, Jonas!\" Asher called from the corner where he was kneeling beside a tub. Jonas saw Fiona nearby, at a different tub. She looked up and smiled at him, but she was busy, gently washing a man who lay in the warm water.",
                                        "Jonas greeted them and the caretaking attendants at work nearby. Then he went to the row of padded lounging chairs where others of the Old were waiting. He had worked here before; he knew what to do.",
                                        "\"Your turn, Larissa,\" he said, reading the nametag on the woman's robe. \"I'll just start the water and then help you up.\" He pressed the button on a nearby empty tub and watched as the warm water flowed in through the many small openings on the sides. The tub would be filled in a minute and the water flow would stop automatically.",
                                        "He helped the woman from the chair, led her to the tub, removed her robe, and steadied her with his hand on her arm as she stepped in and lowered herself. She leaned back and sighed with pleasure, her head on a soft cushioned headrest.",
                                        "\"Comfortable?\" he asked, and she nodded, her eyes closed. Jonas squeezed cleansing lotion onto the clean sponge at the edge of the tub and began to wash her frail body.",
                                        "Last night he had watched as his father bathed the newchild. This was much the same: the fragile skin, the soothing water, the gentle motion of his hand, slippery with soap. The relaxed, peaceful smile on the woman's face reminded him of Gabriel being bathed.",
                                        "And the nakedness, too. It was against the rules for children or adults to look at another's nakedness; but the rule did not apply to newchildren or the Old. Jonas was glad. It was a nuisance to keep oneself covered while changing for games, and the required apology if one had by mistake glimpsed another's body was always awkward. He couldn't see why it was necessary. He liked the feeling of safety here in this warm and quiet room; he liked the expression of trust on the woman's face as she lay in the water unprotected, exposed, and free.",
                                        "From the corner of his eye he could see his friend Fiona help the old man from the tub and tenderly pat his thin, naked body dry with an absorbant cloth. She helped him into his robe.",
                                        "Jonas thought Larissa had drifted into sleep, as the Old often did, and he was careful to keep his motions steady and gentle so he wouldn't wake her. He was surprised when she spoke, her eyes still closed. \"This morning we celebrated the release of Roberto,\" she told him. \"It was wonderful.\"",
                                        "\"I knew Roberto!\" Jonas said. \"I helped with his feeding the last time I was here, just a few weeks ago. He was a very interesting man.\"",
                                        "Larissa opened her eyes happily. \"They told his whole life before they released him,\" she said. \"They always do. But to be honest,\" she whispered with a mischievous look, \"some of the tellings are a little boring. I've even seen some of the Old fall asleep during tellings- when they released Edna recently. Did you know Edna?\"",
                                        "Jonas shook his head. He couldn't recall anyone named Edna.",
                                        "\"Well, they tried to make her life sound meaningful. And of course,\" she added primly, \"all lives are meaningful, I don't mean that they aren't. But Edna. My goodness. She was a Birthmother, and then she worked in Food Production for years, until she came here. She never even had a family unit.\"",
                                        "Larissa lifted her head and looked around to make sure no one else was listening. Then she confided, \"I don't think Edna was very smart.\"",
                                        "Jonas laughed. He rinsed her left arm, laid it back into the water, and began to wash her feet. She murmured with pleasure as he massaged her feet with the sponge.",
                                        "\"But Roberto's life was wonderful,\" Larissa went on, after a moment. \"He had been an Instructor of Elevens --you know how important that is--and he'd been on the Planning Committee. And-- goodness, I don't know how he found the time--he also raised two very successful children,, and he was also the one who did the landscaping design for the Central Plaza. He didn't do the actual labor, of course.\"",
                                        "\"Now your back. Lean forward and I'll help you sit up.\" Jonas put his arm around her and supported her as she sat. He squeezed the sponge against her back and began to rub her sharp-boned shoulders. \"Tell me about the celebration.\"",
                                        "\"Well, there was the telling of his life. That is always first. Then the toast. We all raised our glasses and cheered. We chanted the anthem. He made a lovely goodbye speech. And several of us made little speeches wishing him well. I didn't, though. I've never been fond of public speaking.\"",
                                        "\"He was thrilled. You should have seen the look on his face when they let him go.\"",
                                        "Jonas slowed the strokes of his hand on her back thoughtfully. \"Larissa,\" he asked, \"what happens when they make the actual release? Where exactly did Roberto go?\"",
                                        "She lifted her bare wet shoulders in a small shrug. \"I don't know. I don't think anybody does, except the committee. He just bowed to all of us and then walked, like they all do, through the special door in the Releasing Room. But you should have seen his look. Pure happiness, I'd call it.\"",
                                        "Jonas grinned. \"I wish I'd been there to see it.\"",
                                        "Larissa frowned. \"I don't know why they don't let children come. Not enough room, I guess. They should enlarge the Releasing Room.\"",
                                        "\"We'll have to suggest that to the committee. Maybe they'd study it,\" Jonas said slyly, and Larissa chortled with laughter.",
                                        "\"Right!\" she hooted, and Jonas helped her from the tub."))
```

##Chapter 11
```{r ch11}

annotate_ch11 <- cnlp_annotate(input = c("Jonas felt nothing unusual at first. He felt only the light touch of the old man's hands on his back.",
                                         "He tried to relax, to breathe evenly. The room was absolutely silent, and for a moment Jonas feared that he might disgrace himself now, on the first day of his training, by falling asleep.",
                                         "Then he shivered. He realized that the touch of the hands felt, suddenly, cold. At the same instant, breathing in, he felt the air change, and his very breath was cold. He licked his lips, and in doing so, his tongue touched the suddenly chilled air.",
                                         "It was very startling; but he was not at all frightened, now. He was filled with energy, and he breathed again, feeling the sharp intake of frigid air. Now, too, he could feel cold air swirling around his entire body. He felt it blow against his hands where they lay at his sides, and over his back.",
                                         "The touch of the man's hands seemed to have disappeared.",
                                         "Now he became aware of an entirely new sensation: pinpricks? No, because they were soft and without pain. Tiny, cold, featherlike feelings peppered his body and face. He put out his tongue again, and caught one of the dots of cold upon it. It disappeared from his awareness instantly; but he caught another, and another. The sensation made him smile.",
                                         "One part of his consciousness knew that he was still lying there, on the bed, in the Annex room. Yet another, separate part of his being was upright now, in a sitting position, and beneath him he could feel that he was not on the soft decorated bedcovering at all, but rather seated on a flat, hard surface. His hands now held (though at the same time they were still motionless at his sides) a rough, damp rope.",
                                         "And he could see, though his eyes were closed. He could see a bright, whirling torrent of crystals in the air around him, and he could see them gather on the backs of his hands, like cold fur.",
                                         "His breath was visible.",
                                         "Beyond, through the swirl of what he now, somehow, perceived was the thing the old man had spoken of-- snow-- he could look out and down a great distance. He was up high someplace. The ground was thick with the furry snow, but he sat slightly above it on a hard, flat object.",
                                         "Sled, he knew abruptly. He was sitting on a thing called sled. And the sled itself seemed to be poised at the top of a long, extended mound that rose from the very land where he was. Even as he thought the word \"mound,\" his new consciousness told him hill.",
                                         "Then the sled, with Jonas himself upon it, began to move through the snowfall, and he understood instantly that now he was going downhill. No voice made an explanation. The experience explained itself to him.",
                                         "His face cut through the frigid air as he began the descent, moving through the substance called snow on the vehicle called sled, which propelled itself on what he now knew without doubt to be runners.",
                                         "Comprehending all of those things as he sped downward, he was free to enjoy the breathless glee that overwhelmed him: the speed, the clear cold air, the total silence, the feeling of balance and excitement and peace.",
                                         "Then, as the angle of incline lessened, as the mound--the hill-- flattened, nearing the bottom, the sled's forward motion slowed. The snow was piled now around it, and he pushed with his body, moving it forward, not wanting the exhilarating ride to end.",
                                         "Finally the obstruction of the piled snow was too much for the thin runners of the sled, and he came to a stop. He sat there for a moment, panting, holding the rope in his cold hands. Tentatively he opened his eyes--not his snow-hill-sled eyes, for they had been open throughout the strange ride. He opened his ordinary eyes, and saw that he was still on the bed, that he had not moved at all.",
                                         "The old man, still beside the bed, was watching him. \"How do you feel?\" he asked.",
                                         "Jonas sat up and tried to answer honestly. \"Surprised,\" he said, after a moment.",
                                         "The old man wiped his forehead with his sleeve. \"Whew,\" he said. \"It was exhausting. But you know, even transmitting that tiny memory to you--I think it lightened me just a little.\"",
                                         "\"Do you mean--you did say I could ask questions?\" The man nodded, encouraging his question.",
                                         "\"Do you mean that now you don't have the memory of it--of that ride on the sled--anymore?\"",
                                         "\"That's right. A little weight off this old body.\"",
                                         "\"But it was such fun! And now you don't have it anymore! I took it from you!\"",
                                         "But the old man laughed. \"All I gave you was one ride, on one sled, in one snow, on one hill. I have a whole world of them in my memory. I could give them to you one by one, a thousand times, and there would still be more.\"",
                                         
                                         "\"Are you saying that I--I mean we--could do it again?\" Jonas asked. \"I'd really like to. I think I could steer, by pulling the rope. I didn't try this time, because it was so new.\"",
                                         "The old man, laughing, shook his head. \"Maybe another day, for a treat. But there's no time, really, just to play. I only wanted to begin by showing you how it works.\"",
                                         "\"Now,\" he said, turning businesslike, \"lie back down. I want to --\"",
                                         "Jonas did. He was eager for whatever experience would come next. But he had, suddenly, so many questions.",
                                         "\"Why don't we have snow, and sleds, and hills?\" he asked. \"And when did we, in the past? Did my parents have sleds when they were young? Did you?\"",
                                         "The old man shrugged and gave a short laugh. \"No,\" he told Jonas. \"It's a very distant memory. That's why it was so exhausting--I had to tug it forward from many generations back. It was given to me when I was a new Receiver, and the previous Receiver had to pull it through a long time period, too.\"",
                                         "\"But what happened to those things? Snow, and the rest of it?\"",
                                         "\"Climate Control. Snow made growing food difficult, limited the agricultural periods. And unpredictable weather made transportation almost impossible at times. It wasn't a practical thing, so it became obsolete when we went to Sameness.\"",
                                         "\"And hills, too,\" he added. \"They made conveyance of goods unwieldy. Trucks; buses. Slowed them down. So--\" He waved his hand, as if a gesture had caused hills to disappear. \"Sameness,\" he concluded.",
                                         "Jonas frowned. \"I wish we had those things, still. Just now and then.\"",
                                         "The old man smiled. \"So do I,\" he said. \"But that choice is not ours.\" \"But sir,\" Jonas suggested, \"since you have so much power--\"",
                                         "The man corrected him. \"Honor,\" he said firmly. \"I have great honor. So will you. But you will find that that is not the same as power.\"",
                                         "\"Lie quietly now. Since we've entered into the topic of climate, let me give you something else. And this time I'm not going to tell you the name of it, because I want to test the receiving. You should be able to perceive the name without being told. I gave away snow and sled and downhill and runners by telling them to you in advance.\"",
                                         "Without being instructed, Jonas closed his eyes again. He felt the hands on his back again. He waited.",
                                         "Now it came more quickly, the feelings. This time the hands didn't become cold, but instead began to feel warm on his body. They moistened a little. The warmth spread, extending across his shoulders, up his neck, onto the side of his face. He could feel it through his clothed parts, too: a pleasant, all-over sensation; and when he licked his lips this time, the air was hot and heavy.",
                                        "He didn't move. There was no sled. His posture didn't change. He was simply alone someplace, out of doors, lying down, and the warmth came from far above. It was not as exciting as the ride through the snowy air; but it was pleasurable and comforting.",
                                        "Suddenly he perceived the word for it: sunshine. He perceived that it came from the sky. Then it ended.",
                                        "\"Sunshine,\" he said aloud, opening his eyes.",
                                        "\"Good. You did get the word. That makes my job easier. Not so much explaining.\"",
                                        "\"And it came from the sky.\" \"That's right,\" the old man said. \"Just the way it used to.\"",
                                        "\"Before Sameness. Before Climate Control,\" Jonas added.",
                                        "The man laughed. \"You receive well, and learn quickly. I'm very pleased with you. That's enough for today, I think. We're off to a good start.\"",
                                        "There was a question bothering Jonas. \"Sir,\" he said, \"The Chief Elder told me--she told everyone--and you told me, too, that it would be painful. So I was a little scared. But it didn't hurt at all. I really enjoyed it.\" He looked quizzically at the old man.",
                                        "The man sighed. \"I started you with memories of pleasure. My previous failure gave me the wisdom to do that.\" He took a few deep breaths.\"Jonas,\" he said, \"it will be painful. But it need not be painful yet.\"",
                                        "\"I'm brave. I really am.\" Jonas sat up a little straighter. ",
                                        "The old man looked at him for a moment. He smiled.",
                                        "\"I can see that,\" he said. \"Well, since you asked the question--I think I have enough energy for one more transmission.\"",
                                        "\"Lie down once more. This will be the last today.\"",
                                        "Jonas obeyed cheerfully. He dosed his eyes, waiting, and felt the hands again; then he felt the warmth again, the sunshine again, coming from the sky of this other consciousness that was so new to him. This time, as he lay basking in the wonderful warmth, he felt the passage of time. His real self was aware that it was only a minute or two; but his other, memory-receiving self felt hours pass in the sun. His skin began to sting. Restlessly he moved one arm, bending it, and felt a sharp pain in the crease of his inner arm at the elbow.",
                                        "\"Ouch,\" he said loudly, and shifted on the bed. \"Owwww,\" he said, wincing at the shift, and even moving his mouth to speak made his face hurt.",
                                        "He knew there was a word, but the pain kept him from grasping it.",
                                        "Then it ended. He opened his eyes, wincing with discomfort. \"It hurt,\" he told the man, \"and I couldn't get the word for it.\"",
                                        "\"It was sunburn,\" the old man told him.",
                                        "\"It hurt a lot,\" Jonas said, \"but I'm glad you gave it to me. It was interesting. And now I understand better, what it meant, that there would be pain.\"",
                                        "The man didn't respond. He sat silently for a second. Finally he said, \"Get up, now. It's time for you to go home.\"",
                                        "They both walked to the center of the room. Jonas put his tunic back on. \"Goodbye, sir,\" he said. \"Thank you for my first day.\"",
                                        "The old man nodded to him. He looked drained, and a little sad. \"Sir?\" Jonas said shyly. \"Yes? Do you have a question?\"",
                                        "\"It's just that I don't know your name. I thought you were The Receiver, but you say that now I'm The Receiver. So I don't know what to call you.\"",
                                        "The man had sat back down in the comfortable upholstered chair. He moved his shoulders around as if to ease away an aching sensation. He seemed terribly weary.",
                                        "\"Call me The Giver,\" he told Jonas."))

```

```{r ch11plot}

ch11 <- annotate_ch11$token

#version 1
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Quantity",
       y = "Parts of Speech")

#version 2
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 11 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 11 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch11, head)


```
The above bar chart displays a count of each part of speech within Chapter 11 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

##Chapter 23

```{r ch23}
annotate_ch23 <- cnlp_annotate(input = c("Jonas felt more and more certain that the destination lay ahead of him, very near now in the night that was approaching. None of his senses confirmed it. He saw nothing ahead except the endless ribbon of road unfolding in twisting narrow curves. He heard no sound ahead.",
                                         "Yet he felt it: felt that Elsewhere was not far away. But he had little hope left that he would be able to reach it. His hope diminished further when the sharp, cold air began to blur and thicken with swirling white.",
                                         "Gabriel, wrapped in his inadequate blanket, was hunched, shivering, and silent in his little seat. Jonas stopped the bike wearily, lifted the child down, and realized with heartbreak how cold and weak Gabe had become.",
                                         "Standing in the freezing mound that was thickening around his numb feet, Jonas opened his own tunic, held Gabriel to his bare chest, and tied the torn and dirty blanket around them both. Gabriel moved feebly against him and whimpered briefly into the silence that surrounded them.",
                                         "Dimly, from a nearly forgotten perception as blurred as the substance itself, Jonas recalled what the whiteness was.",
                                         "\"It's called snow, Gabe,\" Jonas whispered. \"Snowflakes. They fall down from the sky, and they're very beautiful.\"",
                                         "There was no response from the child who had once been so curious and alert. Jonas looked down through the dusk at the little head against his chest. Gabriel's curly hair was matted and filthy, and there were tearstains outlined in dirt on his pale cheeks. His eyes were closed. As Jonas watched, a snowflake drifted down and was caught briefly for a moment's sparkle in the tiny fluttering eyelashes.",
                                         "Wearily he remounted the bicycle. A steep hill loomed ahead. In the best of conditions, the hill would have been a difficult, demanding ride. But now the rapidly deepening snow obscured the narrow road and made the ride impossible. His front wheel moved forward imperceptibly as he pushed on the pedals with his numb, exhausted legs. But the bicycle stopped. It would not move.",
                                         "He got off and let it drop sideways into the snow. For a moment he thought how easy it would be to drop beside it himself, to let himself and Gabriel slide into the softness of snow, the darkness of night, the warm comfort of sleep. But he had come this far. He must try to go on.",
                                         "The memories had fallen behind him now, escaping from his protection to return to the people of his community. Were there any left at all? Could he hold onto a last bit of warmth? Did he still have the strength to Give? Could Gabriel still Receive?",
                                         "He pressed his hands into Gabriel's back and tried to remember sunshine. For a moment it seemed that nothing came to him, that his power was completely gone. Then flickered suddenly, and he felt tiny tongues of heat begin to creep across and into his frozen feet and legs. He felt his face begin to glow and the tense, cold skin of his arms and hands relax. For a fleeting second he felt that he wanted to keep it for himself, to let himself bathe in sunlight, unburdened by anything or anyone else.",
                                         "But the moment passed and was followed by an urge, a need, a passionate yearning to share the warmth with the one person left for him to love. Aching from the effort, he forced the memory of warmth into the thin, shivering body in his arms.",
                                         "Gabriel stirred. For a moment they both were bathed in warmth and renewed strength as they stood hugging each other in the blinding snow.",
                                         "Jonas began to walk up the hill.",
                                         "The memory was agonizingly brief. He had trudged no more than a few yards through the night when it was gone and they were cold again.",
                                         "But his mind was alert now. Warming himself ever so briefly had shaken away the lethargy and resignation and restored his will to survive. He began to walk faster on feet that he could no longer feel. But the hill was treacherously steep; he was impeded by the snow and his own lack of strength. He didn't make it very far before he stumbled and fell forward.",
                                         "On his knees, unable to rise, Jonas tried a second time. His consciousness grasped at a wisp of another warm memory, and tried desperately to hold it there, to enlarge it, and pass it into Gabriel. His spirits and strength lifted with the momentary warmth and he stood. Again, Gabriel stirred against him as he began to climb.",
                                         "But the memory faded, leaving him colder than before. If only he had had time to receive more warmth from The Giver before he escaped! Maybe there would be more left for him now. But there was no purpose in if-onlys. His entire concentration now had to be on moving his feet, warming Gabriel and himself, and going forward.",
                                         "He climbed, stopped, and warmed them both briefly again, with a tiny scrap of memory that seemed certainly to be all he had left.",
                                         "The top of the hill seemed so far away, and he did not know what lay beyond. But there was nothing left to do but continue. He trudged upward.",
                                         "As he approached the summit of the hill at last, something began to happen. He was not warmer; if anything, he felt more numb and more cold. He was not less exhausted; on the contrary, his steps were leaden, and he could barely move his freezing, tired legs.",
                                         "But he began, suddenly, to feel happy. He began to recall happy times. He remembered his parents and his sister. He remembered his friends, Asher and Fiona. He remembered The Giver.",
                                         "Memories of joy flooded through him suddenly.",
                                         "He reached the place where the hill crested and he could feel the ground under his snow-covered feet become level. It would not be uphill anymore.",
                                         "\"We're almost there, Gabriel,\" he whispered, feeling quite certain without knowing why. \"I remember this place, Gabe.\" And it was true. But it was not a grasping of a thin and burdensome recollection; this was different. This was something that he could keep. It was a memory of his own.",
                                         "He hugged Gabriel and rubbed him briskly, warming him, to keep him alive. The wind was bitterly cold. The snow swirled, blurring his vision. But somewhere ahead, through the blinding storm, he knew there was warmth and light.",
                                         "Using his final strength, and a special knowledge that was deep inside him, Jonas found the sled that was waiting for them at the top of the hill. Numbly his hands fumbled for the rope.",
                                         "He settled himself on the sled and hugged Gabe close. The hill was steep but the snow was powdery and soft, and he knew that this time there would be no ice, no fall, no pain. Inside his freezing body, his heart surged with hope.",
                                         "They started down.",
                                         "Jonas felt himself losing consciousness and with his whole being willed himself to stay upright atop the sled, clutching Gabriel, keeping him safe. The runners sliced through the snow and the wind whipped at his face as they sped in a straight line through an incision that seemed to lead to the final destination, the place that he had always felt was waiting, the Elsewhere that held their future and their past.",
                                         "He forced his eyes open as they went downward, downward, sliding, and all at once he could see lights, and he recognized them now. He knew they were shining through the windows of rooms, that they were the red, blue, and yellow lights that twinkled from trees in places where families created and kept memories, where they celebrated love.",
                                         "Downward, downward, faster and faster. Suddenly he was aware with certainty and joy that below, ahead, they were waiting for him; and that they were waiting, too, for the baby. For the first time, he heard something that he knew to be music. He heard people singing.",
                                         "Behind him, across vast distances of space and time, from the place he had left, he thought he heard music too. But perhaps it was only an echo."))

```

```{r ch23plot}

ch23 <- annotate_ch23$token

#version 1
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Quantity",
       y = "Parts of Speech")

#version 2
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 23 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 23 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch23, head)

```
The above bar chart displays a count of each part of speech within Chapter 23 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

##Comparison Across Chapters

```{r verbcomparison}

verbch1 <- ch1 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch1")

verbch11 <- ch11 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch11")

verbch23 <- ch23 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch23")

verbcombined <- rbind(verbch1, verbch11, verbch23) %>%
  rename(chapters = new_col) %>%
  select(upos, chapters)

#version 1
verbcombined %>% 
  count(chapters) %>% 
  ggplot(aes(n, chapters)) +
  geom_col() +
  scale_y_discrete(breaks = paste0("ch", c("1", "11", "23")),
                   labels = c("1", "11", "23")) +
  labs(x = "Number of verbs",
       y = "Chapter")

#version 2
verbcombined %>% 
  count(chapters) %>% 
  #mutate(chapters = fct_relevel(chapters, 
                                "1",
                                "11")) %>%
  ggplot(aes(n, chapters)) +
  geom_col() +
  scale_y_discrete(breaks = paste0("ch", c("1", "11", "23")),
                   labels = c("1", "11", "23")) +
  labs(x = "Number of verbs",
       y = "Chapter")

# Interesting finding!! Do you think this could be because  some chapters are longer than others (see my % of words comment from earlier), or this is a meaningful trend? I would love to see in your final version an explanation for why you chose these chapters to look at (i'm assuming beg, middle, and end)/ why you chose verbs to look at. If you want to use gganimate like I suggested, you could put these three plots one after the other and highlight the "verb" bar a color like red. That way you can see how verbs change across chapters in the context of other parts of speech, e.g., you might find that verbs change a much smaller or much larger amount than other parts of speech across chapters

# This graph is extremely effective at demonstrating this effect. I immediately understood what was going on without even needing a title. Beyond what you have here, I think it would be really cool to look specifically at what verbs are most common in the giver. You could present that data in a word cloud, which makes more prevalent words bigger. Since you're interested in change across the book, you could even look at how the verbs change in prevalence across the book 

```
The above graph displays the amount of verbs present in chapters 1, 11, and 23. There is a downward trend in the amount of verbs as you move through the book. 

*I really love this figure! It raises so many questions! I think the use of bar plots was particularly effective here. I'm curious about the possibilities of adding an additional variable to this figure - something like "complexity" or a different metric that shows whether the verbs change in quality as well as quantity over the course of the book. Also, I wonder whether this is impacted by a change in length (i.e., is chapter 1 super long?). That may be a fun aspect to play with in a future draft (or just with this data set)*

*I think you picked great forms for the data you picked, and did a really good job navigating complex data - NLP is tough stuff! I haven't poked around NLP done in R before, so definitely learned a lot examining your code!*
