---
title: "Final Project"
author: "Makayla Whitney"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
#py_install("cleannlp")
#system("pip install cleannlp")
#cnlp_init_spacy(model_name = "en")

#These weren't working for me?? not sure what they are supposed to be doing but i was able to run the rest of the code installing cleannlp from r and not python
```

```{r code_setup}

#install.packages("cleanNLP")
library(cleanNLP)
library(dplyr)
library(tidyr)
#install.packages("gtsummary")
library(gtsummary)
library(magrittr)
library(ggplot2)
library(forcats)
cnlp_init_udpipe()


```

``` {r test}

#this code chunk was to test and learn the package

annotate_test <- cnlp_annotate(input = c(
  "Here is the first text. It is short.",
  "Here's the second. It is short too!",
  "The third text is the shortest."))


d <- annotate_test$token

d %>% 
  count(xpos) %>% 
  filter(xpos != ".") %>% 
  ggplot(aes(n, xpos)) +
  geom_col()


lapply(annotate_test, head)


cat("This is a normal string in R",
  "This is some text \"That has a quote\" and then some other text",
  sep = "\n")

```

#Introduction

14% of adults in the United States are functionally illiterate. "Functionally illiterate" is defined as reading below the 4th grade reading level. Many adults enroll in literacy tutoring programs that assist them with increasing their reading level to either, 1) complete their GED, 2) enroll in college or enter the workforce, 3) assist their children at home, or 4) other personal learning goals. These agencies are not equip with adult reading materials purely because they don't exist. Instead, agencies provide their tutors with children's books that are complementary to their adult learner's reading level. Children books are not topically engaging or relevant to most adult learners. I examine the parts of speech found within *The Giver* - a book found in most 8th grade curriculum across the country - to help conceptualize what concepts should be taught at the 8th grade level. This project is the beginning of a large-scale examination of 4th through 8th grade books to assist with creating a formula to be used in establishing adult reading materials. 

##Chapter 1

```{r ch1}
annotate_ch1 <- cnlp_annotate(input = c("It was almost December, and Jonas was beginning to be frightened.",
                                        "No. Wrong word, Jonas thought.",
                                        "Frightened meant that deep, sickening feeling of something terrible about to happen.",
                                        "Frightened was the way he had felt a year ago when an unidentified aircraft had overflown the community twice.",
                                        "He had seen it both times.",
                                        "Squinting toward the sky, he had seen the sleek jet, almost a blur at its high speed, go past, and a second later heard the blast of sound that followed.",
                                        "Then one more time, a moment later, from the opposite direction, the same plane.",
                                        "At first, he had been only fascinated.",
                                        "He had never seen aircraft so close, for it was against the rules for Pilots to fly over the community.",
                                        "Occasionally, when supplies were delivered by cargo planes to the landing field across the river, the children rode their bicycles to the riverbank and watched, intrigued, the unloading and then the takeoff directed to the west, always away from the community.",
                                        "But the aircraft a year ago had been different.",
                                        "It was not a squat, fat-bellied cargo plane but a needle-nosed single-pilot jet.",
                                        "Jonas, looking around anxiously, had seen others--adults as well as children--stop what they were doing and wait, confused, for an explanation of the frightening event.",
                                        "Then all of the citizens had been ordered to go into the nearest building and stay there. IMMEDIATELY, the rasping voice through the speakers had said. LEAVE YOUR BICYCLES WHERE THEY ARE.",
                                        "Instantly, obediently, Jonas had dropped his bike on its side on the path behind his family's dwelling. He had run indoors and stayed there, alone. His parents were both at work, and his little sister, Lily, was at the Childcare Center where she spent her after-school hours.",
                                        "Looking through the front window, he had seen no people: none of the busy afternoon crew of Street Cleaners, Landscape Workers, and Food Delivery people who usually populated the community at that time of day. He saw only the abandoned bikes here and there on their sides; an upturned wheel on one was still revolving slowly.",
                                        "He had been frightened then. The sense of his own community silent, waiting, had made his stomach churn. He had trembled.",
                                        "But it had been nothing. Within minutes the speakers had crackled again, and the voice, reassuring now and less urgent, had explained that a Pilot-in-Training had misread his navigational instructions and made a wrong turn. Desperately the Pilot had been trying to make his way back before his error was noticed.",
                                        "NEEDLESS TO SAY, HE WILL BE RELEASED, the voice had said, followed by silence. There was an ironic tone to that final message, as if the Speaker found it amusing; and Jonas had smiled a little, though he knew what a grim statement it had been. For a contributing citizen to be released from the community was a final decision, a terrible punishment, an overwhelming statement of failure.",
                                        "Even the children were scolded if they used the term lightly at play, jeering at a teammate who missed a catch or stumbled in a race.",
                                        "Jonas had done it once, had shouted at his best friend, \"That's it, Asher! You're released!\" when Asher's clumsy error had lost a match for his team. He had been taken aside for a brief and serious talk by the coach, had hung his head with guilt and embarrassment, and apologized to Asher after the game.",
                                        "Now, thinking about the feeling of fear as he pedaled home along the river path, he remembered that moment of palpable, stomach- sinking terror when the aircraft had streaked above. It was not what he was feeling now with December approaching. He searched for the right word to describe his own feeling.",
                                        "Jonas was careful about language. Not like his friend, Asher, who talked too fast and mixed things up, scrambling words and phrases until they were barely recognizable and often very funny.",
                                        "Jonas grinned, remembering the morning that Asher had dashed into the classroom, late as usual, arriving breathlessly in the middle of the chanting of the morning anthem. When the class took their seats at the conclusion of the patriotic hymn, Asher remained standing to make his public apology as was required.",
                                        "\"I apologize for inconveniencing my learning community.\" Asher ran through the standard apology phrase rapidly, still catching his breath. The Instructor and class waited patiently for his explanation. The students had all been grinning, because they had listened to Asher's explanations so many times before.",
                                        "\"I left home at the correct time but when I was riding along near the hatchery, the crew was separating some salmon. I guess I just got distraught, watching them.\"",
                                        "\"I apologize to my classmates,\" Asher concluded. He smoothed his rumpled tunic and sat down.",
                                        "\"We accept your apology, Asher.\" The class recited the standard response in unison. Many of the students were biting their lips to keep from laughing.",
                                        "\"I accept your apology, Asher,\" the Instructor said. He was smiling. \"And I thank you, because once again you have provided an opportunity for a lesson in language. 'Distraught' is too strong an adjective to describe salmon-viewing.\" He turned and wrote \"distraught\" on the instructional board. Beside it he wrote \"distracted.\"",
                                        "Jonas, nearing his home now, smiled at the recollection. Thinking, still, as he wheeled his bike into its narrow port beside the door, he realized that frightened was the wrong word to describe his feelings, now that December was almost here. It was too strong an adjective.",
                                        "He had waited a long time for this special December. Now that it was almost upon him, he wasn't frightened, but he was ... eager, he decided. He was eager for it to come. And he was excited, certainly. All of the Elevens were excited about the event that would be coming so soon.",
                                        "But there was a little shudder of nervousness when he thought about it, about what might happen.",
                                        "Apprehensive, Jonas decided. That's what I am.",
                                        "\"Who wants to be the first tonight, for feelings?\" Jonas's father asked, at the conclusion of their evening meal.",
                                        "It was one of the rituals, the evening telling of feelings.",
                                        "Jonas and his sister, Lily, argued over turns, over who would get to go first. Their parents, of course, were part of the ritual; they, too, told their feelings each evening. But like all parents - all adults - they didn't fight and wheedle for their turn.",
                                        "Nor did Jonas, tonight. His feelings were too complicated this evening. He wanted to share them, but he wasn't eager to begin the process of sifting through his own complicated emotions, even with the help that he knew his parents could give.",
                                        "\"You go, Lily,\" he said, seeing his sister, who was much younger-- only a Seven--wiggling with impatience in her chair.",
                                        "\"I felt very angry this afternoon,\" Lily announced. \"My Childcare group was at the play area, and we had a visiting group of Sevens, and they didn't obey the rules at all. One of them--a male; I don't know his name--kept going right to the front of the line for the slide, even though the rest of us were all waiting. I felt so angry at him. I made my hand into a fist, like this.\" She held up a clenched fist and the rest of the family smiled at her small defiant gesture.",
                                        "\"Why do you think the visitors didn't obey the rules?\" Mother asked.",
                                        "Lily considered, and shook her head. \"I don't know. They acted like...like...\"",
                                        "\"Animals?\" Jonas suggested. He laughed.",
                                        "\"That's right,\" Lily said, laughing too. \"Like animals.\" Neither child knew what the word meant, exactly, but it was often used to describe someone uneducated or clumsy, someone who didn't fit in.",
                                        "\"Where were the visitors from?\" Father asked.",
                                        "Lily frowned, trying to remember. \"Our leader told us, when he made the welcome speech, but I can't remember. I guess I wasn't paying attention. It was from another community. They had to leave very early, and they had their midday meal on the bus.\"",
                                        "Mother nodded. \"Do you think it's possible that their rules may be different? And so they simply didn't know what your play area rules were?\"",
                                        "Lily shrugged, and nodded. \"I suppose.\"",
                                        "\"You've visited other communities, haven't you?\" Jonas asked. \"My group has, often.\"",
                                        "Lily nodded again. \"When we were Sixes, we went and shared a whole school day with a group of Sixes in their community.\"",
                                        "\"How did you feel when you were there?\"",
                                        "Lily frowned. \"I felt strange. Because their methods were different. They were learning usages that my group hadn't learned yet, so we felt stupid.\"",
                                        "Father was listening with interest. \"I'm thinking, Lily,\" he said, \"about the boy who didn't obey the rules today. Do you think it's possible that he felt strange and stupid, being in a new place with rules that he didn't know about?\"",
                                        "Lily pondered that. \"Yes,\" she said, finally.",
                                        "\"I feel a little sorry for him,\" Jonas said, \"even though I don't even know him. I feel sorry for anyone who is in a place where he feels strange and stupid.\"",
                                        "\"How do you feel now, Lily?\" Father asked. \"Still angry?\"",
                                        "\"I guess not,\" Lily decided. \"I guess I feel a little sorry for him. And sorry I made a fist.\" She grinned.",
                                        "Jonas smiled back at his sister. Lily's feelings were always straightforward, fairly simple, usually easy to resolve. He guessed that his own had been, too, when he was a Seven.",
                                        "He listened politely, though not very attentively, while his father took his turn, describing a feeling of worry that he'd had that day at work: a concern about one of the newchildren who wasn't doing well. Jonas's father's title was Nurturer. He and the other Nurturers were responsible for all the physical and emotional needs of every newchild during its earliest life. It was a very important job, Jonas knew, but it wasn't one that interested him much.",
                                        "\"What gender is it?\" Lily asked.",
                                        "\"Male,\" Father said. \"He's a sweet little male with a lovely disposition. But he isn't growing as fast as he should, and he doesn't sleep soundly. We have him in the extra care section for supplementary nurturing, but the committee's beginning to talk about releasing him.\"",
                                        "\"Oh, no,\" Mother murmured sympathetically. \"I know how sad that must make you feel.\"",
                                        "Jonas and Lily both nodded sympathetically as well. Release of newchildren was always sad, because they hadn't had a chance to enjoy life within the community yet. And they hadn't done anything wrong.",
                                        "There were only two occasions of release which were not punishment. Release of the elderly, which was a time of celebration for a life well and fully lived; and release of a newchild, which always brought a sense of what-could-we- have-done. This was especially troubling for the Nurturers, like Father, who felt they had failed somehow. But it happened very rarely.",
                                        "\"Well,\" Father said, \"I'm going to keep trying. I may ask the committee for permission to bring him here at night, if you don't mind. You know what the night-crew Nurturers are like. I think this little guy needs something extra.\"",
                                        "\"Of course,\" Mother said, and Jonas and Lily nodded. They had heard Father complain about the night crew before. It was a lesser job, night-crew nurturing, assigned to those who lacked the interest or skills or insight for the more vital jobs of the daytime hours. Most of the people on the night crew had not even been given spouses because they lacked, somehow, the essential capacity to connect to others, which was required for the creation of a family unit.",
                                        "\"Maybe we could even keep him,\" Lily suggested sweetly, trying to look innocent. The look was fake, Jonas knew; they all knew.",
                                        "\"Lily,\" Mother reminded her, smiling, \"you know the rules.\"",
                                        "Two children--one male, one female--to each family unit. It was written very clearly in the rules.",
                                        "Lily giggled. \"Well,\" she said, \"I thought maybe just this once.\"",
                                        "Next, Mother, who held a prominent position at the Department of Justice, talked about her feelings. Today a repeat offender had been brought before her, someone who had broken the rules before. Someone who she hoped had been adequately and fairly punished, and who had been restored to his place: to his job, his home, his family unit. To see him brought before her a second time caused her overwhelming feelings of frustration and anger. And even guilt, that she hadn't made a difference in his life.",
                                        "\"I feel frightened, too, for him,\" she confessed. \"You know that there's no third chance. The rules say that if there's a third transgression, he simply has to be released.\" Jonas shivered. He knew it happened. There was even a boy in his group of Elevens whose father had been released years before. No one ever mentioned it; the disgrace was unspeakable. It was hard to imagine.",
                                        "Lily stood up and went to her mother. She stroked her mother's arm.",
                                        "From his place at the table, Father reached over and took her hand. Jonas reached for the other.",
                                        "One by one, they comforted her. Soon she smiled, thanked them, and murmured that she felt soothed.",
                                        "The ritual continued. \"Jonas?\" Father asked. \"You're last, tonight.\"",
                                        "Jonas sighed. This evening he almost would have preferred to keep his feelings hidden. But it was, of course, against the rules.",
                                        "\"I'm feeling apprehensive,\" he confessed, glad that the appropriate descriptive word had finally come to him. \"Why is that, son?\" His father looked concerned.",
                                        "\"I know there's really nothing to worry about,\" Jonas explained, \"and that every adult has been through it. I know you have, Father, and you too, Mother. But it's the Ceremony that I'm apprehensive about. It's almost December.\"",
                                        "Lily looked up, her eyes wide. \"The Ceremony of Twelve,\" she whispered in an awed voice. Even the smallest children--Lily's age and younger--knew that it lay in the future for each of them.",
                                        "\"I'm glad you told us of your feelings,\" Father said.",
                                        "\"Lily,\" Mother said, beckoning to the little girl, \"Go on now and get into your nightclothes. Father and I are going to stay here and talk to Jonas for a while.\"",
                                        "Lily sighed, but obediently she got down from her chair. \"Privately?\" she asked.",
                                        "Mother nodded. \"Yes,\" she said, \"this talk will be a private one with Jonas.\"" ))
```

```{r ch1plot}

ch1 <- annotate_ch1$token

#version 1
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".") %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "",
       y = "",
       title = "Parts of Speech")

#version 2
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 1 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch1 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 1 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch1, head)

#summarize the percentages of parts of speech present in ch1 
ch1 %>% 
  select(upos) %>%
  tbl_summary()


```
The above bar chart displays a count of each part of speech within Chapter 1 of *The Giver*. 

##Chapter 2
```{r ch2}

annotate_ch2 <- cnlp_annotate(input = c("Jonas watched as his father poured a fresh cup of coffee. He waited.",
                                        "\"You know,\" his father finally said, \"every December was exciting to me when I was young. And it has been for you and Lily, too, I'm sure. Each December brings such changes.\"",
                                        "Jonas nodded. He could remember the Decembers back to when he had become, well, probably a Four. The earlier ones were lost to him. But he observed them each year, and he remembered Lily's earliest Decembers. He remembered when his family received Lily, the day she was named, the day that she had become a One.",
                                        "The Ceremony for the Ones was always noisy and fun. Each December, all the newchildren born in the previous year turned One. One at a time--there were always fifty in each year's group, if none had been released --they had been brought to the stage by the Nurturers who had cared for them since birth. Some were already walking, wobbly on their unsteady legs; others were no more than a few days old, wrapped in blankets, held by their Nurturers.",
                                        "\"I enjoy the Naming,\" Jonas said.",
                                        "His mother agreed, smiling. \"The year we got Lily, we knew, of course, that we'd receive our female, because we'd made our application and been approved. But I'd been wondering and wondering what her name would be.\"",
                                        "\"I could have sneaked a look at the list prior to the ceremony,\" Father confided. \"The committee always makes the list in advance, and it's right there in the office at the Nurturing Center.\"",
                                        "\"As a matter of fact,\" he went on, \"I feel a little guilty about this. But I did go in this afternoon and looked to see if this year's Naming list had been made yet. It was right there in the office, and I looked up number Thirty-six--that's the little guy I've been concerned about-- because it occurred to me that it might enhance his nurturing if I could call him by a name. Just privately, of course, when no one else is around.\"",
                                        "\"Did you find it?\" Jonas asked. He was fascinated. It didn't seem a terribly important rule, but the fact that his father had broken a rule at all awed him. He glanced at his mother, the one responsible for adherence to the rules, and was relieved that she was smiling.",
                                        "His father nodded. \"His name--if he makes it to the Naming without being released, of course--is to be Gabriel. So I whisper that to him when I feed him every four hours, and during exercise and playtime. If no one can hear me.\"",
                                        "\"I call him Gabe, actually,\" he said, and grinned. \"Gabe.\" Jonas tried it out. A good name, he decided. Though Jonas had only become a Five the year that they acquired Lily and learned her name, he remembered the excitement, the conversations at home, wondering about her: how she would look, who she would be, how she would fit into their established family unit. He remembered climbing the steps to the stage with his parents, his father by his side that year instead of with the Nurturers, since it was the year that he would be given a newchild of his own.",
                                        "He remembered his mother taking the newchild, his sister, into her arms, while the document was read to the assembled family units. \"Newchild Twenty-three,\" the Namer had read. \"Lily.\"",
                                        "He remembered his father's look of delight, and that his father had whispered, \"She's one of my favorites. I was hoping for her to be the one.\" The crowd had clapped, and Jonas had grinned. He liked his sister's name. Lily, barely awake, had waved her small fist. Then they had stepped down to make room for the next family unit.",
                                        "\"When I was an Eleven,\" his father said now, \"as you are, Jonas, I was very impatient, waiting for the Ceremony of Twelve. It's a long two days. I remember that I enjoyed the Ones, as I always do, but that I didn't pay much attention to the other ceremonies, except for my sister's. She became a Nine that year, and got her bicycle. I'd been teaching her to ride mine, even though technically I wasn't supposed to.\"",
                                        "Jonas laughed. It was one of the few rules that was not taken very seriously and was almost always broken. The children all received their bicycles at Nine; they were not allowed to ride bicycles before then. But almost always, the older brothers and sisters had secretly taught the younger ones. Jonas had been thinking already about teaching Lily.",
                                        "There was talk about changing the rule and giving the bicycles at an earlier age. A committee was studying the idea. When something went to a committee for study, the people always joked about it. They said that the committee members would become Elders by the time the rule change was made.",
                                        "Rules were very hard to change. Sometimes, if it was a very important rule--unlike the one governing the age for bicycles--it would have to go, eventually, to The Receiver for a decision. The Receiver was the most important Elder. Jonas had never even seen him, that he knew of; someone in a position of such importance lived and worked alone. But the committee would never bother The Receiver with a question about bicycles; they would simply fret and argue about it themselves for years, until the citizens forgot that it had ever gone to them for study.",
                                        "His father continued. \"So I watched and cheered when my sister, Katya, became a Nine and removed her hair ribbons and got her bicycle,\" Father went on. \"Then I didn't pay much attention to the Tens and Elevens. And finally, at the end of the second day, which seemed to go on forever, it was my turn. It was the Ceremony of Twelve.\"",
                                        "Jonas shivered. He pictured his father, who must have been a shy and quiet boy, for he was a shy and quiet man, seated with his group, waiting to be called to the stage. The Ceremony of Twelve was the last of the Ceremonies. The most important.",
                                        "I remember how proud my parents looked--and my sister, too; even though she wanted to be out riding the bicycle publicly, she stopped fidgeting and was very still and attentive when my turn came.",
                                        "\But to be honest, Jonas,\" his father said, \"for me there was not the element of suspense that there is with your Ceremony. Because I was already fairly certain of what my Assignment was to be.\"",
                                        "Jonas was surprised. There was no way, really, to know in advance. It was a secret selection, made by the leaders of the community, the Committee of Elders, who took the responsibility so seriously that there were never even any jokes made about Assignments.",
                                        "His mother seemed surprised, too. \"How could you have known?\" she asked.",
                                        "His father smiled his gentle smile. \"Well, it was clear to me--and my parents later confessed that it had been obvious to them, too--what my aptitude was. I had always loved the newchildren more than anything. When my friends in my age group were holding bicycle races, or building toy vehicles or bridges with their construction sets, or--\"",
                                        "\"All the things I do with my friends,\" Jonas pointed out, and his mother nodded in agreement.",
                                        "\"I always participated, of course, because as children we must experience all of those things. And I studied hard in school, as you do, Jonas. But again and again, during free time, I found myself drawn to the newchildren. I spent almost all of my volunteer hours helping in the Nurturing Center. Of course the Elders knew that, from their observation.\"",
                                        "Jonas nodded. During the past year he had been aware of the increasing level of observation. In school, at recreation time, and during volunteer hours, he had noticed the Elders watching him and the other Elevens. He had seen them taking notes. He knew, too, that the Elders were meeting for long hours with all of the instructors that he and the other Elevens had had during their years of school.",
                                        "\"So I expected it, and I was pleased, but not at all surprised, when my Assignment was announced as Nurturer,\" Father explained.",
                                        "\"Did everyone applaud, even though they weren't surprised?\" Jonas asked.",
                                        "\"Oh, of course. They were happy for me, that my Assignment was what I wanted most. I felt very fortunate.\" His father smiled.",
                                        "\"Were any of the Elevens disappointed, your year?\" Jonas asked. Unlike his father, he had no idea what his Assignment would be. But he knew that some would disappoint him. Though he respected his father's work, Nurturer would not be his wish. And he didn't envy Laborers at all.",
                                        "His father thought. \"No, I don't think so. Of course the Elders are so careful in their observations and selections.\"",
                                        "\"I think it's probably the most important job in our community,\" his mother commented.",
                                        "\"My friend Yoshiko was surprised by her selection as Doctor,\" Father said, \"but she was thrilled. And let's see, there was Andrei--I remember that when we were boys he never wanted to do physical things. He spent all the recreation time he could with his construction set, and his volunteer hours were always on building sites. The Elders knew that, of course. Andrei was given the Assignment of Engineer and he was delighted.\"",
                                        "\"Andrei later designed the bridge that crosses the river to the west of town,\" Jonas's mother said. \"It wasn't there when we were children.\"",
                                        "\"There are very rarely disappointments, Jonas. I don't think you need to worry about that,\" his father reassured him. \"And if there are, you know there's an appeal process.\" But they all laughed at that--an appeal went to a committee for study.",
                                        "\"I worry a little about Asher's Assignment,\" Jonas confessed. \"Asher's such fun. But he doesn't really have any serious interests. He makes a game out of everything.\"",
                                        "His father chuckled. \"You know,\" he said, \"I remember when Asher was a newchild at the Nurturing Center, before he was named. He never cried. He giggled and laughed at everything. All of us on the staff enjoyed nurturing Asher.\"",
                                        "\"The Elders know Asher,\" his mother said. \"They'll find exactly the right Assignment for him. I don't think you need to worry about him. But, Jonas, let me warn you about something that may not have occurred to you. I know I didn't think about it until after my Ceremony of Twelve.\"",
                                        "\"What's that?\"",
                                        "\"Well, it's the last of the Ceremonies, as you know. After Twelve, age isn't important. Most of us even lose track of how old we are as time passes, though the information is in the Hall of Open Records, and we could go and look it up if we wanted to. What's important is the preparation for adult life, and the training you'll receive in your Assignment.\"",
                                        "\"I know that,\" Jonas said. \"Everyone knows that.\"",
                                        "\"But it means,\" his mother went on, \"that you'll move into a new group. And each of your friends will. You'll no longer be spending your time with your group of Elevens. After the Ceremony of Twelve, you'll be with your Assignment group, with those in training. No more volunteer hours. No more recreation hours. So your friends will no longer be as close.\"",
                                        "Jonas shook his head. \"Asher and I will always be friends,\" he said firmly. \"And there will still be school.\"",
                                        "\"That's true,\" his father agreed. \"But what your mother said is true as well. There will be changes.\"",
                                        "\"Good changes, though,\" his mother pointed out. \"After my Ceremony of Twelve, I missed my childhood recreation. But when I entered my training for Law and Justice, I found myself with people who shared my interests. I made friends on a new level, friends of all ages.\"",
                                        "\"Did you still play at all, after Twelve?\" Jonas asked.",
                                        "\"Occasionally,\" his mother replied. \"But it didn't seem as important to me.\"",
                                        "\"I did,\" his father said, laughing. \"I still do. Every day, at the Nurturing Center, I play bounce-on-the-knee, and peek-a-boo, and hug-the-teddy.\" He reached over and stroked Jonas's neatly trimmed hair. \"Fun doesn't end when you become Twelve.\"",
                                        "Lily appeared, wearing her nightclothes, in the doorway. She gave an impatient sigh. \"This is certainly a very long private conversation,\" she said. \"And there are certain people waiting for their comfort object.\"",
                                        "\"Lily,\" her mother said fondly, \"you're very dose to being an Eight, and when you're an Eight, your comfort object will be taken away. It will be recycled to the younger children. You should be starting to go off to sleep without it.\"",
                                        "But her father had already gone to the shelf and taken down the stuffed elephant which was kept there. Many of the comfort objects, like Lily's, were soft, stuffed, imaginary creatures. Jonas's had been called a bear.",
                                        "\"Here you are, Lily-billy,\" he said. \"I'll come help you remove your hair ribbons.\"",
                                        "Jonas and his mother rolled their eyes, yet they watched affectionately as Lily and her father headed to her sleeping-room with the stuffed elephant that had been given to her as her comfort object when she was born. His mother moved to her big desk and opened her briefcase; her work never seemed to end, even when she was at home in the evening. Jonas went to his own desk and began to sort through his school papers for the evening's assignment. But his mind was still on December and the coming Ceremony.",
                                        "Though he had been reassured by the talk with his parents, he hadn't the slightest idea what Assignment the Elders would be selecting for his future, or how he might feel about it when the day came."))

```

##Chapter 3
```{r ch3}
annotate_ch3 <- cnlp_annotate(input = c("\"Oh, look!\" Lily squealed in delight. \"Isn't he cute? Look how tiny he is! And he has funny eyes like yours, Jonas!\" Jonas glared at her. He didn't like it that she had mentioned his eyes. He waited for his father to chastise Lily. But Father was busy unstrapping the carrying basket from the back of his bicycle. Jonas walked over to look.",
                                        "It was the first thing Jonas noticed as he looked at the newchild peering up curiously from the basket. The pale eyes.",
                                        "Almost every citizen in the community had dark eyes. His parents did, and Lily did, and so did all of his group members and friends. But there were a few exceptions: Jonas himself, and a female Five who he had noticed had the different, lighter eyes. No one mentioned such things; it was not a rule, but was considered rude to call attention to things that were unsettling or different about individuals. Lily, he decided, would have to learn that soon, or she would be called in for chastisement because of her insensitive chatter.",
                                        "Father put his bike into its port. Then he picked up the basket and carried it into the house. Lily followed behind, but she glanced back over her shoulder at Jonas and teased, \"Maybe he had the same Birthmother as you.\"",
                                        "Jonas shrugged. He followed them inside. But he had been startled by the newchild's eyes. Mirrors were rare in the community; they weren't forbidden, but there was no real need of them, and Jonas had simply never bothered to look at himself very often even when he found himself in a location where a mirror existed. Now, seeing the newchild and its expression, he was reminded that the light eyes were not only a rarity but gave the one who had them a certain look--what was it? Depth, he decided; as if one were looking into the clear water of the river, down to the bottom, where things might lurk which hadn't been discovered yet. He felt self-conscious, realizing that he, too, had that look.",
                                        "He went to his desk, pretending not to be interested in the newchild. On the other side of the room, Mother and Lily were bending over to watch as Father unwrapped its blanket.",
                                        "\"What's his comfort object called?\" Lily asked, picking up the stuffed creature which had been placed beside the newchild in his basket.",
                                        "Father glanced at it. \"Hippo,\" he said.",
                                        "Lily giggled at the strange word. \"Hippo,\" she repeated, and put the comfort object down again. She peered at the unwrapped newchild, who waved his arms.",
                                        "\"I think newchildren are so cute,\" Lily sighed. \"I hope I get assigned to be a Birthmother.\"",
                                        "\"Lily!\" Mother spoke very sharply. \"Don't say that. There's very little honor in that Assignment.\"",
                                        "\"But I was talking to Natasha. You know the Ten who lives around the corner? She does some of her volunteer hours at the Birthing Center. And she told me that the Birthmothers get wonderful food, and they have very gentle exercise periods, and most of the time they just play games and amuse themselves while they're waiting. 1 think I'd like that,\" Lily said petulantly.",
                                        "\"Three years,\" Mother told her firmly. \"Three births and that's all. After that they are Laborers for the rest of their adult lives, until the day that they enter the House of the Old. Is that what you want, Lily? Three lazy years, and then hard physical labor until you are old?\"",
                                        "\"Well, no, I guess not,\" Lily acknowledged reluctantly Father turned the newchild onto his tummy in the basket. He sat beside it and rubbed its small back with a rhythmic motion. \"Anyway, Lily-billy,\" he said affectionately, \"the Birthmothers never even get to see newchildren. If you enjoy the little ones so much, you should hope for an Assignment as Nurturer.\"",
                                        "\"When you're an Eight and start your volunteer hour, you can try some at the Nurturing Center,\" Mother suggested.",
                                        "\"Yes, I think I will,\" Lily said. She knelt beside the basket. \"What did you say his name is? Gabriel? Hello, Gabriel,\" she said in a singsong voice. Then she giggled \"Ooops,\" she whispered. \"I think he's alseep. I guess I'd better be quiet.\"",
                                        "Jonas turned to the school assignments on his desk Some chance of that, he thought. Lily was never quiet Probably she should hope for an Assignment as Speaker so that she could sit in the office with the microphone a1l day, making announcements. He laughed silently to himself, picturing his sister droning on in the self- important voice that all the Speakers seemed to develop, saying things like, ATTENTION. THIS IS A REMINDER TO FEMALES UNDER NINE THAT HAIR RIBBONS ARE TO BE NEATLY TIED AT ALL TIMES.",
                                        "He turned toward Lily and noticed to his satisfaction that her ribbons were, as usual, undone and dangling. There would be an announcement like that quite soon, he felt certain, and it would be directed mainly at Lily, though her name, of course, would not be mentioned. Everyone would know.",
                                        "Everyone had known, he remembered with humiliation, that the announcement ATTENTION. THIS IS A REMINDER TO MALE ELEVENS THAT OBJECTS ARE NOT TO BE REMOVED FROM THE RECREATION AREA AND THAT SNACKS ARE TO BE EATEN, NOT HOARDED had been specifically directed at him, the day last month that he had taken an apple home. No one had mentioned it, not even his parents, because the public announcement had been sufficient to produce the appropriate remorse. He had, of course, disposed of the apple and made his apology to the Recreation Director the next morning, before school.",
                                        "Jonas thought again about that incident. He was still bewildered by it. Not by the announcement or the necessary apology; those were standard procedures, and he had deserved them--but by the incident itself. He probably should have brought up his feeling of bewilderment that very evening when the family unit had shared their feelings of the day. But he had not been able to sort out and put words to the source of his confusion, so he had let it pass.",
                                        "It had happened during the recreation period, when he had been playing with Asher. Jonas had casually picked up an apple from the basket where the snacks were kept, and had thrown it to his friend. Asher had thrown it back, and they had begun a simple game of catch.",
                                        "There had been nothing special about it; it was an activity that he had performed countless times: throw, catch; throw, catch. It was effortless for Jonas, and even boring, though Asher enjoyed it, and playing catch was a required activity for Asher because it would improve his hand-eye coordination, which was not up to standards.",
                                        "But suddenly Jonas had noticed, following the path of the apple through the air with his eyes, that the piece of fruit had--well, this was the part that he couldn't adequately understand--the apple had changed. Just for an instant. It had changed in mid-air, he remembered. Then it was in his hand, and he looked at it carefully, but it was the same apple. Unchanged. The same size and shape: a perfect sphere. The same nondescript shade, about the same shade as his own tunic.",
                                        "There was absolutely nothing remarkable about that apple. He had tossed it back and forth between his hands a few times, then thrown it again to Asher. And again--in the air, for an instant only--it had changed.",
                                        "It had happened four times. Jonas had blinked, looked around, and then tested his eyesight, squinting at the small print on the identification badge attached to his tunic. He read his name quite clearly. He could also clearly see Asher at the other end of the throwing area. And he had had no problem catching the apple.",
                                        "Jonas had been completely mystified.",
                                        "\"Ash?\" he had called. \"Does anything seem strange to you? About the apple?\"",
                                        "\"Yes,\" Asher called back, laughing. \"It jumps out of my hand onto the ground!\" Asher had just dropped it once again.",
                                        "So Jonas laughed too, and with his laughter tried to ignore his uneasy conviction that something had happened. But he had taken the apple home, against the recreation area rules. That evening, before his parents and Lily arrived at the dwelling, he had held it in his hands and looked at it carefully. It was slightly bruised now, because Asher had dropped it several times. But there was nothing at all unusual about the apple.",
                                        "He had held a magnifying glass to it. He had tossed it several times across the room, watching, and then rolled it around and around on his desktop, waiting for the thing to happen again.",
                                        "But it hadn't. The only thing that happened was the announcement later that evening over the speaker, the announcement that had singled him out without using his name, that had caused both of his parents to glance meaningfully at his desk where the apple still lay.",
                                        "Now, sitting at his desk, staring at his schoolwork as his family hovered over the newchild in its basket, he shook his head, trying to forget the odd incident. He forced himself to arrange his papers and try to study a little before the evening meal. The newchild, Gabriel, stirred and whimpered, and Father spoke softly to Lily, explaining the feeding procedure as he opened the container that held the formula and equipment.",
                                        "The evening proceeded as all evenings did in the family unit, in the dwelling, in the community: quiet, reflective, a time for renewal and preparation for the day to come. It was different only in the addition to it of the newchild with his pale, solemn, knowing eyes."))
```

##Chapter 4
```{r ch4}
annotate_ch4 <- cnlp_annotate(input = c("Jonas rode at a leisurely pace, glancing at the bikeports beside the buildings to see if he could spot Asher's. He didn't often do his volunteer hours with his friend because Asher frequently fooled around and made serious work a little difficult. But now, with Twelve coming so soon and the volunteer hours ending, it didn't seem to matter.",
                                        "The freedom to choose where to spend those hours had always seemed a wonderful luxury to Jonas; other hours of the day were so carefully regulated.",
                                        "He remembered when he had become an Eight, as Lily would do shortly, and had been faced with that freedom of choice. The Eights always set out on their first volunteer hour a little nervously, giggling and staying in groups of friends. They almost invariably did their hours on Recreation Duty first, helping with the younger ones in a place where they still felt comfortable. But with guidance, as they developed self-confidence and maturity, they moved on to other jobs, gravitating toward those that would suit their own interests and skills.",
                                        "A male Eleven named Benjamin had done his entire nearly-Four years in the Rehabilitation Center, working with citizens who had been injured. It was rumored that he was as skilled now as the Rehabilitation Directors themselves, and that he had even developed some machines and methods to hasten rehabilitation. There was no doubt that Benjamin would receive his Assignment to that field and would probably be permitted to bypass most of the training.",
                                        "Jonas was impressed by the things Benjamin had achieved. He knew him, of course, since they had always been groupmates, but they had never talked about the boy's accomplishments because such a conversation would have been awkward for Benjamin.
There was never any comfortable way to mention or discuss one's successes without breaking the rule against bragging, even if one didn't mean to. It was a minor rule, rather like rudeness, punishable only by gentle chastisement. But still. Better to steer clear of an occasion governed by a rule which would be so easy to break.",
                                        "The area of dwellings behind him, Jonas rode past the community structures, hoping to spot Asher's bicycle parked beside one of the small factories or office buildings. He passed the Childcare Center where Lily stayed after school, and the play areas surrounding it.
He rode through the Central Plaza and the large Auditorium where public meetings were held.",
                                        "Jonas slowed and looked at the nametags on the bicycles lined up outside the Nurturing Center. Then he checked those outside Food Distribution; it was always fun to help with the deliveries, and he hoped he would find his friend there so that they could go together on the daily rounds, carrying the cartons of supplies into the dwellings of the community. But he finally found Asher's bicycle-- leaning, as usual, instead of upright in its port, as it should have been--at the House of the Old.",
                                        "There was only one other child's bicycle there, that of a female Eleven named Fiona. Jonas liked Fiona. She was a good student, quiet and polite, but she had a sense of fun as well, and it didn't surprise him that she was working with Asher today. He parked his bicycle neatly in the port beside theirs and entered the building.",
                                        "\"Hello, Jonas,\" the attendant at the front desk said. She handed him the sign-up sheet and stamped her own official seal beside his signature. All of his volunteer hours would be carefully tabulated at the Hall of Open Records. Once, long ago, it was whispered among the children, an Eleven had arrived at the Ceremony of Twelve only to hear a public announcement that he had not completed the required number of volunteer hours and would not, therefore, be given his Assignment. He had been permitted an additional month in which to complete the hours, and then given his Assignment privately, with no applause, no celebration: a disgrace that had clouded his entire future.",
                                        "\"It's good to have some volunteers here today,\" the attendant told him. \"We celebrated a release this morning, and that always throws the schedule off a little, so things get backed up.\" She looked at a printed sheet. \"Let's see. Asher and Fiona are helping in the bathing room. Why don't you join them there? You know where it is, don't you?\"",
                                        "Jonas nodded, thanked her, and walked down the long hallway. He glanced into the rooms on either side. The Old were sitting quietly, some visiting and talking with one another, others doing handwork and simple crafts. A few were asleep. Each room was comfortably furnished, the floors covered with thick carpeting. It was a serene and slow-paced place, unlike the busy centers of manufacture and distribution where the daily work of the community occurred.",
                                        "Jonas was glad that he had, over the years, chosen to do his hours in a variety of places so that he could experience the differences. He realized, though, that not focusing on one area meant he was left with not the slightest idea-not even a guess of what his Assignment would be.",
                                        "He laughed softly. Thinking about the Ceremony again, Jonas? he teased himself. But he suspected that with the date so near, probably all of his friends were, too.",
                                        "He passed a Caretaker walking slowly with one of the Old in the hall. \"Hello, Jonas,\" the young uniformed man said, smiling pleasantly. The woman beside him, whose arm he held, was hunched over as she shuffled along in her soft slippers. She looked toward Jonas and smiled, but her dark eyes were clouded and blank. He realized she was blind.",
                                        "He entered the bathing room with its warm moist air and scent of cleansing lotions. He removed his tunic, hung it carefully on a wall hook, and put on the volunteer's smock that was folded on a shelf.",
                                        "\"Hi, Jonas!\" Asher called from the corner where he was kneeling beside a tub. Jonas saw Fiona nearby, at a different tub. She looked up and smiled at him, but she was busy, gently washing a man who lay in the warm water.",
                                        "Jonas greeted them and the caretaking attendants at work nearby. Then he went to the row of padded lounging chairs where others of the Old were waiting. He had worked here before; he knew what to do.",
                                        "\"Your turn, Larissa,\" he said, reading the nametag on the woman's robe. \"I'll just start the water and then help you up.\" He pressed the button on a nearby empty tub and watched as the warm water flowed in through the many small openings on the sides. The tub would be filled in a minute and the water flow would stop automatically.",
                                        "He helped the woman from the chair, led her to the tub, removed her robe, and steadied her with his hand on her arm as she stepped in and lowered herself. She leaned back and sighed with pleasure, her head on a soft cushioned headrest.",
                                        "\"Comfortable?\" he asked, and she nodded, her eyes closed. Jonas squeezed cleansing lotion onto the clean sponge at the edge of the tub and began to wash her frail body.",
                                        "Last night he had watched as his father bathed the newchild. This was much the same: the fragile skin, the soothing water, the gentle motion of his hand, slippery with soap. The relaxed, peaceful smile on the woman's face reminded him of Gabriel being bathed.",
                                        "And the nakedness, too. It was against the rules for children or adults to look at another's nakedness; but the rule did not apply to newchildren or the Old. Jonas was glad. It was a nuisance to keep oneself covered while changing for games, and the required apology if one had by mistake glimpsed another's body was always awkward. He couldn't see why it was necessary. He liked the feeling of safety here in this warm and quiet room; he liked the expression of trust on the woman's face as she lay in the water unprotected, exposed, and free.",
                                        "From the corner of his eye he could see his friend Fiona help the old man from the tub and tenderly pat his thin, naked body dry with an absorbant cloth. She helped him into his robe.",
                                        "Jonas thought Larissa had drifted into sleep, as the Old often did, and he was careful to keep his motions steady and gentle so he wouldn't wake her. He was surprised when she spoke, her eyes still closed. \"This morning we celebrated the release of Roberto,\" she told him. \"It was wonderful.\"",
                                        "\"I knew Roberto!\" Jonas said. \"I helped with his feeding the last time I was here, just a few weeks ago. He was a very interesting man.\"",
                                        "Larissa opened her eyes happily. \"They told his whole life before they released him,\" she said. \"They always do. But to be honest,\" she whispered with a mischievous look, \"some of the tellings are a little boring. I've even seen some of the Old fall asleep during tellings- when they released Edna recently. Did you know Edna?\"",
                                        "Jonas shook his head. He couldn't recall anyone named Edna.",
                                        "\"Well, they tried to make her life sound meaningful. And of course,\" she added primly, \"all lives are meaningful, I don't mean that they aren't. But Edna. My goodness. She was a Birthmother, and then she worked in Food Production for years, until she came here. She never even had a family unit.\"",
                                        "Larissa lifted her head and looked around to make sure no one else was listening. Then she confided, \"I don't think Edna was very smart.\"",
                                        "Jonas laughed. He rinsed her left arm, laid it back into the water, and began to wash her feet. She murmured with pleasure as he massaged her feet with the sponge.",
                                        "\"But Roberto's life was wonderful,\" Larissa went on, after a moment. \"He had been an Instructor of Elevens --you know how important that is--and he'd been on the Planning Committee. And-- goodness, I don't know how he found the time--he also raised two very successful children,, and he was also the one who did the landscaping design for the Central Plaza. He didn't do the actual labor, of course.\"",
                                        "\"Now your back. Lean forward and I'll help you sit up.\" Jonas put his arm around her and supported her as she sat. He squeezed the sponge against her back and began to rub her sharp-boned shoulders. \"Tell me about the celebration.\"",
                                        "\"Well, there was the telling of his life. That is always first. Then the toast. We all raised our glasses and cheered. We chanted the anthem. He made a lovely goodbye speech. And several of us made little speeches wishing him well. I didn't, though. I've never been fond of public speaking.\"",
                                        "\"He was thrilled. You should have seen the look on his face when they let him go.\"",
                                        "Jonas slowed the strokes of his hand on her back thoughtfully. \"Larissa,\" he asked, \"what happens when they make the actual release? Where exactly did Roberto go?\"",
                                        "She lifted her bare wet shoulders in a small shrug. \"I don't know. I don't think anybody does, except the committee. He just bowed to all of us and then walked, like they all do, through the special door in the Releasing Room. But you should have seen his look. Pure happiness, I'd call it.\"",
                                        "Jonas grinned. \"I wish I'd been there to see it.\"",
                                        "Larissa frowned. \"I don't know why they don't let children come. Not enough room, I guess. They should enlarge the Releasing Room.\"",
                                        "\"We'll have to suggest that to the committee. Maybe they'd study it,\" Jonas said slyly, and Larissa chortled with laughter.",
                                        "\"Right!\" she hooted, and Jonas helped her from the tub.")) 
```

##Chapter 5
```{r ch5}
annotate_ch5 <- cnlp_annotate(input = c("Usually, at the morning ritual when the family members told their dreams, Jonas didn't contribute much. He rarely dreamed. Sometimes he awoke with a feeling of fragments afloat in his sleep, but he couldn't seem to grasp them and put them together into something worthy of telling at the ritual.",
                                        "But this morning was different. He had dreamed very vividly the night before.",
                                        "His mind wandered while Lily, as usual, recounted a lengthy dream, this one a frightening one in which she had, against the rules, been riding her mother's bicycle and been caught by the Security Guards.",
                                        "They all listened carefully and discussed with Lily the warning that the dream had given.",
                                        "\"Thank you for your dream, Lily.\" Jonas said the standard phrase automatically, and tried to pay better attention while his mother told of a dream fragment, a disquieting scene where she had been chastised for a rule infraction she didn't understand. Together they agreed that it probably resulted from her feelings when she had reluctantly dealt punishment to the citizen who had broken the major rules a second time.",
                                        "Father said that he had had no dreams.",
                                        "\"Gabe?\" Father asked, looking down at the basket where the newchild lay gurgling after his feeding, ready to be taken back to the Nurturing Center for the day.",
                                        "They all laughed. Dream-telling began with Threes. If newchildren dreamed, no one knew.",
                                        "\"Jonas?\" Mother asked. They always asked, though they knew how rarely Jonas had a dream to tell.",
                                        "\"I did dream last night,\" Jonas told them. He shifted in his chair, frowning.",
                                        "\"Good,\" Father said. \"Tell us.\"",
                                        "\"The details aren't clear, really,\" Jonas explained, trying to recreate the odd dream in his mind. \"I think I was in the bathing room at the House of the Old.\"",
                                        "\"That's where you were yesterday,\" Father pointed out.",
                                        "Jonas nodded. \"But it wasn't really the same. There was a tub, in the dream. But only one. And the real bathing room has rows and rows of them. But the room in the dream was warm and damp. And I had taken off my tunic, but hadn't put on the smock, so my chest was bare. I was perspiring, because it was so warm. And Fiona was there, the way she was yesterday.\"",
                                        "\"Asher, too?\" Mother asked.",
                                        "Jonas shook his head. \"No. It was only me and Fiona, alone in the room, standing beside the tub. She was laughing. But I wasn't. I was almost a little angry at her, in the dream, because she wasn't taking me seriously.\"",
                                        "\"Seriously about what?\" Lily asked.",
                                        "Jonas looked at his plate. For some reason that he didn't understand, he felt slightly embarrassed. \"I think I was trying to convince her that she should get into the tub of water.\"",
                                        "He paused. He knew he had to tell it all, that it was not only all right but necessary to tell all of a dream. So he forced himself to relate the part that made him uneasy.",
                                        "\"I wanted her to take off her clothes and get into the tub,\" he explained quickly. \"I wanted to bathe her. I had the sponge in my hand. But she wouldn't. She kept laughing and saying no.\"",
                                        "He looked up at his parents. \"That's all,\" he said.",
                                        "\"Can you describe the strongest feeling in your dream, son?\" Father asked.",
                                        "Jonas thought about it. The details were murky and vague. But the feelings were clear, and flooded him again now as he thought. \"The wanting,\" he said. \"I knew that she wouldn't. And I think I knew that she shouldn't. But I wanted it so terribly. I could feel the wanting all through me.\"",
                                        "\"Thank you for your dream, Jonas,\" Mother said after a moment. She glanced at Father.",
                                        "\"Lily,\" Father said, \"it's time to leave for school. Would you walk beside me this morning and keep an eye on the newchild's basket? We want to be certain he doesn't wiggle himself loose.\"",
                                        "Jonas began to rise to collect his schoolbooks. He thought it surprising that they hadn't talked about his dream at length before the thank you. Perhaps they found it as confusing as he had.",
                                        "\"Wait, Jonas,\" Mother said gently. \"I'll write an apology to your instructor so that you won't have to speak one for being late.\"",
                                        "He sank back down into his chair, puzzled. He waved to Father and Lily as they left the dwelling, carrying Gabe in his basket. He watched while Mother tidied the remains of the morning meal and placed the tray by the front door for the Collection Crew.",
                                        "Finally she sat down beside him at the table. \"Jonas,\" she said with a smile, \"the feeling you described as the wanting? It was your first Stirrings. Father and I have been expecting it to happen to you. It happens to everyone. It happened to Father when he was your age. And it happened to me. It will happen someday to Lily.\"",
                                        "\"And very often,\" Mother added, \"it begins with a dream.\"",
                                        "Stirrings. He had heard the word before. He remembered that there was a reference to the Stirrings in the Book of Rules, though he didn't remember what it said. And now and then the Speaker mentioned it. ATTENTION. A REMINDER THAT STIRRINGS MUST BE REPORTED IN ORDER FOR TREATMENT TO TAKE PLACE.",
                                        "He had always ignored that announcement because he didn't understand it and it had never seemed to apply to him in any way. He ignored, as most citizens did, many of the commands and reminders read by the Speaker.",
                                        "\"Do I have to report it?\" he asked his mother.",
                                        "She laughed. \"You did, in the dream-telling. That's enough.\"",
                                        "\"But what about the treatment? The Speaker says that treatment must take place.\" Jonas felt miserable. Just when the Ceremony was about to happen, his Ceremony of Twelve, would he have to go away someplace for treatment? Just because of a stupid dream?",
                                        "But his mother laughed again in a reassuring, affectionate way. \"No, no,\" she said. \"It's just the pills. You're ready for the pills, that's all. That's the treatment for Stirrings.\"",
                                        "Jonas brightened. He knew about the pills. His parents both took them each morning. And some of his friends did, he knew. Once he had been heading off to school with Asher, both of them on their bikes, when Asher's father had called from their dwelling doorway, \"You forgot your pill, Asher!\" Asher had groaned good-naturedly, turned his bike, and ridden back while Jonas waited.",
                                        "It was the sort of thing one didn't ask a friend about because it might have fallen into that uncomfortable category of \"being different.\" Asher took a pill each morning; Jonas did not. Always better, less rude, to talk about things that were the same.",
                                        "Now he swallowed the small pill that his mother handed him.",
                                        "\"That's all?\" he asked.",
                                        "\"That's all,\" she replied, returning the bottle to the cupboard. \"But you mustn't forget. I'll remind you for the first weeks, but then you must do it on your own. If you forget, the Stirrings will come back. The dreams of Stirrings will come back. Sometimes the dosage must be adjusted.\"",
                                        "\"Asher takes them,\" Jonas confided.",
                                        "His mother nodded, unsurprised. \"Many of your groupmates probably do. The males, at least. And they all will, soon. Females too.\"",
                                        "\"How long will I have to take them?\"",
                                        "\"Until you enter the House of the Old,\" she explained. \"All of your adult life. But it becomes routine; after a while you won't even pay much attention to it.\"",
                                        "She looked at her watch. \"If you leave right now, you won't even be late for school. Hurry along.\"",
                                        "\"And thank you again, Jonas,\" she added, as he went to the door, \"for your dream.\"",
                                        "Pedaling rapidly down the path, Jonas felt oddly proud to have joined those who took the pills. For a moment, though, he remembered the dream again. The dream had felt pleasurable. Though the feelings were confused, he thought that he had liked the feelings that his mother had called Stirrings. He remembered that upon waking, he had wanted to feel the Stirrings again.",
                                        "Then, in the same way that his own dwelling slipped away behind him as he rounded a corner on his bicycle, the dream slipped away from his thoughts. Very briefly, a little guiltily, he tried to grasp it back. But the feelings had disappeared. The Stirrings were gone."))

```

##Chapter 6
```{r ch6}
annotate_ch5 <- cnlp_annotate(input = c("\"Lily, please hold still,\" Mother said again.",
                                        "Lily, standing in front of her, fidgeted impatiently. \"I can tie them myself,\" she complained. \"I always have.\"",
                                        "\"I know that,\" Mother replied, straightening the hair ribbons on the little girl's braids. \"But I also know that they constantly come loose and more often than not, they're dangling down your back by afternoon. Today, at least, we want them to be neatly tied and to stay neatly tied.\"",
                                        "\"I don't like hair ribbons. I'm glad I only have to wear them one more year,\" Lily said irritably. \"Next year I get my bicycle, too,\" she added more cheerfully.",
                                        "\"There are good things each year,\" Jonas reminded her. \"This year you get to start your volunteer hours. And remember last year, when you became a Seven, you were so happy to get your front- buttoned jacket?\"",
                                        "The little girl nodded and looked down at herself, at the jacket with its row of large buttons that designated her as a Seven. Fours, Fives, and Sixes all wore jackets that fastened down the back so that they would have to help each other dress and would learn interdependence.",
                                        "The front-buttoned jacket was the first sign of independence, the first very visible symbol of growing up. The bicycle, at Nine, would be the powerful emblem of moving gradually out into the community, away from the protective family unit.",
                                        "Lily grinned and wriggled away from her mother. \"And this year you get your Assignment,\" she said to Jonas in an excited voice. \"I hope you get Pilot. And that you take me flying!\"",
                                        "\"Sure I will,\" said Jonas. \"And I'll get a special little parachute that just fits you, and I'll take you up to, oh, maybe twenty thousand feet, and open the door, and--\"", 
                                        "\"Jonas,\" Mother warned.",
                                        "\"I was only joking,\" Jonas groaned. \"I don't want Pilot, anyway. If I get Pilot I'll put in an appeal.\"",
                                        "\"Come on,\" Mother said. She gave Lily's ribbons a final tug. \"Jonas? Are you ready? Did you take your pill? I want to get a good seat in the Auditorium.\" She prodded Lily to the front door and Jonas followed.",
                                        "It was a short ride to the Auditorium, Lily waving to her friends from her seat on the back of Mother's bicycle. Jonas stowed his bicycle beside Mother's and made his way through the throng to find his group.",
                                        "The entire community attended the Ceremony each year. For the parents, it meant two days holiday from work; they sat together in the huge hall. Children sat with their groups until they went, one by one, to the stage.",
                                        "Father, though, would not join Mother in the audience right away. For the earliest ceremony, the Naming, the Nurturers brought the newchildren to the stage. Jonas, from his place in the balcony with the Elevens, searched the Auditorium for a glimpse of Father. It wasn't at all hard to spot the Nurturers' section at the front; coming from it were the wails and howls of the newchildren who sat squirming on the Nurturers' laps. At every other public ceremony, the audience was silent and attentive. But once a year, they all smiled indulgently at the commotion from the little ones waiting to receive their names and families.",
                                        "Jonas finally caught his father's eye and waved. Father grinned and waved back, then held up the hand of the newchild on his lap, making it wave, too.",
                                        "It wasn't Gabriel. Gabe was back at the Nurturing Center today, being cared for by the night crew. He had been given an unusual and special reprieve from the committee, and granted an additional year of nurturing before his Naming and Placement. Father had gone before the committee with a plea on behalf of Gabriel, who had not yet gained the weight appropriate to his days of life nor begun to sleep soundly enough at night to be placed with his family unit. Normally such a newchild would be labeled Inadequate and released from the community.",
                                        "Instead, as a result of Father's plea, Gabriel had been labeled Uncertain and given the additional year. He would continue to be nurtured at the Center and would spend his nights with Jonas's family unit. Each family member, including Lily, had been required to sign a pledge that they would not become attached to this little temporary guest, and that they would relinquish him without protest At least, Jonas thought, after Gabriel was placed next year, they would still see him often because he would be part of the community. If he were released, they would not see him again. Ever. Those who were released--even as newchildren--were sent Elsewhere and never returned to the community.",
                                        "Father had not had to release a single newchild this year, so Gabriel would have represented a real failure and sadness. Even Jonas, though he didn't hover over the little one the way Lily and his father did, was glad that Gabe had not been released.",
                                        "The first Ceremony began right on time, and Jonas watched as one after another each newchild was given a name and handed by the Nurturers to its new family unit. For some, it was a first child. But many came to the stage accompanied by another child beaming with pride to receive a little brother or sister, the way Jonas had when he was about to be a Five.",
                                        "Asher poked Jonas's arm. \"Remember when we got Phillipa?\" he asked in a loud whisper. Jonas nodded. It had only been last year. Asher's parents had waited quite a long time before applying for a second child. Maybe, Jonas suspected, they had been so exhausted by Asher's lively foolishness that they had needed a little time.",
                                        "Two of their group, Fiona and another female named Thea, were missing temporarily, waiting with their parents to receive newchildren. But it was rare that there was such an age gap between children in a family unit.",
                                        "When her family's ceremony was completed, Fiona took the seat that had been saved for her in the row ahead of Asher and Jonas. She turned and whispered to them, \"He's cute. But I don't like his name very much.\" She made a face and giggled. Fiona's new brother had been named Bruno. It wasn't a great name, Jonas thought, like--well, like Gabriel, for example. But it was okay.",
                                        "The audience applause, which was enthusiastic at each Naming, rose in an exuberant swell when one parental pair, glowing with pride, took a male newchild and heard him named Caleb. This new Caleb was a replacement child. The couple had lost their first Caleb, a cheerful little Four. Loss of a child was very, very rare. The community was extraordinarily safe, each citizen watchful and protective of all children. But somehow the first little Caleb had wandered away unnoticed, and had fallen into the river. The entire community had performed the Ceremony of Loss together, murmuring the name Caleb throughout an entire day, less and less frequently, softer in volume, as the long and somber day went on, so that the little Four seemed to fade away gradually from everyone's consciousness.",
                                        "Now, at this special Naming, the community performed the brief Murmur-of-Replacement Ceremony, repeating the name for the first time since the loss: softly and slowly at first, then faster and with greater volume, as the couple stood on the stage with the newchild sleeping in the mother's arms. It was as if the first Caleb were returning.",
                                        "Another newchild was given the name Roberto, and Jonas remembered that Roberto the Old had been released only last week. But there was no Murmur-of-Replacement Ceremony for the new little Roberto. Release was not the same as Loss.",
                                        "He sat politely through the ceremonies of Two and Three and Four, increasingly bored as he was each year. Then a break for midday meal--served outdoorsand back again to the seats, for the Fives, Sixes, Sevens, and finally, last of the first day's ceremonies, the Eights.",
                                        "Jonas watched and cheered as Lily marched proudly to the stage, became an Eight and received the identifying jacket that she would wear this year, this one with smaller buttons and, for the first time, pockets, indicating that she was mature enough now to keep track of her own small belongings. She stood solemnly listening to the speech of firm instructions on the responsibilities of Eight and doing volunteer hours for the first time. But Jonas could see that Lily, though she seemed attentive, was looking longingly at the row of gleaming bicycles, which would be presented tomorrow morning to the Nines.",
                                        "Next year, Lily-billy, Jonas thought.",
                                        "It was an exhausting day, and even Gabriel, retrieved in his basket from the Nurturing Center, slept soundly that night.",
                                        "Finally it was the morning of the Ceremony of Twelve.",
                                        "Now Father sat beside Mother in the audience. Jonas could see them applauding dutifully as the Nines, one by one, wheeled their new bicycles, each with its gleaming nametag attached to the back, from the stage. He knew that his parents cringed a little, as he did, when Fritz, who lived in the dwelling next door to theirs, received his bike and almost immediately bumped into the podium with it. Fritz was a very awkward child who had been summoned for chastisement again and again. His transgressions were small ones, always: shoes on the wrong feet, schoolwork misplaced, failure to study adequately for a quiz. But each such error reflected negatively on his parents' guidance and infringed on the community's sense of order and success. Jonas and his family had not been looking forward to Fritz's bicycle, which they realized would probably too often be dropped on the front walk instead of wheeled neatly into its port.",
                                        "Finally the Nines were all resettled in their seats, each having wheeled a bicycle outside where it would be waiting for its owner at the end of the day. Everyone always chuckled and made small jokes when the Nines rode home for the first time. \"Want me to show you how to ride?\" older friends would call. \"I know you've never been on a bike before!\" But invariably the grinning Nines, who in technical violation of the rule had been practicing secretly for weeks, would mount and ride off in perfect balance, training wheels never touching the ground.",
                                        "Then the Tens. Jonas never found the Ceremony of Ten particularly interesting- only time-consuming, as each child's hair was snipped neatly into its distinguishing cut: females lost their braids at Ten, and males, too, relinquished their long childish hair and took on the more manly short style which exposed their ears.",
                                        "Laborers moved quickly to the stage with brooms and swept away the mounds of discarded hair. Jonas could see the parents of the new Tens stir and murmur, and he knew that this evening, in many dwellings, they would be snipping and straightening the hastily done haircuts, trimming them into a neater line.",
                                        "Elevens. It seemed a short time ago that Jonas had undergone the Ceremony of Eleven, but he remembered that it was not one of the more interesting ones. By Eleven, one was only waiting to be Twelve. It was simply a marking of time with no meaningful changes. There was new clothing: different undergarments for the females, whose bodies were beginning to change; and longer trousers for the males, with a specially shaped pocket for the small calculator that they would use this year in school; but those were simply presented in wrapped packages without an accompanying speech.",
                                        "Break for midday meal. Jonas realized he was hungry. He and his groupmates congregated by the tables in front of the Auditorium and took their packaged food. Yesterday there had been merriment at lunch, a lot of teasing and energy. But today the group stood anxiously, separate from the other children. Jonas watched the new Nines gravitate toward their waiting bicycles, each one admiring his or her nametag. He saw the Tens stroking their new shortened hair, the females shaking their heads to feel the unaccustomed lightness without the heavy braids they had worn so long.",
                                        "\"I heard about a guy who was absolutely certain he was going to be assigned Engineer,\" Asher muttered as they ate, \"and instead they gave him Sanitation Laborer. He went out the next day, jumped into the river, swam across, and joined the next community he came to. Nobody ever saw him again.\"",
                                        "Jonas laughed. \"Somebody made that story up, Ash,\" he said. \"My father said he heard that story when he was a Twelve.\"",
                                        "But Asher wasn't reassured. He was eyeing the river where it was visible behind the Auditorium. \"I can't even swim very well,\" he said.",
                                        "\"My swimming instructor said that I don't have the right boyishness or something.\"",
                                        "\"Buoyancy,\" Jonas corrected him. \"Whatever. I don't have it. I sink.\"",
                                        "\"Anyway,\" Jonas pointed out, \"have you ever once known of anyone--I mean really known for sure, Asher, not just heard a story about it--who joined another community?\"",
                                        "\"No,\" Asher admitted reluctantly. \"But you can. It says so in the rules. If you don't fit in, you can apply for Elsewhere and be released. My mother says that once, about ten years ago, someone applied and was gone the next day.\" Then he chuckled. \"She told me that because I was driving her crazy. She threatened to apply for Elsewhere.\"",
                                        "\"She was joking.\"",
                                        "\"I know. But it was true, what she said, that someone did that once. She said that it was really true. Here today and gone tomorrow.\"",
                                        "\"Never seen again. Not even a Ceremony of Release.\"",
                                        "Jonas shrugged. It didn't worry him. How could someone not fit in? The community was so meticulously ordered, the choices so carefully made.",
                                        "Even the Matching of Spouses was given such weighty consideration that sometimes an adult who applied to receive a spouse waited months or even years before a Match was approved and announced. All of the factors--disposition, energy level, intelligence, and interests--had to correspond and to interact perfectly. Jonas's mother, for example, had higher intelligence than his father; but his father had a calmer disposition. They balanced each other. Their Match, which like all Matches had been monitored by the Committee of Elders for three years before they could apply for children, had always been a successful one.",
                                        "Like the Matching of Spouses and the Naming and Placement of newchildren, the Assignments were scrupulously thought through by the Committee of Elders.",
                                        "He was certain that his Assignment, whatever it was to be, and Asher's too, would be the right one for them. He only wished that the midday break would conclude, that the audience would reenter the Auditorium, and the suspense would end.",
                                        "As if in answer to his unspoken wish, the signal came and the crowd began to move toward the doors."))

```

##Chapter 7
```{r ch7}
annotate_ch7 <- cnlp_annotate(input = c("Now Jonas's group had taken a new place in the Auditorium, trading with the new Elevens, so that they sat in the very front, immediately before the stage.",
                                        "They were arranged by their original numbers, the numbers they had been given at birth. The numbers were rarely used after the Naming. But each child knew his number, of course. Sometimes parents used them in irritation at a child's misbehavior, indicating that mischief made one unworthy of a name. Jonas always chuckled when he heard a parent, exasperated, call sharply to a whining toddler, \"That's enough, Twenty-three!\"",
                                        "Jonas was Nineteen. He had been the nineteenth newchild born his year. It had meant that at his Naming, he had been already standing and bright-eyed, soon to walk and talk. It had given him a slight advantage the first year or two, a little more maturity than many of his groupmates who had been born in the later months of that year. But it evened out, as it always did, by Three.",
                                        "After Three, the children progressed at much the same level, though by their first number one could always tell who was a few months older than others in his group. Technically, Jonas's full number was Eleven-nineteen, since there were other Nineteens, of course, in each age group. And today, now that the new Elevens had been advanced this morning, there were two Eleven-nineteens. At the midday break he had exchanged smiles with the new one, a shy female named Harriet.",
                                        "But the duplication was only for these few hours. Very soon he would not be an Eleven but a Twelve, and age would no longer matter. He would be an adult, like his parents, though a new one and untrained still.",
                                        "Asher was Four, and sat now in the row ahead of Jonas. He would receive his Assignment fourth.",
                                        "Fiona, Eighteen, was on his left; on his other side sat Twenty, a male named Pierre whom Jonas didn't like much. Pierre was very serious, not much fun, and a worrier and tattletale, too. \"Have you checked the rules, Jonas?\" Pierre was always whispering solemnly. \"I'm not sure that's within the rules.\" Usually it was some foolish thing that no one cared about--opening his tunic if it was a day with a breeze; taking a brief try on a friend's bicycle, just to experience the different feel of it.",
                                        "The initial speech at the Ceremony of Twelve was made by the Chief Elder, the leader of the community who was elected every ten years. The speech was much the same each year: recollection of the time of childhood and the period of preparation, the coming responsibilities of adult life, the profound importance of Assignment, the seriousness of training to come.",
                                        "Then the Chief Elder moved ahead in her speech. \"This is the time,\" she began, looking directly at them, \"when we acknowledge differences. You Elevens have spent all your years till now learning to fit in, to standardize your behavior, to curb any impulse that might set you apart from the group.\"",
                                        "\"But today we honor your differences. They have determined your futures.\"",
                                        "She began to describe this year's group and its variety of personalities, though she singled no one out by name. She mentioned that there was one who had singular skills at caretaking, another who loved newchildren, one with unusual scientific aptitude, and a fourth for whom physical labor was an obvious pleasure. Jonas shifted in his seat, trying to recognize each reference as one of his groupmates. The caretaking skills were no doubt those of Fiona, on his left; he remembered noticing the tenderness with which she had bathed the Old. Probably the one with scientific aptitude was Benjamin, the male who had devised new, important equipment for the Rehabilitation Center. He heard nothing that he recognized as himself, Jonas. Finally the Chief Elder paid tribute to the hard work of her committee, which had performed the observations so meticulously all year. The Committee of Elders stood and was acknowledged by applause. Jonas noticed Asher yawn slightly, covering his mouth politely with his hand.",
                                        "Then, at last, the Chief Elder called number One to the stage, and the Assignments began.",
                                        "Each announcement was lengthy, accompanied by a speech directed at the new Twelve. Jonas tried to pay attention as One, smiling happily, received her Assignment as Fish Hatchery Attendant along with words of praise for her childhood spent doing many volunteer hours there, and her obvious interest in the important process of providing nourishment for the community.",
                                        "Number One--her name was Madeline--returned, finally, amidst applause, to her seat, wearing the new badge that designated her Fish Hatchery Attendant. Jonas was certainly glad that that Assignment was taken; he wouldn't have wanted it. But he gave Madeline a smile of congratulation.",
                                        "When Two, a female named Inger, received her Assignment as Birthmother, Jonas remembered that his mother had called it a job without honor. But he thought that the Committee had chosen well. Inger was a nice girl though somewhat lazy, and her body was strong. She would enjoy the three years of being pampered that would follow her brief training; she would give birth easily and well; and the task of Laborer that would follow would use her strength, keep her healthy, and impose self-discipline. Inger was smiling when she resumed her seat. Birthmother was an important job, if lacking in prestige.",
                                        "Jonas noticed that Asher looked nervous. He kept turning his head and glancing back at Jonas until the group leader had to give him a silent chastisement, a motion to sit still and face forward.",
                                        "Three, Isaac, was given an Assignment as Instructor of Sixes, which obviously pleased him and was well deserved. Now there were three Assignments gone, none of them ones that Jonas would have liked not that he could have been a Birthmother, anyway, he realized with amusement. He tried to sort through the list in his mind, the possible Assignments that remained. But there were so many he gave it up; and anyway, now it was Asher's turn. He paid strict attention as his friend went to the stage and stood self- consciously beside the Chief Elder.",
                                        "\"All of us in the community know and enjoy Asher,\" the Chief Elder began. Asher grinned and scratched one leg with the other foot. The audience chuckled softly.",
                                        "\"When the committee began to consider Asher's Assignment,\" she went on, \"there were some possibilities that were immediately discarded. Some that would clearly not have been right for Asher.\"",
                                          "\For example,\" she said, smiling, \"we did not consider for an instant designating Asher an Instructor of Threes.\"",
                                        "The audience howled with laughter. Asher laughed, too, looking sheepish but pleased at the special attention. The Instructors of Threes were in charge of the acquisition of correct language.",
                                        "\In fact,\" the Chief Elder continued, chuckling a little herself, \"we even gave a little thought to some retroactive chastisement for the one who had been Asher's Instructor of Threes so long ago. At the meeting where Asher was discussed, we retold many of the stories that we all remembered from his days of language acquisition.\"",
                                        "\"Especially,\" she said, chuckling, \"the difference between snack and smack. Remember, Asher?\"",
                                        "Asher nodded ruefully, and the audience laughed aloud. Jonas did, too. He remembered, though he had been only a Three at the time himself.",
                                        "The punishment used for small children was a regulated system of smacks with the discipline wand: a thin, flexible weapon that stung painfully when it was wielded. The Childcare specialists were trained very carefully in the discipline methods: a quick smack across the hands for a bit of minor misbehavior; three sharper smacks on the bare legs for a second offense.",
                                        "Poor Asher, who always talked too fast and mixed up words, even as a toddler. As a Three, eager for his juice and crackers at snacktime, he one day said \"smack\" instead of \"snack\" as he stood waiting in line for the morning treat.",
                                        "Jonas remembered it clearly. He could still see little Asher, wiggling with impatience in the line. He remembered the cheerful voice call out, \"I want my smack!\"",
                                        "The other Threes, including Jonas, had laughed nervously. \"Snack!\" they corrected. \"You meant snack, Asher!\" But the mistake had been made. And precision of language was one of the most important tasks of small children. Asher had asked for a smack.",
                                        "The discipline wand, in the hand of the Childcare worker, whistled as it came down across Asher's hands. Asher whimpered, cringed, and corrected himself instantly. \"Snack,\" he whispered.",
                                        "But the next morning he had done it again. And again the following week. He couldn't seem to stop, though for each lapse the discipline wand came again, escalating to a series of painful lashes that left marks on Asher's legs. Eventually, for a period of time, Asher stopped talking altogether, when he was a Three.",
                                        "\"For a while,\" the Chief Elder said, relating the story, \"we had a silent Asher! But he learned.\"",
                                        "She turned to him with a smile. \"When he began to talk again, it was with greater precision. And now his lapses are very few. His corrections and apologies are very prompt. And his good humor is unfailing.\" The audience murmured in agreement. Asher's cheerful disposition was well-known throughout the community.",
                                        "\"Asher.\" She lifted her voice to make the official announcement. \"We have given you the Assignment of Assistant Director of Recreation.\"",
                                        "She clipped on his new badge as he stood beside her, beaming. Then he turned and left the stage as the audience cheered. When he had taken his seat again, the Chief Elder looked down at him and said the words that she had said now four times, and would say to each new Twelve. Somehow she gave it special meaning for each of them.",
                                        "\"Asher,\" she said, \"thank you for your childhood.\"",
                                        "The Assignments continued, and Jonas watched and listened, relieved now by the wonderful Assignment his best friend had been given. But he was more and more apprehensive as his own approached. Now the new Twelves in the row ahead had all received their badges. They were fingering them as they sat, and Jonas knew that each one was thinking about the training that lay ahead. For some--one studious male had been selected as Doctor, a female as Engineer, and another for Law and Justice--it would be years of hard work and study. Others, like Laborers and Birthmothers, would have a much shorter training period.",
                                        "Eighteen, Fiona, on his left, was called. Jonas knew she must be nervous, but Fiona was a calm female. She had been sitting quietly, serenely, throughout the Ceremony.",
                                        "Even the applause, though enthusiastic, seemed serene when Fiona was given the important Assignment of Caretaker of the Old. It was perfect for such a sensitive, gentle girl, and her smile was satisfied and pleased when she took her seat beside him again.",
                                        "Jonas prepared himself to walk to the stage when the applause ended and the Chief Elder picked up the next folder and looked down to the group to call forward the next new Twelve. He was calm now that his turn had come. He took a deep breath and smoothed his hair with his hand.",
                                        "\"Twenty,\" he heard her voice say clearly. \"Pierre.\"",
                                        "She skipped me, Jonas thought, stunned. Had he heard wrong? No. There was a sudden hush in the crowd, and he knew that the entire community realized that the Chief Elder had moved from Eighteen to Twenty, leaving a gap. On his right, Pierre, with a startled look, rose from his seat and moved to the stage.",
                                        "A mistake. She made a mistake. But Jonas knew, even as he had the thought, that she hadn't. The Chief Elder made no mistakes. Not at the Ceremony of Twelve.",
                                        "He felt dizzy, and couldn't focus his attention. He didn't hear what Assignment Pierre received, and was only dimly aware of the applause as the boy returned, wearing his new badge. Then: Twenty-one. Twenty-two.",
                                        "The numbers continued in order. Jonas sat, dazed, as they moved into the Thirties and then the Forties, nearing the end. Each time, at each announcement, his heart jumped for a moment, and he thought wild thoughts. Perhaps now she would call his name. Could he have forgotten his own number? No. He had always been Nineteen. He was sitting in the seat marked Nineteen.",
                                        "But she had skipped him. He saw the others in his group glance at him, embarrassed, and then avert their eyes quickly. He saw a worried look on the face of his group leader. He hunched his shoulders and tried to make himself smaller in the seat. He wanted to disappear, to fade away, not to exist. He didn't dare to turn and find his parents in the crowd. He couldn't bear to see their faces darkened with shame.",
                                        "Jonas bowed his head and searched through his mind. What had he done wrong?"))

```

##Chapter 8
```{r ch8}
annotate_ch8 <- cnlp_annotate(input = c("The audience was clearly ill at ease. They applauded at the final Assignment; but the applause was piecemeal, no longer a crescendo of united enthusiasm. There were murmurs of confusion.",
                                        "Jonas moved his hands together, clapping, but it was an automatic, meaningless gesture that he wasn't even aware of. His mind had shut out all of the earlier emotions: the anticipation, excitement, pride, and even the happy kinship with his friends. Now he felt only humiliation and terror.",
                                        "The Chief Elder waited until the uneasy applause subsided. Then she spoke again.",
                                        "\"I know,\" she said in her vibrant, gracious voice, \"that you are all concerned. That you feel I have made a mistake.\"",
                                        "She smiled. The community, relieved from its discomfort very slightly by her benign statement, seemed to breathe more easily. It was very silent.",
                                        "Jonas looked up.",
                                        "\"I have caused you anxiety,\" she said. \"I apologize to my community.\" Her voice flowed over the assembled crowd.",
                                        "\"We accept your apology,\" they all uttered together.",
                                        "\"Jonas,\" she said, looking down at him, \"I apologize to you in particular. I caused you anguish.\"",
                                        "\"I accept your apology,\" Jonas replied shakily.",
                                        "\"Please come to the stage now.\"",
                                        "Earlier that day, dressing in his own dwelling, he had practiced the kind of jaunty, self-assured walk that he hoped he could make to the stage when his turn came. All of that was forgotten now. He simply willed himself to stand, to move his feet that felt weighted and clumsy, to go forward, up the steps and across the platform until he stood at her side.",
                                        "Reassuringly she placed her arm across his tense shoulders.",
                                        "\"Jonas has not been assigned,\" she informed the crowd, and his heart sank.",
                                        "Then she went on. \"Jonas has been selected.\"",
                                        "He blinked. What did that mean? He felt a collective, questioning stir from the audience. They, too, were puzzled.",
                                        "In a firm, commanding voice she announced, \"Jonas has been selected to be our next Receiver of Memory.\"",
                                        "Then he heard the gasp--the sudden intake of breath, drawn sharply in astonishment, by each of the seated citizens. He saw their faces; the eyes widened in awe.",
                                        "And still he did not understand.",
                                        "\"Such a selection is very, very rare,\" the Chief Elder told the audience. \"Our community has only one Receiver. It is he who trains his successor.\"",
                                        "\"We have had our current Receiver for a very long time,\" she went on. Jonas followed her eyes and saw that she was looking at one of the Elders. The Committee of Elders was sitting together in a group; and the Chief Elder's eyes were now on one who sat in the midst but seemed oddly separate from them. It was a man Jonas had never noticed before, a bearded man with pale eyes. He was watching Jonas intently.",
                                        "\"We failed in our last selection,\" the Chief Elder said solemnly. \"It was ten years ago, when Jonas was just a toddler. I will not dwell on the experience because it causes us all terrible discomfort.\"",
                                        "Jonas didn't know what she was referring to, but he could sense the discomfort of the audience. They shifted uneasily in their seats.",
                                        "\"We have not been hasty this time,\" she continued. \"We could not afford another failure.\"",
                                        "\"Sometimes,\" she went on, speaking now in a lighter tone, relaxing the tension in the Auditorium, \"we are not entirely certain about the Assignments, even after the most painstaking observations. Sometimes we worry that the one assigned might not develop, through training, every attribute necessary. Elevens are still children, after all. What we observe as playfulness and patience-- the requirements to become Nurturer--could, with maturity, be revealed as simply foolishness and indolence. So we continue to observe during training, and to modify behavior when necessary.\"",
                                        "\"But the Receiver-in-training cannot be observed, cannot be modified. That is stated quite clearly in the rules. He is to be alone, apart, while he is prepared by the current Receiver for the job which is the most honored in our community.\"",
                                        "\"Alone? Apart? Jonas listened with increasing unease. \"Therefore the selection must be sound. It must be a unanimous choice of the Committee. They can have no doubts, however fleeting. If, during the process, an Elder reports a dream of uncertainty, that dream has the power to set a candidate aside instantly.\"",
                                        "\"Jonas was identified as a possible Receiver many years ago. We have observed him meticulously. There were no dreams of uncertainty.\"",
                                        "\"He has shown all of the qualities that a Receiver must have.\"",
                                        "With her hand still firmly on his shoulder, the Chief Elder listed the qualities.",
                                        "\"Intelligence,\" she said. \"We are all aware that Jonas has been a top student throughout his school days.\"",
                                        "\"Integrity,\" she said next. \"Jonas has, like all of us, committed minor transgressions.\" She smiled at him. \"We expect that. We hoped, also, that he would present himself promptly for chastisement, and he has always done so.\"",
                                        "\"Courage,\" she went on. \"Only one of us here today has ever undergone the rigorous training required of a Receiver. He, of course, is the most important member of the Committee: the current Receiver. It was he who reminded us, again and again, of the courage required.\"",
                                        "\"Jonas,\" she said, turning to him, but speaking in a voice that the entire community could hear, \"the training required of you involves pain. Physical pain.\"",
                                        "He felt fear flutter within him.",
                                        "\"You have never experienced that. Yes, you have scraped your knees in falls from your bicycle. Yes, you crushed your finger in a door last year.\"",
                                      "Jonas nodded, agreeing, as he recalled the incident, and its accompanying misery.",
                                      "\"But you will be faced, now,\" she explained gently, \"with pain of a magnitude that none of us here can comprehend because it is beyond our experience. The Receiver himself was not able to describe it, only to remind us that you would be faced with it, that you would need immense courage. We cannot prepare you for that.\"",
                                      "\"But we feel certain that you are brave,\" she said to him. He did not feel brave at all. Not now.
\"The fourth essential attribute,\" the Chief Elder said, \"is wisdom. Jonas has not yet acquired that. The acquisition of wisdom will come through his training.\"",
                                      "\"We are convinced that Jonas has the ability to acquire wisdom. That is what we looked for.\"",
                                      "\"Finally, The Receiver must have one more quality, and it is one which I can only name, but not describe. I do not understand it. You members of the community will not understand it, either. Perhaps Jonas will, because the current Receiver has told us that Jonas already has this quality. He calls it the Capacity to See Beyond.\"",
                                      "The Chief Elder looked at Jonas with a question in her eyes. The audience watched him, too. They were silent.",
                                      "For a moment he froze, consumed with despair. He didn't have it, the whatever-she-had-said. He didn't know what it was. Now was the moment when he would have to confess, to say, \"No, I don't. I can't,\" and throw himself on their mercy, ask their forgiveness, to explain that he had been wrongly chosen, that he was not the right one at all.",
                                      "But when he looked out across the crowd, the sea of faces, the thing happened again. The thing that had happened with the apple.",
                                      "They changed.",
                                      "He blinked, and it was gone. His shoulders straightened slightly. Briefly he felt a tiny sliver of sureness for the first time.",
                                      "She was still watching him. They all were.",
                                      "\"I think it's true,\" he told the Chief Elder and the community. \"I don't understand it yet. I don't know what it is. But sometimes I see something. And maybe it's beyond.\"",
                                      "She took her arm from his shoulders.",
                                      "\"Jonas,\" she said, speaking not to him alone but to the entire community of which he was a part, \"you will be trained to be our next Receiver of Memory. We thank you for your childhood.\"",
                                      "Then she turned and left the stage, left him there alone, standing and facing the crowd, which began spontaneously the collective murmur of his name.",
                                      "\"Jonas.\" It was a whisper at first: hushed, barely audible. \"Jonas. Jonas.\"",
                                      "Then louder, faster. \"JONAS. JONAS. JONAS.\" With the chant, Jonas knew, the community was accepting him and his new role, giving him life, the way they had given it to the newchild Caleb. His heart swelled with gratitude and pride.",
                                      "But at the same time he was filled with fear. He did not know what his selection meant. He did not know what he was to become.",
                                      "Or what would become of him."))

```

##Chapter 9
```{r ch9}
annotate_ch9 <- cnlp_annotate(input = c("Now, for the first time in his twelve years of life, Jonas felt separate, different. He remembered what the Chief Elder had said: that his training would be alone and apart.",
                                         "But his training had not yet begun and already, upon leaving the Auditorium, he felt the apartness. Holding the folder she had given him, he made his way through the throng, looking for his family unit and for Asher. People moved aside for him. They watched him. He thought he could hear whispers.",
                                         "\"Ash!\" he called, spotting his friend near the rows of bicycles. \"Ride back with me?\"",
                                         "\"Sure.\" Asher smiled, his usual smile, friendly and familiar. But Jonas felt a moment of hesitation from his friend, an uncertainty.",
                                         "\"Congratulations,\" Asher said.",
                                         "\"You too,\" Jonas replied. \"It was really funny, when she told about the smacks. You got more applause than almost anybody else.\"",
                                         "The other new Twelves clustered nearby, placing their folders carefully into the carrying containers on the backs of the bikes. In each dwelling tonight they would be studying the instructions for the beginning of their training. Each night for years the children had memorized the required lessons for school, often yawning with boredom. Tonight they would all begin eagerly to memorize the rules for their adult Assignments.",
                                         "\"Congratulations, Asher!\" someone called. Then that hesitation again. \"You too, Jonas!\"",
                                         "Asher and Jonas responded with congratulations to their groupmates. Jonas saw his parents watching him from the place where their own bicycles were waiting. Lily had already been strapped into her seat.",
                                         "He waved. They waved back, smiling, but he noticed that Lily was watching him solemnly, her thumb in her mouth.",
                                         "He rode directly to his dwelling, exchanging only small jokes and unimportant remarks with Asher.",
                                         "\"See you in the morning, Recreation Director!\" he called, dismounting by his door as Asher continued on.",
                                         "\"Right! See you!\" Asher called back. Once again, there was just a moment when things weren't quite the same, weren't quite as they had always been through the long friendship. Perhaps he had imagined it. Things couldn't change, with Asher.",
                                         "The evening meal was quieter than usual. Lily chattered about her plans for volunteer work; she would begin, she said, at the Nurturing Center, since she was already an expert at feeding Gabriel.",
                                         "\I know,\" she added quickly, when her father gave her a warning glance, \"I won't mention his name. I know I'm not supposed to know his name.\"",
                                         "\I can't wait for tomorrow to come,\" she said happily.",
                                         "Jonas sighed uneasily. \"I can,\" he muttered.",
                                         "\"You've been greatly honored,\" his mother said. \"Your father and I are very proud.\"",
                                         "\"It's the most important job in the community,\" Father said.",
                                         "\"But just the other night, you said that the job of making Assignments was the most important!\"",
                                         "Mother nodded. \"This is different. It's not a job, really. I never thought, never expected--\" She paused. \"There's only one Receiver.\"",
                                         "\"But the Chief Elder said that they had made a selection before, and that it failed. What was she talking about?\"",
                                         "Both of his parents hesitated. Finally his father described the previous selection. \"It was very much as it was today, Jonas- the same suspense, as one Eleven had been passed over when the Assignments were given. Then the announcement, when they singled out the one--\"",
                                         "Jonas interrupted. \"What was his name?\"",
                                         "His mother replied, \"Her, not his. It was a female. But we are never to speak the name, or to use it again for a newchild.\"",
                                         "Jonas was shocked. A name designated Not-to-Be-Spoken indicated the highest degree of disgrace.",
                                         "\"What happened to her?\" he asked nervously.",
                                         "But his parents looked blank. \"We don't know,\" his father said uncomfortably. \"We never saw her again.\"",
                                         "A silence fell over the room. They looked at each other. Finally his mother, rising from the table, said, \"You've been greatly honored, Jonas. Greatly honored.\"",
                                         "Alone in his sleepingroom, prepared for bed, Jonas opened his folder at last. Some of the other Twelves, he had noticed, had been given folders thick with printed pages. He imagined Benjamin, the scientific male in his group, beginning to read pages of rules and instructions with relish. He pictured Fiona smiling her gentle smile as she bent over the lists of duties and methods that she would be required to learn in the days to come.",
                                         "But his own folder was startlingly close to empty. Inside there was only a single printed sheet. He read it twice.",
                                         "JONAS",
                                         "RECEIVER OF MEMORY",
                                         "1.	Go immediately at the end of school hours each day to the Annex entrance behind the House of the Old and present yourself to the attendant.",
                                         "2.	Go immediately to your dwelling at the conclusion of Training Hours each day.",
                                         "3.	From this moment you are exempted from rules governing rudeness. You may ask any question of any citizen and you will receive answers.",
                                         "4.	Do not discuss your training with any other member of the community, including parents and Elders.",
                                         "5.	From this moment you are prohibited from dream-telling.",
                                         "6.	Except for illness or injury unrelated to your training, do not apply for any medication.",
                                         "7.	You are not permitted to apply for release.",
                                         "8.	You may lie.",
                                         "Jonas was stunned. What would happen to his friendships? His mindless hours playing ball, or riding his bike along the river? Those had been happy and vital times for him. Were they to be completely taken from him, now? The simple logistic instructions-- where to go, and when--were expected. Every Twelve had to be told, of course, where and how and when to report for training. But he was a little dismayed that his schedule left no time, apparently, for recreation.",
                                         "The exemption from rudeness startled him. Reading it again, however, he realized that it didn't compel him to be rude; it simply allowed him the option. He was quite certain he would never take advantage of it. He was so completely, so thoroughly accustomed to courtesy within the community that the thought of asking another citizen an intimate question, of calling someone's attention to an area of awkwardness, was unnerving.",
                                         "The prohibition of dream-telling, he thought, would not be a real problem. He dreamed so rarely that the dream-telling did not come easily to him anyway, and he was glad to be excused from it. He wondered briefly, though, how to deal with it at the morning meal. What if he did dream--should he simply tell his family unit, as he did so often, anyway, that he hadn't? That would be a lie. Still, the final rule said ... well, he wasn't quite ready to think about the final rule on the page.",
                                         "The restriction of medication unnerved him. Medication was always available to citizens, even to children, through their parents. When he had crushed his finger in the door, he had quickly, gasping into the speaker, notified his mother; she had hastily requisitioned relief- of-pain medication which had promptly been delivered to his dwelling. Almost instantly the excruciating pain in his hand had diminished to the throb which was, now, all he could recall of the experience.",
                                         "Re-reading rule number 6, he realized that a crushed finger fell into the category of \"unrelated to training.\" So if it ever happened again--and he was quite certain it wouldn't; he had been very careful near heavy doors since the accident!--he could still receive medication.",
                                         "The pill he took now, each morning, was also unrelated to training. So he would continue to receive the pill.",
                                         "But he remembered uneasily what the Chief Elder had said about the pain that would come with his training. She had called it indescribable.",
                                         "Jonas swallowed hard, trying without success to imagine what such pain might be like, with no medication at all. But it was beyond his comprehension.",
                                         "He felt no reaction to rule number 7 at all. It had never occurred to him that under any circumstances, ever, he might apply for release.",
                                         "Finally he steeled himself to read the final rule again. He had been trained since earliest childhood, since his earliest learning of language, never to lie. It was an integral part of the learning of precise speech. Once, when he had been a Four, he had said, just prior to the midday meal at school, \"I'm starving.\"",
                                         "Immediately he had been taken aside for a brief private lesson in language precision. He was not starving, it was pointed out. He was hungry. No one in the community was starving, had ever been starving, would ever be starving. To say \"starving\" was to speak a lie. An unintentioned lie, of course. But the reason for precision of language was to ensure that unintentional lies were never uttered. Did he understand that? they asked him. And he had.",
                                         "He had never, within his memory, been tempted to lie. Asher did not lie. Lily did not lie. His parents did not lie. No one did. Unless ...",
                                         "Now Jonas had a thought that he had never had before. This new thought was frightening. What if others--adult-- had, upon becoming Twelves, received in their instructions the same terrifying sentence?",
                                         "What if they had all been instructed: You may lie?",
                                         "His mind reeled. Now, empowered to ask questions of utmost rudeness and promised answers--he could, conceivably (though it was almost unimaginable), ask someone, some adult, his father perhaps: \"Do you lie?\"",
                                         "But he would have no way of knowing if the answer he received were true."))

```

##Chapter 10
```{r ch10}
annotate_ch10 <- cnlp_annotate(input = c("\"I go in here, Jonas,\" Fiona told him when they reached the front door of the House of the Old after parking their bicycles in the designated area.",
                                         "\"I don't know why I'm nervous,\" she confessed. \"I've been here so often before.\" She turned her folder over in her hands.",
                                         "\"Well, everything's different now,\" Jonas reminded her.",
                                         "\"Even the nameplates on our bikes,\" Fiona laughed. During the night the nameplate of each new Twelve had been removed by the Maintenance Crew and replaced with the style that indicated citizen-in-training.",
                                         "\"I don't want to be late,\" she said hastily, and started up the steps. \"If we finish at the same time, I'll ride home with you.\"",
                                         "Jonas nodded, waved to her, and headed around the building toward the Annex, a small wing attached to the back. He certainly didn't want to be late for his first day of training, either.",
                                         "The Annex was very ordinary, its door unremarkable. He reached for the heavy handle, then noticed a buzzer on the wall. So he buzzed instead.",
                                         "\"Yes?\" The voice came through a small speaker above the buzzer. \"It's, uh, Jonas. I'm the new--I mean--\" \"Come in.\" A click indicated that the door had been unlatched. The lobby was very small and contained only a desk at which a female Attendant sat working on some papers. She looked up when he entered; then, to his surprise, she stood. It was a small thing, the standing; but no one had ever stood automatically to acknowledge Jonas's presence before.",
                                         "\"Welcome, Receiver of Memory,\" she said respectfully.",
                                         "\"Oh, please,\" he replied uncomfortably. \"Call me Jonas.\"",
                                         "She smiled, pushed a button, and he heard a click that unlocked the door to her left. \"You may go right on in,\" she told him.",
                                         "Then she seemed to notice his discomfort and to realize its origin. No doors in the community were locked, ever. None that Jonas knew of, anyway.",
                                         "\"The locks are simply to insure The Receiver's privacy because he needs concentration,\" she explained. \"It would be difficult if citizens wandered in, looking for the Department of Bicycle Repair, or something.\"",
                                         "Jonas laughed, relaxing a little. The woman seemed very friendly, and it was true--in fact it was a joke throughout the community--that the Department of Bicycle Repair, an unimportant little office, was relocated so often that no one ever knew where it was.",
                                         "\"There is nothing dangerous here,\" she told him.",
                                         "\"But,\" she added, glancing at the wall clock, \"he doesn't like to be kept waiting.\"",
                                         "Jonas hurried through the door and found himself in a comfortably furnished living area. It was not unlike his own family unit's dwelling. Furniture was standard throughout the community: practical, sturdy, the function of each piece clearly defined. A bed for sleeping. A table for eating. A desk for studying.",
                                         "All of those things were in this spacious room, though each was slightly different from those in his own dwelling. The fabrics on the upholstered chairs and sofa were slightly thicker and more luxurious; the table legs were not straight like those at home, but slender and curved, with a small carved decoration at the foot. The bed, in an alcove at the far end of the room, was draped with a splendid cloth embroidered over its entire surface with intricate designs.",
                                         "But the most conspicuous difference was the books. In his own dwelling, there were the necessary reference volumes that each household contained: a dictionary, and the thick community volume which contained descriptions of every office, factory, building, and committee. And the Book of Rules, of course.",
                                         "The books in his own dwelling were the only books that Jonas had ever seen. He had never known that other books existed.",
                                         "But this room's walls were completely covered by bookcases, filled, which reached to the ceiling. There must have been hundreds-- perhaps thousands--of books, their titles embossed in shiny letters.",
                                         "Jonas stared at them. He couldn't imagine what the thousands of pages contained. Could there be rules beyond the rules that governed the community? Could there be more descriptions of offices and factories and committees?",
                                         "He had only a second to look around because he was aware that the man sitting in a chair beside the table was watching him. Hastily he moved forward, stood before the man, bowed slightly, and said, \"I'm Jonas.\"",
                                         "\"I know. Welcome, Receiver of Memory.\"",
                                         "Jonas recognized the man. He was the Elder who had seemed separate from the others at the Ceremony, though he was dressed in the same special clothing that only Elders wore.",
                                         "Jonas looked self-consciously into the pale eyes that mirrored his own.",
                                         "\"Sir, I apologize for my lack of understanding.....\"",
                                         "He waited, but the man did not give the standard accepting-of- apology response.",
                                         "After a moment, Jonas went on, \"But I thought--I mean I think,\" he corrected, reminding himself that if precision of language were ever to be important, it was certainly important now, in the presence of this man, \"that you are the Receiver of Memory. I'm only, well, I was only assigned, I mean selected, yesterday. I'm not anything at all. Not yet.\"",
                                         "The man looked at him thoughtfully, silently. It was a look that combined interest, curiosity, concern, and perhaps a little sympathy as well.",
                                         "Finally he spoke. \"Beginning today, this moment, at least to me, you are the Receiver.\"",
                                         "\"I have been the Receiver for a long time. A very, very long time. You can see that, can't you?\"",
                                         "Jonas nodded. The man was wrinkled, and his eyes, though piercing in their unusual lightness, seemed tired.",
                                         "The flesh around them was darkened into shadowed circles.",
                                         "\"I can see that you are very old,\" Jonas responded with respect. The Old were always given the highest respect.",
                                         "The man smiled. He touched the sagging flesh on his own face with amusement. \"I am not, actually, as old as I look,\" he told Jonas. \"This job has aged me. I know I look as if I should be scheduled for release very soon. But actually I have a good deal of time left.\"",
                                         "\"I was pleased, though, when you were selected. It took them a long time. The failure of the previous selection was ten years ago, and my energy is starting to diminish. I need what strength I have remaining for your training. We have hard and painful work to do, you and I.\"",
                                         "\"Please sit down,\" he said, and gestured toward the nearby chair. Jonas lowered himself onto the soft cushioned seat.",
                                         "The man dosed his eyes and continued speaking. \"When I became a Twelve, I was selected, as you were. I was frightened, as I'm sure you are.\" He opened his eyes for a moment and peered at Jonas, who nodded.",
                                         "The eyes closed again. \"I came to this very room to begin my training. It was such a long time ago.\"",
                                         "\"The previous Receiver seemed just as old to me as I do to you. He was just as tired as I am today.\"",
                                         "He sat forward suddenly, opened his eyes, and said, \"You may ask questions. I have so little experience in describing this process. It is forbidden to talk of it.\"",
                                         "\"I know, sir. I have read the instructions,\" Jonas said.",
                                         "\"So I may neglect to make things as clear as I should.\" The man chuckled. \"My job is important and has enormous honor. But that does not mean I am perfect, and when I tried before to train a successor, I failed. Please ask any questions that will help you.\"",
                                         "In his mind, Jonas had questions. A thousand. A million questions. As many questions as there were books lining the walls. But he did not ask one, not yet.",
                                         "The man sighed, seeming to put his thoughts in order. Then he spoke again. \"Simply stated,\" he said, \"although it's not really simple at all, my job is to transmit to you all the memories I have within me. Memories of the past.\"",
                                         "\"Sir,\" Jonas said tentatively, \"I would be very interested to hear the story of your life, and to listen to your memories.\",
                                         "\"I apologize for interrupting,\" he added quickly.",
                                         "The man waved his hand impatiently. \"No apologies in this room. We haven't time.\"",
                                         "\"Well,\" Jonas went on, uncomfortably aware that he might be interrupting again, \"I am really interested, I don't mean that I'm not. But I don't exactly understand why it's so important. I could do some adult job in the community, and in my recreation time I could come and listen to the stories from your childhood. I'd like that. Actually,\" he added, \"I've done that already, in the House of the Old. The Old like to tell about their childhoods, and it's always fun to listen.\"",
                                         "The man shook his head. \"No, no,\" he said. \"I'm not being clear. It's not my past, not my childhood that I must transmit to you.\"",
                                         "He leaned back, resting his head against the back of the upholstered chair. \"It's the memories of the whole world,\" he said with a sigh. \"Before you, before me, before the previous Receiver, and generations before him.\"",
                                         "Jonas frowned. \"The whole world?\" he asked. \"I don't understand. Do you mean not just us? Not just the community? Do you mean Elsewhere, too?\" He tried, in his mind, to grasp the concept. \"I'm sorry, sir. I don't understand exactly. Maybe I'm not smart enough. I don't know what you mean when you say 'the whole world' or 'generations before him.' I thought there was only us. I thought there was only now.\"",
                                         "\"There's much more. There's all that goes beyond--all that is Elsewhere--and all that goes back, and back, and back. I received all of those, when I was selected. And here in this room, all alone, I re-experience them again and again. It is how wisdom comes. And how we shape our future.\"",
                                         "He rested for a moment, breathing deeply. \"I am so weighted with them,\" he said.",
                                         "Jonas felt a terrible concern for the man, suddenly.",
                                         "\"It's as if ... \" The man paused, seeming to search his mind for the right words of description. \"It's like going downhill through deep snow on a sled,\" he said, finally. \"At first it's exhilarating: the speed; the sharp, clear air; but then the snow accumulates, builds up on the runners, and you slow, you have to push hard to keep going, and--\"",
                                         "He shook his head suddenly, and peered at Jonas. \"That meant nothing to you, did it?\" he asked.",
                                         "Jonas was confused. \"I didn't understand it, sir.\"",
                                         "\"Of course you didn't. You don't know what snow is, do you?\"",
                                         "Jonas shook his head.",
                                         "\"Or a sled? Runners?\"",
                                         "\"No, sir,\" Jonas said.",
                                         "\"Downhill? The term means nothing to you?\"",
                                         "\"Nothing, sir.\"",
                                         "\"Well, it's a place to start. I'd been wondering how to begin. Move to the bed, and lie face down. Remove your tunic first.\"",
                                         "Jonas did so, a little apprehensively. Beneath his bare chest, he felt the soft folds of the magnificent cloth that covered the bed. He watched as the man rose and moved first to the wall where the speaker was. It was the same sort of speaker that occupied a place in every dwelling, but one thing about it was different. This one had a switch, which the man deftly snapped to the end that said OFF.",
                                         "Jonas almost gasped aloud. To have the power to turn the speakeroff! It was an astonishing thing.",
                                         "Then the man moved with surprising quickness to the corner where the bed was. He sat on a chair beside Jonas, who was motionless, waiting for what would happen next.",
                                         "\"Close your eyes. Relax. This will not be painful.\"",
                                         "Jonas remembered that he was allowed, that he had even been encouraged, to ask questions. \"What are you going to do, sir?\" he asked, hoping that his voice didn't betray his nervousness.",
                                         "\"I am going to transmit the memory of snow,\" the old man said, and placed his hands on Jonas's bare back."))

```

##Chapter 11
```{r ch11}

annotate_ch11 <- cnlp_annotate(input = c("Jonas felt nothing unusual at first. He felt only the light touch of the old man's hands on his back.",
                                         "He tried to relax, to breathe evenly. The room was absolutely silent, and for a moment Jonas feared that he might disgrace himself now, on the first day of his training, by falling asleep.",
                                         "Then he shivered. He realized that the touch of the hands felt, suddenly, cold. At the same instant, breathing in, he felt the air change, and his very breath was cold. He licked his lips, and in doing so, his tongue touched the suddenly chilled air.",
                                         "It was very startling; but he was not at all frightened, now. He was filled with energy, and he breathed again, feeling the sharp intake of frigid air. Now, too, he could feel cold air swirling around his entire body. He felt it blow against his hands where they lay at his sides, and over his back.",
                                         "The touch of the man's hands seemed to have disappeared.",
                                         "Now he became aware of an entirely new sensation: pinpricks? No, because they were soft and without pain. Tiny, cold, featherlike feelings peppered his body and face. He put out his tongue again, and caught one of the dots of cold upon it. It disappeared from his awareness instantly; but he caught another, and another. The sensation made him smile.",
                                         "One part of his consciousness knew that he was still lying there, on the bed, in the Annex room. Yet another, separate part of his being was upright now, in a sitting position, and beneath him he could feel that he was not on the soft decorated bedcovering at all, but rather seated on a flat, hard surface. His hands now held (though at the same time they were still motionless at his sides) a rough, damp rope.",
                                         "And he could see, though his eyes were closed. He could see a bright, whirling torrent of crystals in the air around him, and he could see them gather on the backs of his hands, like cold fur.",
                                         "His breath was visible.",
                                         "Beyond, through the swirl of what he now, somehow, perceived was the thing the old man had spoken of-- snow-- he could look out and down a great distance. He was up high someplace. The ground was thick with the furry snow, but he sat slightly above it on a hard, flat object.",
                                         "Sled, he knew abruptly. He was sitting on a thing called sled. And the sled itself seemed to be poised at the top of a long, extended mound that rose from the very land where he was. Even as he thought the word \"mound,\" his new consciousness told him hill.",
                                         "Then the sled, with Jonas himself upon it, began to move through the snowfall, and he understood instantly that now he was going downhill. No voice made an explanation. The experience explained itself to him.",
                                         "His face cut through the frigid air as he began the descent, moving through the substance called snow on the vehicle called sled, which propelled itself on what he now knew without doubt to be runners.",
                                         "Comprehending all of those things as he sped downward, he was free to enjoy the breathless glee that overwhelmed him: the speed, the clear cold air, the total silence, the feeling of balance and excitement and peace.",
                                         "Then, as the angle of incline lessened, as the mound--the hill-- flattened, nearing the bottom, the sled's forward motion slowed. The snow was piled now around it, and he pushed with his body, moving it forward, not wanting the exhilarating ride to end.",
                                         "Finally the obstruction of the piled snow was too much for the thin runners of the sled, and he came to a stop. He sat there for a moment, panting, holding the rope in his cold hands. Tentatively he opened his eyes--not his snow-hill-sled eyes, for they had been open throughout the strange ride. He opened his ordinary eyes, and saw that he was still on the bed, that he had not moved at all.",
                                         "The old man, still beside the bed, was watching him. \"How do you feel?\" he asked.",
                                         "Jonas sat up and tried to answer honestly. \"Surprised,\" he said, after a moment.",
                                         "The old man wiped his forehead with his sleeve. \"Whew,\" he said. \"It was exhausting. But you know, even transmitting that tiny memory to you--I think it lightened me just a little.\"",
                                         "\"Do you mean--you did say I could ask questions?\" The man nodded, encouraging his question.",
                                         "\"Do you mean that now you don't have the memory of it--of that ride on the sled--anymore?\"",
                                         "\"That's right. A little weight off this old body.\"",
                                         "\"But it was such fun! And now you don't have it anymore! I took it from you!\"",
                                         "But the old man laughed. \"All I gave you was one ride, on one sled, in one snow, on one hill. I have a whole world of them in my memory. I could give them to you one by one, a thousand times, and there would still be more.\"",
                                         
                                         "\"Are you saying that I--I mean we--could do it again?\" Jonas asked. \"I'd really like to. I think I could steer, by pulling the rope. I didn't try this time, because it was so new.\"",
                                         "The old man, laughing, shook his head. \"Maybe another day, for a treat. But there's no time, really, just to play. I only wanted to begin by showing you how it works.\"",
                                         "\"Now,\" he said, turning businesslike, \"lie back down. I want to --\"",
                                         "Jonas did. He was eager for whatever experience would come next. But he had, suddenly, so many questions.",
                                         "\"Why don't we have snow, and sleds, and hills?\" he asked. \"And when did we, in the past? Did my parents have sleds when they were young? Did you?\"",
                                         "The old man shrugged and gave a short laugh. \"No,\" he told Jonas. \"It's a very distant memory. That's why it was so exhausting--I had to tug it forward from many generations back. It was given to me when I was a new Receiver, and the previous Receiver had to pull it through a long time period, too.\"",
                                         "\"But what happened to those things? Snow, and the rest of it?\"",
                                         "\"Climate Control. Snow made growing food difficult, limited the agricultural periods. And unpredictable weather made transportation almost impossible at times. It wasn't a practical thing, so it became obsolete when we went to Sameness.\"",
                                         "\"And hills, too,\" he added. \"They made conveyance of goods unwieldy. Trucks; buses. Slowed them down. So--\" He waved his hand, as if a gesture had caused hills to disappear. \"Sameness,\" he concluded.",
                                         "Jonas frowned. \"I wish we had those things, still. Just now and then.\"",
                                         "The old man smiled. \"So do I,\" he said. \"But that choice is not ours.\" \"But sir,\" Jonas suggested, \"since you have so much power--\"",
                                         "The man corrected him. \"Honor,\" he said firmly. \"I have great honor. So will you. But you will find that that is not the same as power.\"",
                                         "\"Lie quietly now. Since we've entered into the topic of climate, let me give you something else. And this time I'm not going to tell you the name of it, because I want to test the receiving. You should be able to perceive the name without being told. I gave away snow and sled and downhill and runners by telling them to you in advance.\"",
                                         "Without being instructed, Jonas closed his eyes again. He felt the hands on his back again. He waited.",
                                         "Now it came more quickly, the feelings. This time the hands didn't become cold, but instead began to feel warm on his body. They moistened a little. The warmth spread, extending across his shoulders, up his neck, onto the side of his face. He could feel it through his clothed parts, too: a pleasant, all-over sensation; and when he licked his lips this time, the air was hot and heavy.",
                                        "He didn't move. There was no sled. His posture didn't change. He was simply alone someplace, out of doors, lying down, and the warmth came from far above. It was not as exciting as the ride through the snowy air; but it was pleasurable and comforting.",
                                        "Suddenly he perceived the word for it: sunshine. He perceived that it came from the sky. Then it ended.",
                                        "\"Sunshine,\" he said aloud, opening his eyes.",
                                        "\"Good. You did get the word. That makes my job easier. Not so much explaining.\"",
                                        "\"And it came from the sky.\" \"That's right,\" the old man said. \"Just the way it used to.\"",
                                        "\"Before Sameness. Before Climate Control,\" Jonas added.",
                                        "The man laughed. \"You receive well, and learn quickly. I'm very pleased with you. That's enough for today, I think. We're off to a good start.\"",
                                        "There was a question bothering Jonas. \"Sir,\" he said, \"The Chief Elder told me--she told everyone--and you told me, too, that it would be painful. So I was a little scared. But it didn't hurt at all. I really enjoyed it.\" He looked quizzically at the old man.",
                                        "The man sighed. \"I started you with memories of pleasure. My previous failure gave me the wisdom to do that.\" He took a few deep breaths.\"Jonas,\" he said, \"it will be painful. But it need not be painful yet.\"",
                                        "\"I'm brave. I really am.\" Jonas sat up a little straighter. ",
                                        "The old man looked at him for a moment. He smiled.",
                                        "\"I can see that,\" he said. \"Well, since you asked the question--I think I have enough energy for one more transmission.\"",
                                        "\"Lie down once more. This will be the last today.\"",
                                        "Jonas obeyed cheerfully. He dosed his eyes, waiting, and felt the hands again; then he felt the warmth again, the sunshine again, coming from the sky of this other consciousness that was so new to him. This time, as he lay basking in the wonderful warmth, he felt the passage of time. His real self was aware that it was only a minute or two; but his other, memory-receiving self felt hours pass in the sun. His skin began to sting. Restlessly he moved one arm, bending it, and felt a sharp pain in the crease of his inner arm at the elbow.",
                                        "\"Ouch,\" he said loudly, and shifted on the bed. \"Owwww,\" he said, wincing at the shift, and even moving his mouth to speak made his face hurt.",
                                        "He knew there was a word, but the pain kept him from grasping it.",
                                        "Then it ended. He opened his eyes, wincing with discomfort. \"It hurt,\" he told the man, \"and I couldn't get the word for it.\"",
                                        "\"It was sunburn,\" the old man told him.",
                                        "\"It hurt a lot,\" Jonas said, \"but I'm glad you gave it to me. It was interesting. And now I understand better, what it meant, that there would be pain.\"",
                                        "The man didn't respond. He sat silently for a second. Finally he said, \"Get up, now. It's time for you to go home.\"",
                                        "They both walked to the center of the room. Jonas put his tunic back on. \"Goodbye, sir,\" he said. \"Thank you for my first day.\"",
                                        "The old man nodded to him. He looked drained, and a little sad. \"Sir?\" Jonas said shyly. \"Yes? Do you have a question?\"",
                                        "\"It's just that I don't know your name. I thought you were The Receiver, but you say that now I'm The Receiver. So I don't know what to call you.\"",
                                        "The man had sat back down in the comfortable upholstered chair. He moved his shoulders around as if to ease away an aching sensation. He seemed terribly weary.",
                                        "\"Call me The Giver,\" he told Jonas."))

```

```{r ch11plot}

ch11 <- annotate_ch11$token

#version 1
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Quantity",
       y = "Parts of Speech")

#version 2
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 11 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch11 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 11 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch11, head)


```
The above bar chart displays a count of each part of speech within Chapter 11 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 




##Chapter 23

```{r ch23}
annotate_ch23 <- cnlp_annotate(input = c("Jonas felt more and more certain that the destination lay ahead of him, very near now in the night that was approaching. None of his senses confirmed it. He saw nothing ahead except the endless ribbon of road unfolding in twisting narrow curves. He heard no sound ahead.",
                                         "Yet he felt it: felt that Elsewhere was not far away. But he had little hope left that he would be able to reach it. His hope diminished further when the sharp, cold air began to blur and thicken with swirling white.",
                                         "Gabriel, wrapped in his inadequate blanket, was hunched, shivering, and silent in his little seat. Jonas stopped the bike wearily, lifted the child down, and realized with heartbreak how cold and weak Gabe had become.",
                                         "Standing in the freezing mound that was thickening around his numb feet, Jonas opened his own tunic, held Gabriel to his bare chest, and tied the torn and dirty blanket around them both. Gabriel moved feebly against him and whimpered briefly into the silence that surrounded them.",
                                         "Dimly, from a nearly forgotten perception as blurred as the substance itself, Jonas recalled what the whiteness was.",
                                         "\"It's called snow, Gabe,\" Jonas whispered. \"Snowflakes. They fall down from the sky, and they're very beautiful.\"",
                                         "There was no response from the child who had once been so curious and alert. Jonas looked down through the dusk at the little head against his chest. Gabriel's curly hair was matted and filthy, and there were tearstains outlined in dirt on his pale cheeks. His eyes were closed. As Jonas watched, a snowflake drifted down and was caught briefly for a moment's sparkle in the tiny fluttering eyelashes.",
                                         "Wearily he remounted the bicycle. A steep hill loomed ahead. In the best of conditions, the hill would have been a difficult, demanding ride. But now the rapidly deepening snow obscured the narrow road and made the ride impossible. His front wheel moved forward imperceptibly as he pushed on the pedals with his numb, exhausted legs. But the bicycle stopped. It would not move.",
                                         "He got off and let it drop sideways into the snow. For a moment he thought how easy it would be to drop beside it himself, to let himself and Gabriel slide into the softness of snow, the darkness of night, the warm comfort of sleep. But he had come this far. He must try to go on.",
                                         "The memories had fallen behind him now, escaping from his protection to return to the people of his community. Were there any left at all? Could he hold onto a last bit of warmth? Did he still have the strength to Give? Could Gabriel still Receive?",
                                         "He pressed his hands into Gabriel's back and tried to remember sunshine. For a moment it seemed that nothing came to him, that his power was completely gone. Then flickered suddenly, and he felt tiny tongues of heat begin to creep across and into his frozen feet and legs. He felt his face begin to glow and the tense, cold skin of his arms and hands relax. For a fleeting second he felt that he wanted to keep it for himself, to let himself bathe in sunlight, unburdened by anything or anyone else.",
                                         "But the moment passed and was followed by an urge, a need, a passionate yearning to share the warmth with the one person left for him to love. Aching from the effort, he forced the memory of warmth into the thin, shivering body in his arms.",
                                         "Gabriel stirred. For a moment they both were bathed in warmth and renewed strength as they stood hugging each other in the blinding snow.",
                                         "Jonas began to walk up the hill.",
                                         "The memory was agonizingly brief. He had trudged no more than a few yards through the night when it was gone and they were cold again.",
                                         "But his mind was alert now. Warming himself ever so briefly had shaken away the lethargy and resignation and restored his will to survive. He began to walk faster on feet that he could no longer feel. But the hill was treacherously steep; he was impeded by the snow and his own lack of strength. He didn't make it very far before he stumbled and fell forward.",
                                         "On his knees, unable to rise, Jonas tried a second time. His consciousness grasped at a wisp of another warm memory, and tried desperately to hold it there, to enlarge it, and pass it into Gabriel. His spirits and strength lifted with the momentary warmth and he stood. Again, Gabriel stirred against him as he began to climb.",
                                         "But the memory faded, leaving him colder than before. If only he had had time to receive more warmth from The Giver before he escaped! Maybe there would be more left for him now. But there was no purpose in if-onlys. His entire concentration now had to be on moving his feet, warming Gabriel and himself, and going forward.",
                                         "He climbed, stopped, and warmed them both briefly again, with a tiny scrap of memory that seemed certainly to be all he had left.",
                                         "The top of the hill seemed so far away, and he did not know what lay beyond. But there was nothing left to do but continue. He trudged upward.",
                                         "As he approached the summit of the hill at last, something began to happen. He was not warmer; if anything, he felt more numb and more cold. He was not less exhausted; on the contrary, his steps were leaden, and he could barely move his freezing, tired legs.",
                                         "But he began, suddenly, to feel happy. He began to recall happy times. He remembered his parents and his sister. He remembered his friends, Asher and Fiona. He remembered The Giver.",
                                         "Memories of joy flooded through him suddenly.",
                                         "He reached the place where the hill crested and he could feel the ground under his snow-covered feet become level. It would not be uphill anymore.",
                                         "\"We're almost there, Gabriel,\" he whispered, feeling quite certain without knowing why. \"I remember this place, Gabe.\" And it was true. But it was not a grasping of a thin and burdensome recollection; this was different. This was something that he could keep. It was a memory of his own.",
                                         "He hugged Gabriel and rubbed him briskly, warming him, to keep him alive. The wind was bitterly cold. The snow swirled, blurring his vision. But somewhere ahead, through the blinding storm, he knew there was warmth and light.",
                                         "Using his final strength, and a special knowledge that was deep inside him, Jonas found the sled that was waiting for them at the top of the hill. Numbly his hands fumbled for the rope.",
                                         "He settled himself on the sled and hugged Gabe close. The hill was steep but the snow was powdery and soft, and he knew that this time there would be no ice, no fall, no pain. Inside his freezing body, his heart surged with hope.",
                                         "They started down.",
                                         "Jonas felt himself losing consciousness and with his whole being willed himself to stay upright atop the sled, clutching Gabriel, keeping him safe. The runners sliced through the snow and the wind whipped at his face as they sped in a straight line through an incision that seemed to lead to the final destination, the place that he had always felt was waiting, the Elsewhere that held their future and their past.",
                                         "He forced his eyes open as they went downward, downward, sliding, and all at once he could see lights, and he recognized them now. He knew they were shining through the windows of rooms, that they were the red, blue, and yellow lights that twinkled from trees in places where families created and kept memories, where they celebrated love.",
                                         "Downward, downward, faster and faster. Suddenly he was aware with certainty and joy that below, ahead, they were waiting for him; and that they were waiting, too, for the baby. For the first time, he heard something that he knew to be music. He heard people singing.",
                                         "Behind him, across vast distances of space and time, from the place he had left, he thought he heard music too. But perhaps it was only an echo."))

```

```{r ch23plot}

ch23 <- annotate_ch23$token

#version 1
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".") %>% 
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Quantity",
       y = "Parts of Speech")

#version 2
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col()+
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 23 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter")

#version 3 (final version)
ch23 %>% 
  count(upos) %>% 
  filter(upos != ".",
         upos != "X",
         upos != "PUNCT") %>% 
  mutate(upos = recode(upos,
                           "VERB" = "Verb",
                           "SCONJ" = "Subordinating Conjunction",
                           "PROPN" = "Proper Noun",
                           "PRON" = "Pronoun",
                           "PART" = "Particle",
                           "NUM" = "Numeral",
                           "NOUN" = "Noun",
                           "INTJ" = "Interjection",
                           "DET" = "Determiner",
                           "CCONJ" = "Coordinating Conjunction",
                           "AUX" = "Auxillary",
                           "ADV" = "Adverb",
                           "ADP" = "Adposition",
                           "ADJ" = "Adjective")) %>%
  mutate(upos = fct_reorder(upos, n)) %>%
  ggplot(aes(n, upos)) +
  geom_col() +
  labs(x = "Number of Occurances",
       y = "",
       title = "Parts of Speech Present in Chapter 23 of The Giver",
       subtitle = "Displays comparison of parts of speech throughout the chapter") +
  theme_minimal()

lapply(annotate_ch23, head)

```
The above bar chart displays a count of each part of speech within Chapter 23 of *The Giver*. I will be cleaning up the parts of speech categories, modifying the visualization, and adding color to decrease the cognitive load. 

##Comparison Across Chapters

```{r verbcomparison}

verbch1 <- ch1 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch1")

verbch11 <- ch11 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch11")

verbch23 <- ch23 %>%
  filter(upos == "VERB") %>%
  mutate(new_col = "ch23")

verbcombined <- rbind(verbch1, verbch11, verbch23) %>%
  rename(chapters = new_col) %>%
  select(upos, chapters)

#version 1
verbcombined %>% 
  count(chapters) %>% 
  ggplot(aes(n, chapters)) +
  geom_col() +
  scale_y_discrete(breaks = paste0("ch", c("1", "11", "23")),
                   labels = c("1", "11", "23")) +
  labs(x = "Number of verbs",
       y = "Chapter")

#version 2
verbcombined %>% 
  count(chapters) %>% 
  #mutate(chapters = fct_relevel(chapters, 
                                "1",
                                "11")) %>%
  ggplot(aes(n, chapters)) +
  geom_col() +
  scale_y_discrete(breaks = paste0("ch", c("1", "11", "23")),
                   labels = c("1", "11", "23")) +
  labs(x = "Number of verbs",
       y = "Chapter")

# Interesting finding!! Do you think this could be because  some chapters are longer than others (see my % of words comment from earlier), or this is a meaningful trend? I would love to see in your final version an explanation for why you chose these chapters to look at (i'm assuming beg, middle, and end)/ why you chose verbs to look at. If you want to use gganimate like I suggested, you could put these three plots one after the other and highlight the "verb" bar a color like red. That way you can see how verbs change across chapters in the context of other parts of speech, e.g., you might find that verbs change a much smaller or much larger amount than other parts of speech across chapters

# This graph is extremely effective at demonstrating this effect. I immediately understood what was going on without even needing a title. Beyond what you have here, I think it would be really cool to look specifically at what verbs are most common in the giver. You could present that data in a word cloud, which makes more prevalent words bigger. Since you're interested in change across the book, you could even look at how the verbs change in prevalence across the book 

```
The above graph displays the amount of verbs present in chapters 1, 11, and 23. There is a downward trend in the amount of verbs as you move through the book. 

*I really love this figure! It raises so many questions! I think the use of bar plots was particularly effective here. I'm curious about the possibilities of adding an additional variable to this figure - something like "complexity" or a different metric that shows whether the verbs change in quality as well as quantity over the course of the book. Also, I wonder whether this is impacted by a change in length (i.e., is chapter 1 super long?). That may be a fun aspect to play with in a future draft (or just with this data set)*

*I think you picked great forms for the data you picked, and did a really good job navigating complex data - NLP is tough stuff! I haven't poked around NLP done in R before, so definitely learned a lot examining your code!*
